---
title: "SI Text 3: Model selection"
subtitle: "Beaked and golden redfish: biomass density"
# author:
#    - name: Mikko Vihtakari
#      affiliations:
#       - ref: imr
# affiliations: 
#    - id: imr 
#      name: Institute of Marine Research, Norway
date: today
published-title: "Version"
format:
  html:
    self-contained: true
    fig-width: 9
    fig-height: 6
    code-fold: true
    output-file: SI-Text-3-model-selection
    toc: true
bibliography: ../manuscript/zotero_library.bib
csl: ../manuscript/ices_jms.csl
execute:
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

```{r setup, eval = TRUE, include = FALSE}
# Clear the workspace for manual code runs

# rm(list = ls()) 

knitr::opts_knit$set(root.dir = '../../') 

knitr::knit_hooks$set(timeit = local({
  now = NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res = round(difftime(Sys.time(), now, units = "mins"), 1)
      now <<- NULL
      # use options$label if you want the chunk label as well
      print(paste('Time:', as.character(res), "min"))
    }
  }})
)

# Settings
recompile_models <- FALSE # switch to compile the model. If FALSE, the model is loaded from the data/model_output folder.
recalculate_index <- recompile_models # switch to estimate index values from the predictive skill fold (10%)
recompile_data <- FALSE # switch to recompile data from sources. FALSE saves a couple of minutes when include_temperature = TRUE

component_ranking <- FALSE # run component ranking?
model_selection <- FALSE
family_selection <- FALSE
spatiotemporal_selection <- FALSE

development_mode <- FALSE # Only uses a subset of data and coarse meshes to able quick compilation of the models. Turn this option off once the document is finished and rerun. 
```

This supplementary information file for @vihtakari2026 provides a reproducible overview of the model selection process using generalized linear mixed-effects models fitted with the [sdmTMB package](https://CRAN.R-project.org/package=sdmTMB) [@sdmTMB] in R [@rcoreteam]. The four-step model selection procedure employs 10-fold cross-validation via the sdmTMB_cv() function. Models are compared based on total predicted (out-of-sample) negative log-likelihood (NLL) and Akaike information criterion (AIC) values. To further assess the impact of model alternatives on survey indices, visual comparisons are made with the best-fitting model. This document was written with [Quarto](https://quarto.org/) and can be compiled from the project's [GitHub repository](https://github.com/DeepWaterIMR/ref-index-paper). Foldable code fields are included to display the code used for model compilation.

The model selection routine comprises four key steps:

1. The fixed- and random-effect model components are initially [ranked according to model fit](@comp-ranking), as measured by NLL. The fixed effects include gear type (factor), year (factor), month (continuous), bottom depth (continuous), and modelled bottom temperature (continuous). Spatial random fields and IID (independent and identically distributed) spatiotemporal fields are incorporated as random effects. Each fixed effect is evaluated within models that include both spatial and spatiotemporal random fields, while models without fixed effects serve as null models for comparison.
2. From the previous step all fixed effects are combined into a single model. The influence of each fixed effect on model performance is then assessed by [sequentially removing](@model-selection) one effect at a time, with the goal of identifying the best overall model for both species. All models are fitted with spatial random fields and IID spatiotemporal fields as random effects. A model with only spatial- and spatiotemporal fields is used as a null model for comparison. 
3. After determining the best-fitting model formula, alternative model families (delta-lognormal, delta-gamma, delta gamma poisson link, and Tweedie distributions) are compared using this formula.
4. Finally, the influence of different spatiotemporal field structures is evaluated using the best-fitting model and distribution family from previous steps. These structures include IID, first-order autoregressive AR(1), random walk (RW), and the absence of spatiotemporal fields.

Overall, the best fitting model had a formula of rho_biomass ~ 0 + geartype + s(month) + s(depth) + s(temperature), AR1 spatiotemporal random fields and delta lognormal model family as shown below. 

```{r packages}
#| timeit: !expr if(recompile_data) TRUE else NULL

############## ####
#### Libraries ----

## Package names
packages <- 
  c("tidyverse", "sf", "ggOceanMaps", "ggspatial", "sdmTMB", 
    "ggeffects", "scales", "concaveman", "plotly", "data.table",
    "dtplyr", "remotes", "sdmTMBextra", "lubridate", "inlabru", 
    "future", "pbapply")

## Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())

if("remotes" %in% packages[!installed_packages]) {
  install.packages("remotes")
}

if (any(installed_packages == FALSE)) {
  
  if("sdmTMBextra" %in% packages[!installed_packages]) {
    remotes::install_github("pbs-assess/sdmTMBextra", dependencies = TRUE,
                            upgrade = "never")
  }
  
  installed_packages <- packages %in% rownames(installed.packages())
  
  install.packages(packages[!installed_packages])
}

## Load packages
invisible(lapply(packages, function(x) {
  suppressPackageStartupMessages(library(x, character.only = TRUE))}))

# Backend for parallel

if(Sys.info()["sysname"] == "Linux") {
  library(flexiblas)
  flexiblas_switch(flexiblas_load_backend("OPENBLAS-THREADS"))
}

################### ###
# Custom functions ----

source("R/legacy/make_table.R") # a function to make tables
source("R/legacy/bubble_map.R")

################ ##
# ggplot theme ----

theme_cust <- theme_classic(base_size = 12) %+replace%
  theme(strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank(),
        legend.background = element_blank(),
        legend.box.background = element_blank(),
        plot.margin = margin(5.5, 10, 5.5, 5.5))

theme_set(theme_cust) # Set default theme globally for the entire project

# Parallel processing settings

n_cores <- 10 # Number of cores to be used in parallel processing within pblapply. Use NULL for sequential processing. "future" uses the n_sessions argument
n_sessions <- 10 # Number of parallel R sessions to be used in parallel processing within future::plan
```

<details>
<summary>Unfold to see R package versions and system information used to generate the document</summary>
```{r}
#| label: sys-info
sessionInfo()
```
</details>

# Data

Compile the data: 

```{r data}
#| timeit: !expr if(recompile_data) TRUE else NULL

# Detailed settings

run_fixef <- TRUE # make fixed effect plot s
run_index <- TRUE # switch to ex/include the index estimation

include_temperature <- TRUE # switch to include temperature as a predictor variable
segmental_save <- TRUE # save model files segmentally. Adds flexibility in case of crashes, but increases processing time considerably.
save_model_output <- FALSE # save model output. Takes tens of minutes. Alternative is to run without saving, which goes quicker, but does not allow reopening the models. 

## Response variable
response_variable <- "biomass density" # options: "catch biomass" meaning trawl catch weight of the target species (biomass column in dt). Log transformed swept area by each trawl will be used as offset argument. "biomass density" referring to the rho_biomass column giving readily calculated biomass density (biomass/swept area). "catch abundance" and "abundance density" for similar variables using abundance/count data. Note that the width of the trawl influences the relative difference between gear types, which is included to the model, and only affect the absolute level of the index estimate in the end. 

## Model family (use this to set the family)
model_family <- delta_lognormal()
model_family_name <- ifelse(is.null(model_family$clean_name), 
                            model_family$family, model_family$clean_name)

### Link function inversion function
if(length(model_family$family) > 1) {
  linkinv <- model_family[[2]]$linkinv
} else {
  linkinv <- model_family$linkinv
}

## Mesh cutoff argument (smaller values create meshes with more knots)
mesh_cutoff <- ifelse(development_mode, 150, 75) # Minimum distance between knots before a new mesh vertex is added in kilometers

####################
# Read the data ----

if(recompile_data) {
  
  load("data/out/Compiled_redfish_survey_data.rda")
  
  dt <- lapply(c("snabeluer", "vanlig uer"), function(filter_sp) {
    out <- stn %>%
      distinct() %>% 
      filter(commonname %in% filter_sp) %>% 
      rename(lon = longitudestart,
             lat = latitudestart,
             year = startyear,
             depth = bottomdepthstart) %>% 
      mutate(gear_width = 25) %>% 
      mutate(cruiseseries = factor(cruiseseries),
             fyear = factor(year),
             id = paste(cruiseseries, year, platformname, serialnumber, sep = "-"),
             swept_area = (gear_width/1852)*distance) %>% 
      mutate(yday = lubridate::yday(stationstartdate),
             .before = stationstopdate) %>% 
      droplevels() %>% 
      mutate(rho_biomass = biomass/swept_area,
             rho_abundance = count/swept_area,
             value = rho_biomass) %>% # this defines the response type
      filter(!is.na(value))
    
    ## Zero catch survey removal
    tmp <- out %>% 
      group_by(cruiseseries, cruise, year, platformname) %>% 
      reframe(sum = sum(value)) %>% 
      filter(sum == 0) %>% 
      mutate(tmpid = paste(cruise, year, platformname))
    
    out %>% 
      mutate(tmpid = paste(cruise, year, platformname)) %>% 
      filter(!tmpid %in% tmp$tmpid) %>% 
      dplyr::select(-tmpid)
  }) %>% 
    bind_rows()
  
  ## Reduce data size when running in development mode
  
  if(development_mode) {
    dt <- lapply(unique(dt$commonname), function(k) {
      dt %>% 
        filter(year > 2003) %>% 
        filter(commonname == k) %>% 
        mutate(sample = rep(1:4, length.out = nrow(.))) %>% 
        filter(sample == 1) %>% 
        dplyr::select(-sample)
    }) %>% bind_rows()
  }
  
  ## Response variable
  
  if(response_variable == "biomass density") {
    dt <- dt %>% mutate(value = rho_biomass, .after = rho_abundance) %>% 
      filter(!is.na(value))
  } else if(response_variable == "abundance density") {
    dt <- dt %>% mutate(value = rho_abundance, .after = rho_abundance) %>% 
      filter(!is.na(value))
  } else if(response_variable == "catch biomass") {
    dt <- dt %>% mutate(value = biomass, .after = rho_abundance) %>% 
      filter(!is.na(value))
  } else if(response_variable == "catch abundance") {
    dt <- dt %>% mutate(value = count, .after = rho_abundance) %>% 
      filter(!is.na(value))
  } else {
    stop("The response_variable object has to be 'biomass density', 'abundance density', 'catch biomass', or 'catch abundance'")
  }
  
  ## CRS
  
  dt_crs <- get_crs(dt, ll_names = c("lon", "lat"))
  
  dt <- sdmTMB::add_utm_columns(dt, ll_names = c("lon", "lat"), utm_crs = dt_crs)
  
  # Response variable type definitions for automatic text
  
  model_type <- ifelse(response_variable %in% c("biomass density", "catch biomass"), "Biomass", "Abundance")
  
  # Add temperature predictor variable (takes time)
  if(include_temperature) {
    load("data/model_output/temperature-model-fit.rda")
    
    ## Predict using the data
    
    dt <- predict(
      fit[[1]]$mod,
      newdata = dt, 
      type = "response",
      model = NA,
      return_tmb_object = FALSE) %>% 
      dplyr::select(-est_non_rf, -est_rf, -omega_s, -epsilon_st) %>% 
      rename("temperature" = "est") %>% 
      relocate("temperature", .after = "depth")
  }
  
  ## Response variables and units for figure labels
  if(response_variable == "biomass density") {
    value_name <- expression(paste("Density (kg ", nmi^-2, ")"))
  } else if(response_variable == "catch biomass") {
    value_name <- "Catch weight (kg)"
  } else if(response_variable == "abundance density") {
    value_name <- expression(paste("Density (n ", nmi^-2, ")"))
  } else {
    value_name <- "Catch count (n)"
  }
  
  if(response_variable == "biomass density") {
    value_unit <- quote(kg~nmi^-2)
  } else if(response_variable == "catch biomass") {
    value_unit <- "kg"
  } else if(response_variable == "abundance density") {
    value_unit <- quote(n~nmi^-2)
  } else {
    value_unit <- "n"
  }
  
  ## Response variables and units for text
  if(response_variable == "biomass density") {
    value_text_name <- "Swept area biomass density (kg nmi$^{-2}$)"
  } else if(response_variable == "catch biomass") {
    value_text_name <- "Catch weight (kg)"
  } else if(response_variable == "abundance density") {
    value_text_name <- "Swept area abundance density (n nmi$^{-2}$)"
  } else {
    value_text_name <- "Catch count (n)"
  }
  
  if(response_variable == "biomass density") {
    value_text_unit <- "kg nmi$^{-2}$"
  } else if(response_variable == "catch biomass") {
    value_text_unit <- "kg"
  } else if(response_variable == "abundance density") {
    value_text_unit <- "n nmi$^{-2}$"
  } else {
    value_text_unit <- "n"
  }
  
  ## Species
  
  filter_sp <- "both"
  sp_name <- case_when(filter_sp == "snabeluer" ~ "beaked redfish", 
                       filter_sp == "vanlig uer" ~ "golden redfish", 
                       filter_sp == "both" ~ "beaked and golden redfish")
  Sp_name <- case_when(filter_sp == "snabeluer" ~ "Beaked redfish", 
                       filter_sp == "vanlig uer" ~ "Golden redfish", 
                       filter_sp == "both" ~ "Beaked and golden redfish")
  
  dt <- dt %>% 
    mutate(
      species = case_when(
        commonname == "snabeluer" ~ "Beaked redfish", 
        commonname == "vanlig uer" ~ "Golden redfish"
      ),
      .after = "commonname")
  
  save(dt, dt_crs, model_type, value_name, value_text_name, value_text_unit,
       sp_name, Sp_name,
       file = "data/model_output/model_selection/model-selection-input-data.rda",
       compress = "xz")
  
} else {
  load("data/model_output/model_selection/model-selection-input-data.rda")
}
```

Swept area `r tolower(model_type)` density (`r ifelse(model_type == "Biomass", "kg", "n")` nmi$^{-2}$), called 'value' henceforth, was used as the response variable.

# Component ranking {#comp-ranking}

```{r component-ranking}
#| code-fold: true
#| warning: false
#| message: false
#| results: hide
#| timeit: !expr if(recompile_models & component_ranking) TRUE else NULL

if(recompile_models & component_ranking) {
  
  component_ranking_list <- tibble(
    name = c(
      "Bottom depth",
      "Year",
      "Month",
      "Gear type",
      "Temperature",
      "Spatial and spatiotemporal",
      "Only spatiotemporal"
    ),
    formula = list(
      formula(value ~ s(depth)),
      formula(value ~ fyear),
      formula(value ~ s(month)),
      formula(value ~ geartype),
      formula(value ~ s(temperature)),
      formula(value ~ 1),
      formula(value ~ 1)
    ),
    spatial = c(
      "on",
      "on",
      "on",
      "on",
      "on",
      "on",
      "off"
    ),
    spatiotemporal = c(
      "iid",
      "iid",
      "iid",
      "iid",
      "iid",
      "iid",
      "iid"
    )
  )
  
  if(!include_temperature) {
    component_ranking_list <- 
      component_ranking_list %>% filter(name != "Temperature")
  }
  
  component_ranking_list <- component_ranking_list %>% 
    replicate_df("commonname", levels(dt$commonname)) %>% 
    relocate("commonname", .before = 1)
  
  if(!dir.exists("data/model_output/model_selection")) {
    dir.create("data/model_output/model_selection")
  }
  
  saveRDS(component_ranking_list, 
          file = "data/model_output/model_selection/component_ranking_list.rds")
  
  future::plan("multisession", workers = n_sessions)
  options(future.globals.maxSize = 1000 * 1024^2, future.seed = TRUE)
  
  ## Fit the models 
  
  message("Cross-validation (takes an hour on a powerful computer)")
  
  tmp <- pbapply::pblapply(1:nrow(component_ranking_list), function(i) {
    
    x <- dt %>% 
      filter(commonname == component_ranking_list$commonname[i]) %>%
      arrange(stationstartdate, year, missionid, serialnumber) %>% 
      droplevels()
    
    comp_ranking_mesh <- make_mesh(x, xy_cols = c("X", "Y"), cutoff = mesh_cutoff)
    clust <- sample(seq_len(10), size = nrow(x), replace = TRUE)
    
    print(i)
    term <- all.vars(component_ranking_list$formula[[i]])[-1]
    continuous_predictor <- 
      any(grepl(paste0("s\\(", term, "\\)"),
                as.character(component_ranking_list$formula[[i]])))
    
    y <- x %>% {
      if(continuous_predictor) {
        mutate(., {{term}} := as.numeric(as.character(get(term))))
      } else {
        if(length(term) > 0) mutate(., {{term}} := as.factor(get(term))) else .
      }}
    
    out <- try({
      sdmTMB_cv(
        component_ranking_list$formula[[i]],
        data = y,
        mesh = comp_ranking_mesh,
        fold_ids = clust,
        family = model_family,
        spatial = component_ranking_list$spatial[[i]],
        spatiotemporal = component_ranking_list$spatiotemporal[[i]],
        time = "year",
        control = sdmTMBcontrol(newton_loops = 2)
      )
    })
    
    if(inherits(out, "try-error")) {
      warning(paste(paste(component_ranking_list$commonname,
                          component_ranking_list$name, sep = "_")[i], "crashed"))
    }
    
    # Save / return
    if(segmental_save) {
      saveRDS(
        list(mod = out, mesh = comp_ranking_mesh),
        file =
          paste0("data/model_output/model_selection/Component-ranking-models-",
                 gsub("\\(.*\\)", "", model_family_name), "_",
                 paste(component_ranking_list$commonname, 
                       gsub(" ", "-", component_ranking_list$name), sep = "_")[i],
                 ".rds"),
        compress = "xz"
      )
    } else {
      return(list(mod = out, mesh = comp_ranking_mesh))
    }
  })
  
  # Load and make objects
  
  if(segmental_save) {
    component_ranking_models <- 
      pbapply::pblapply(1:nrow(component_ranking_list), function(i) {
        readRDS(
          file =
            paste0("data/model_output/model_selection/Component-ranking-models-",
                   gsub("\\(.*\\)", "", model_family_name), "_",
                   paste(component_ranking_list$commonname, 
                         gsub(" ", "-", component_ranking_list$name), sep = "_")[i],
                   ".rds")
        )$mod
      })
    
    component_ranking_models <- 
      setNames(component_ranking_models, 
               paste(component_ranking_list$name, 
                     component_ranking_list$commonname, sep = "_"))
    
    component_ranking_mesh <- 
      pbapply::pblapply(
        unique(dt$commonname), function(k) {
          readRDS(
            dir("data/model_output/model_selection/", 
                pattern = paste0("Component-ranking-models-",
                                 gsub("\\(.*\\)", "", model_family_name), "_", k),
                full.names = TRUE)[1])$mesh
        })
  } else {
    component_ranking_models <- lapply(tmp, function(k) k$mod)
    component_ranking_models <- 
      setNames(component_ranking_models, 
               paste(component_ranking_list$name, 
                     component_ranking_list$commonname, sep = "_"))
    
    component_ranking_mesh <- 
      lapply(c(1,8), function(i) tmp[[i]]$mesh)
    
    component_ranking_mesh <- 
      setNames(component_ranking_mesh, levels(dt$commonname))
    
    if(save_model_output) {
      saveRDS(component_ranking_models, 
              file = paste0(
                "data/model_output/model_selection/Component-ranking-models-",
                gsub("\\(.*\\)", "", model_family_name), ".rds"),
              compress = "xz"
      ) 
    }
    
    saveRDS(component_ranking_mesh,
            file = paste0(
              "data/model_output/model_selection/Component-ranking-meshes-",
              gsub("\\(.*\\)", "", model_family_name), ".rds"),
            compress = "xz"
    )  
  }
  
  ## Save required components
  component_ranking_params <- 
    list(
      model_names = names(component_ranking_models)
    )
  
  saveRDS(
    component_ranking_params,
    file = 
      paste0(
        "data/model_output/model_selection/Component-ranking-parameters-",
        gsub("\\(.*\\)", "", model_family_name), ".rds")
  )
  
} else {
  component_ranking_list <- 
    readRDS(
      file = "data/model_output/model_selection/component_ranking_list.rds")
  
  component_ranking_mesh <- 
    readRDS(paste0(
      "data/model_output/model_selection/Component-ranking-meshes-",
      gsub("\\(.*\\)", "", model_family_name), ".rds")
    )
  
  component_ranking_params <-
    readRDS(paste0(
      "data/model_output/model_selection/Component-ranking-parameters-",
      gsub("\\(.*\\)", "", model_family_name), ".rds")
    )
}
```  

```{r}
#| label: tbl-component-ranking
#| timeit: !expr if(recompile_models & component_ranking) TRUE else NULL
#| tbl-cap: !expr paste0("Ranking of single-component models for ", sp_name, " based on ten-fold cross-validation. The models contain spatial and spatiotemporal random fields, unless specified otherwise. Each model is compared to the best fitting single component model ($\\Delta$NLL and $\\Delta$NLL % columns). The negative log-likelihood value is given in the NLL column and AIC difference to the best fitting model in the $\\Delta$AIC column.")

## Table
if(recompile_models & component_ranking) {
  component_ranking_tab <- 
    pbapply::pblapply(seq_along(component_ranking_models), function(i) {
      if(inherits(component_ranking_models[[i]], "try-error")) {
        tibble(
          species = case_when(
            component_ranking_list$commonname[i] == "snabeluer" ~ "Beaked redfish", 
            component_ranking_list$commonname[i] == "vanlig uer" ~ "Golden redfish"
          ),
          model = component_ranking_list$name[i],
          formula = paste(trimws(format(component_ranking_list$formula[[i]])), 
                          collapse = ""),
          spatial = component_ranking_list$spatial[[i]],
          spatiotemporal = component_ranking_list$spatiotemporal[[i]],
          crashed = TRUE,
          converged = FALSE,
          sanity = FALSE,
          loglik = NA,
          AIC = NA
        )
      } else {
        tibble(
          species = case_when(
            component_ranking_list$commonname[i] == "snabeluer" ~ "Beaked redfish", 
            component_ranking_list$commonname[i] == "vanlig uer" ~ "Golden redfish"
          ),
          model = component_ranking_list$name[i],
          formula = paste(trimws(format(component_ranking_models[[i]]$models[[1]]$formula[[1]])), 
                          collapse = ""),
          spatial = unique(component_ranking_models[[i]]$models[[1]]$spatial),
          spatiotemporal = unique(component_ranking_models[[i]]$models[[1]]$spatiotemporal),
          crashed = FALSE,
          converged = component_ranking_models[[i]]$converged,
          sanity = all(unlist(sanity(component_ranking_models[[i]]$models[[1]],
                                     silent = TRUE))),
          loglik = component_ranking_models[[i]]$sum_loglik,
          AIC = median(sapply(component_ranking_models[[i]]$models, AIC))
        )
      }
    }, cl = n_cores) %>% 
    bind_rows() %>% 
    group_by(species) %>% 
    arrange(-loglik) %>% 
    mutate(
      deltaNLL = loglik - max(loglik, na.rm = TRUE),
      deltaNLL_pr = 100*(1-loglik/max(loglik, na.rm = TRUE)),
      deltaAIC = AIC - min(AIC, na.rm = TRUE),
      deltaAIC_pr = 100*(AIC/min(AIC, na.rm = TRUE) - 1)
    ) %>% 
    ungroup()
  
  ## Save
  write_csv(
    component_ranking_tab,
    file =
      paste0("data/model_output/model_selection/Component-ranking-table-", 
             gsub("\\(.*\\)", "", model_family_name),
             ".csv")
  )
} else {
  component_ranking_tab <- read_csv(
    paste0("data/model_output/model_selection/Component-ranking-table-", 
           gsub("\\(.*\\)", "", model_family_name),
           ".csv")
  )
}

## Print the table
component_ranking_tab %>% 
  filter(model != "Month categorical") %>%
  mutate(model = gsub(" continuous", "", model)) %>% 
  relocate(formula, .after = last_col()) %>%
  dplyr::select(-spatial, -spatiotemporal, -AIC, -deltaAIC_pr) %>% 
  mutate(loglik = abs(loglik), 
         deltaNLL = abs(deltaNLL),
         deltaNLL_pr = abs(deltaNLL_pr)) %>% 
  rename(
    "Species" = "species", "Model" = "model", "Formula" = "formula", 
    "Crashed" = "crashed", "Converged" = "converged", "Sanity test" = "sanity", 
    "NLL" = "loglik", "ΔNLL" = "deltaNLL", "ΔNLL %" =  "deltaNLL_pr", 
    "ΔAIC" = "deltaAIC"
  ) %>% 
  make_table(digits = c(0,0,0,0,0,0,0,1,0,0))

# More compact alternative:
# component_ranking_tab %>% 
#   mutate(deltaNLL = abs(deltaNLL)) %>% 
#   dplyr::select(species, model, formula, sanity, deltaNLL) %>%
#   ungroup() %>% 
#   pivot_wider(id_cols = c(model, formula, sanity), 
#               names_from = species,
#               values_from = deltaNLL) %>% 
#   filter(!model %in% "Full model s(logdepth)") %>% 
#   rename("Model" = "model", "Formula" = "formula", 
#          "Sanity test" = "sanity") %>%
#   make_table(digits = 2)
```

```{r}
#| label: fig-component-ranking
#| fig-cap: !expr paste("Ranking of single-component models for", sp_name, "based on ten-fold cross-validation. The y-axis lists the model components. The x-axis shows the negative log-likelihood for each model, where higher values indicate better model fit. Symbol color and shape reflect model diagnostics; black circles denote models without issues, while grey symbols indicate models that failed the sanity check or did not converge.")

tmp <- component_ranking_tab %>%
  filter(model != "Month continuous") %>%
  mutate(model = gsub(" categorical", "", model)) %>% 
  group_by(species) %>% 
  mutate(status = case_when(
    !sanity & !converged ~ "Nonconvergent &\nsanity test issues",
    !sanity ~ "Sanity test issues",
    !converged ~ "Nonconvergent",
    .default = "All OK"
  ), .after = "sanity") %>% 
  mutate(deltaNLL = abs(loglik - max(loglik, na.rm = TRUE))) %>% 
  mutate(deltaNLL = abs(ifelse(crashed, Inf, deltaNLL))) %>% 
  rename("value" = "deltaNLL") %>% 
  ungroup()

tmp <- tmp %>% 
  mutate(
    model = factor(
      model,
      levels = tmp %>% group_by(model) %>% reframe(value = mean(value)) %>% 
        arrange(value) %>% pull(model) %>% rev())
  )

ggplot(tmp, aes(x = model, y = value, color = status, shape = status)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c(16,15,17,18)) +
  scale_color_manual(values = c("black", rep("grey", 3))) +
  facet_grid(~species, scales = "free") +
  coord_flip() + 
  # scale_y_continuous(transform = "sqrt") + 
  labs(y = "Negative log likelihood difference",
       x = "Variable", shape = "Model diagnostics", color = "Model diagnostics") +
  theme(legend.position = "bottom",
        panel.grid.major = element_line(color = "grey", linewidth = LS(0.5)))
```

## Fixed effects

```{r}
#| label: fig-fixed-effects
#| fig-height: 12
#| timeit: !expr if(recompile_models & component_ranking) TRUE else NULL
#| fig-cap: Conditional fixed effects for single component models. 2024 was used as the temporal reference point. Error bars and shading indicate 95% confidence intervals.

if(run_fixef) {
  
  if(recompile_models & component_ranking) {
    terms <- grep("spatial|spatiotemporal", names(component_ranking_models), 
                  ignore.case = TRUE, invert = TRUE, value = TRUE)
    
    fixef_dt <- 
      pbapply::pblapply(terms, function(k) {
        
        if(inherits(component_ranking_models[[k]], "try-error")) return(NULL)
        
        model <- component_ranking_models[[k]]$models[[1]]
        
        term <- 
          all.vars(component_ranking_models[[k]]$models[[1]]$formula[[1]])[-1] 
        model_data <- component_ranking_models[[k]]$data
        
        if(inherits(model_data[[term]], "numeric")) {
          tt <- 
            seq(min(model_data[[term]]), max(model_data[[term]]), 
                length.out = 50) %>% 
            tibble() %>% 
            setNames(term) 
        } else {
          tt <- 
            unique(model_data[[term]]) %>% 
            sort() %>% 
            tibble() %>% 
            setNames(term)
        }
        
        if(!grepl("Year", k)) {
          tt <- tt %>% 
            mutate(year = max(model_data[["year"]]))
        } else {
          tt <- tt %>% 
            mutate(year = as.integer(as.character(get(term))))
        }
        
        tmp <- predict(
          model, tt,
          se_fit = TRUE, re_form = NA, model = NA
        ) %>%
          mutate(lwr = linkinv(est - est_se), 
                 upr = linkinv(est + est_se), 
                 est = linkinv(est)) 
        
        ymax <- ifelse(max(tmp$upr) > 1.5*max(tmp$est), 1.5*max(tmp$est), 
                       max(tmp$upr))
        
        tmp <- tmp %>% 
          mutate(commonname = unique(model_data$commonname), .before = 1)
        
        list(dt = tmp, ymax = ymax, term = term)
      }, cl = n_cores)
    
    
    fixef_dt <- setNames(fixef_dt, terms)
    
    ## Save
    
    saveRDS(
      fixef_dt,
      file =
        paste0(
          "data/model_output/model_selection/Component-ranking-fixed-effects-", 
          gsub("\\(.*\\)", "", model_family_name),
          ".rds"),
      compress = "xz"
    )
    
  } else {
    terms <- grep("spatial|spatiotemporal", component_ranking_params$model_names,
                  ignore.case = TRUE, invert = TRUE, value = TRUE)
    
    fixef_dt <- readRDS(
      file =
        paste0(
          "data/model_output/model_selection/Component-ranking-fixed-effects-", 
          gsub("\\(.*\\)", "", model_family_name),
          ".rds")
    )
  }
  
  cowplot::plot_grid(
    plotlist = 
      lapply(unique(dt$commonname), function(k) {
        # print(k)
        tmp <- fixef_dt[grep(k, terms, ignore.case = TRUE, value = TRUE)]
        
        cowplot::plot_grid(
          plotlist = 
            lapply(names(tmp), function(kk) {
              # print(kk)
              g <- tmp[[kk]]
              
              if(is.null(g) | inherits(g, "try-error")) {
                return({
                  ggplot() + 
                    geom_blank() + 
                    annotate(
                      "text", x = 0, y = 0, 
                      label = paste(kk, "crashed", sep = "\n"),
                      size = FS(10)) +
                    theme_void()
                })
              }
              
              if(inherits(g$dt[[g$term]], "numeric")) {
                g$dt %>%
                  ggplot(aes(x = get(g$term), y = est, ymin = lwr, ymax = upr)) +
                  geom_ribbon(alpha = 0.3, color = NA) +
                  geom_path() +
                  coord_cartesian(expand = FALSE, ylim = c(0, g$ymax)) +
                  labs(y = value_name, x = g$term) +
                  theme_bw(base_size = 10) +
                  theme(legend.position = c(0.1,0.7),
                        strip.background = element_blank(),
                        panel.background = element_blank(),
                        plot.background = element_blank(),
                        legend.background = element_blank(),
                        legend.box.background = element_blank(),
                        plot.margin = margin(5.5, 10, 5.5, 5.5))
              } else {
                g$dt %>%
                  ggplot(aes(x = get(g$term), y = est, ymin = lwr, ymax = upr)) +
                  geom_errorbar(width = 0.2, position = position_dodge(0.5)) +
                  geom_point(position = position_dodge(0.5)) +
                  coord_cartesian(expand = FALSE, ylim = c(0, g$ymax)) + {
                    if(g$term == "fyear") {
                      scale_x_discrete(
                        breaks = unique(g$dt$fyear),
                        labels = substr(unique(g$dt$fyear), 3, 4),
                        guide = guide_axis(n.dodge = 2)
                      ) 
                    }
                  } +
                  labs(y = value_name, x = g$term) +
                  theme_bw(base_size = 10) +
                  theme(legend.position = "none",
                        strip.background = element_blank(),
                        panel.background = element_blank(),
                        plot.background = element_blank(),
                        legend.background = element_blank(),
                        legend.box.background = element_blank(),
                        plot.margin = margin(5.5, 10, 5.5, 5.5))
              }
            })
        )
        
      }),
    labels = unique(dt$species), ncol = 1
  )
}
```

## Survey indices

```{r comp-rank-survey-indices}
#| timeit: !expr if(recalculate_index & component_ranking) TRUE else NULL

if(run_index) {
  if(recalculate_index & component_ranking) {
    
    proj_dt <- read_rds("data/model_output/prediction-grid-data.rds") %>% 
      distinct(X, Y, depth, year) 
    
    ## Prediction grid cell side length
    grid_side_size <- attributes(proj_dt)$grid_side_size # kilometers due to UTM projection and division by 1000
    
    if(development_mode) {
      proj_dt <- proj_dt %>% 
        filter(year > 2003) %>% 
        mutate(sample = rep(1:4, length.out = nrow(.))) %>% 
        filter(sample == 1) %>% 
        dplyr::select(-sample)
    }
    
    models <- component_ranking_models
    
    index <- pbapply::pblapply(seq_along(models), function(i) {
      
      # print(i)
      
      k <- models[[i]]
      if(inherits(k, "try-error")) return(NULL)
      k_name <- names(models)[i]
      
      terms <- trimws(unlist(strsplit(
        as.character(k$models[[1]]$formula[[1]])[3], "\\+")))
      terms <- terms[terms != "0"]
      terms <- terms[!grepl(":", terms)]
      
      pred_dt <- proj_dt
      
      if(grepl("Spatial and spatiotemporal", k_name)) {
        pred_dt <- pred_dt %>% distinct(X, Y, year)
      } else {
        fyear <- any(grepl("fyear", terms))
        geartype <- any(grepl("geartype", terms))
        logdepth <- any(grepl("logdepth", terms))
        cruiseseries <- any(grepl("cruiseseries", terms))
        season <- any(grepl("season", terms))
        depth <- any(grepl("depth", terms))
        month <- any(grepl("month", terms))
        smonth <- any(grepl("s\\(month\\)", terms))
        yday <- any(grepl("yday", terms))
        distance <- any(grepl("distance", terms))
        commonname <- any(grepl("commonname", terms))
        temperature <- any(grepl("temperature", terms))
        
        if(!any(c(depth, temperature))) pred_dt <- pred_dt %>% distinct(X, Y, year)
        if(logdepth) pred_dt <- pred_dt %>% mutate(logdepth = log(depth))
        if(geartype) pred_dt <- pred_dt %>% mutate(geartype = "Shrimp")
        if(fyear) pred_dt <- pred_dt %>% mutate(fyear = as.factor(year))
        if(season) pred_dt <- pred_dt %>% mutate(season = "H2")
        if(month) pred_dt <- pred_dt %>% replicate_df("month", if(smonth) as.integer(8) else factor(8))
        if(yday) pred_dt <- pred_dt %>% mutate(yday = 225)
        if(distance) pred_dt <- pred_dt %>% mutate(distance = 1) # assuming 3 knot trawling speed for 20 min
        
        if(include_temperature & temperature) {
          
          ### Temperature model ###
          load("data/model_output/temperature-model-fit.rda")
          
          ## Predict using the data
          
          pred_dt <- predict(
            fit[[1]]$mod,
            newdata = pred_dt %>% 
              mutate(fyear = as.factor(year),
                     month = as.integer(8)),
            type = "response",
            model = NA,
            return_tmb_object = FALSE) %>% 
            dplyr::select(-est_non_rf, -est_rf, -omega_s, -epsilon_st) %>% 
            rename("temperature" = "est") 
         
          if(!smonth) {
            pred_dt <- pred_dt %>% mutate(month = factor(month))
          }
           
        }
        
      }
      
      ## Slower more official way with CIs
      # lapply(seq_along(k$models), function(j) {
      #   predict(k$models[[j]], newdata = proj_dt, type = "response", model = NA,
      #           return_tmb_object = TRUE) %>% 
      #     get_index(
      #       bias_correct = FALSE,
      #       area = (grid_side_size/1.852)^2) %>% # grid cell area in square nautical miles
      #     mutate(est = est/1e6, # convert kg to kilotonnes
      #            lwr = lwr/1e6,
      #            upr = upr/1e6, 
      #            cv = sqrt(exp(se^2) - 1)) %>% 
      #     mutate(model = k_name, .before = 1) %>% 
      #     mutate(fold = j)
      # }) %>% 
      #   bind_rows()
      
      ## Quicker way
      mod_pred <- lapply(seq_along(k$models), function(j) {
        predict(k$models[[j]], newdata = pred_dt, type = "response", model = NA,
                return_tmb_object = FALSE) %>%
          mutate(fold = j)
      })
      
      mod_pred %>%
        bind_rows() %>%
        group_by(fold, year, X, Y) %>%
        reframe(est = mean(est)) %>%
        group_by(fold, year) %>%
        reframe(
          # Y = weighted.mean(Y, est),
          # X = weighted.mean(X, est),
          # depth = weighted.mean(depth, est),
          est = (sum(est)*(grid_side_size/1.852)^2)/1e6
        ) %>%
        mutate(lwr = NA, upr = NA, cv = NA, se = NA) %>%
        mutate(commonname = unique(k$data$commonname),
               model = k_name, 
               .before = 1)
    }, cl = length(models))
    
    names(index) <- names(models)
    
    write_rds(
      index,
      file =
        paste0("data/model_output/model_selection/Component-ranking-indices-", 
               gsub("\\(.*\\)", "", model_family_name),
               ".rds"), 
      compress = "xz"
    )
    
    sc_indices <- index
    
  } else {
    
    proj_dt <- read_rds("data/model_output/prediction-grid-data.rds") %>% 
      distinct(X, Y, depth, year)
    
    sc_indices <- read_rds(
      paste0("data/model_output/model_selection/Component-ranking-indices-", 
             gsub("\\(.*\\)", "", model_family_name),
             ".rds"), 
    ) 
  }
}
```

```{r}
#| code-fold: true
#| label: fig-comp-rank-index
#| fig-height: !expr ceiling(length(sc_indices)/3)*2.5
#| fig-cap: !expr paste0("Survey indices estimated using single-component models for ", sp_name, " using spatiotemporal generalized linear mixed effects models with the sdmTMB package (colored solid line). Black lines with shading indicate the full biomass models with multiple components; 0 + geartype + s(month) + s(depth) + s(temperature). ", value_text_name, " is used as the response variable, and the models use meshes with ", paste(sapply(component_ranking_mesh, function(k) k$mesh$n), collapse = " and "), " vertices (knots), as well as ", model_family_name," model family. Predictions are made to a grid with ", proj_dt %>% distinct(X,Y) %>% nrow(), " cells with a side of ", attributes(proj_dt)$grid_side_size, " km. Each line represents one fold from the 10-fold cross-validation. See Table 1 for more detailed model specification.")

if(run_index) {
  tmp <- lapply(sc_indices, function(k) {
    if(inherits(k, "try-error")) return(NULL)
    if(any(!is.na(k$upr))) {
      if(max(k$upr, na.rm = TRUE) > 2*max(k$est, na.rm = TRUE)) {
        k[k$upr > 1.5*max(k$est, na.rm = TRUE),]$upr <- ceiling(1.5*max(k$est, na.rm = TRUE))
      }}
    k
  }) %>%
    bind_rows() %>% 
    filter(!is.na(year)) %>% 
    mutate(model = gsub("_snabeluer|_vanlig uer", "", model)) %>% 
    # mutate(model = factor(model, levels = names(sc_indices))) %>% 
    mutate(
      species = case_when(
        commonname == "snabeluer" ~ "Beaked redfish", 
        commonname == "vanlig uer" ~ "Golden redfish"
      ),
      .after = "commonname")
  
  load("data/model_output/index-bm.rda")
  
  index <- index %>% 
    bind_rows() %>% 
    rename("species" = "length_group")
  
  cowplot::plot_grid(
    plotlist = 
      lapply(unique(tmp$species), function(k) {
        ggplot() +
          geom_ribbon(
            data = index %>% filter(species == k),
            aes(year, est, ymin = lwr, ymax = upr), alpha = 0.3, color = NA, 
            fill = "grey") +
          geom_path(
            data = index %>% filter(species == k),
            aes(year, est), alpha = 0.5) +
          geom_path(
            data = tmp %>% filter(species == k),
            aes(year, est, color = model, group = fold), alpha = 0.5) +
          facet_wrap(~model, scales = "free") + 
          coord_cartesian(expand = FALSE, ylim = c(0, NA)) +
          scale_x_continuous(
            breaks = seq(1980, 2030, 4),
            labels = substr(seq(1980, 2030, 4),3,4)
          ) +
          labs(
            x = 'Year',
            y = ifelse(model_type == "Biomass", 'Biomass (kt)', 
                       'Abundance (millions)')
          ) +
          theme_bw() +
          theme(legend.position = "none")
      }),
    labels = unique(tmp$species), ncol = 1
  )
}
```

# Model selection {#model-selection}

The fixed effects depth, temperature, month, year and gear type were selected for the model selection step. Elements were removed from the full model to find the best fitting model formula. 

```{r model-selection}
#| code-fold: true
#| warning: false
#| message: false
#| results: hide
#| timeit: !expr if(recompile_models & model_selection) TRUE else NULL

if(recompile_models & model_selection) {
  
  model_list <- tibble(
    name = c(
      "Full model",
      "No year",
      "No year & depth",
      "No year & temperature",
      "No year & gear type",
      "No year & month",
      "No spatial",
      "No spatiotemporal",
      "Null model"
    ), 
    formula = list(
      formula(value ~ 0 + geartype + s(month) + fyear + s(depth) + s(temperature)),
      formula(value ~ 0 + geartype + s(month) + s(depth) + s(temperature)),
      formula(value ~ 0 + geartype + s(month) + s(temperature)),
      formula(value ~ 0 + geartype + s(month) + s(depth)),
      formula(value ~ 0 + s(month) + s(depth) + s(temperature)),
      formula(value ~ 0 + geartype + s(depth) + s(temperature)),
      formula(value ~ 0 + geartype + s(month) + fyear + s(depth) + s(temperature)),
      formula(value ~ 0 + geartype + s(month) + fyear + s(depth) + s(temperature)),
      formula(value ~ 1)
    ),
    spatial = c(
      "on",
      "on",
      "on",
      "on",
      "on",
      "on",
      "off",
      "on",
      "on"
    ),
    spatiotemporal = c(
      "iid", 
      "iid", 
      "iid", 
      "iid", 
      "iid", 
      "iid", 
      "iid", 
      "off",
      "iid"
    )
  )
  
  model_list <- model_list %>% 
    replicate_df("commonname", levels(dt$commonname)) %>% 
    relocate("commonname", .before = 1)
  
  if(!dir.exists("data/model_output/model_selection")) {
    dir.create("data/model_output/model_selection")
  }
  
  saveRDS(model_list, 
          file = "data/model_output/model_selection/model_selection_list.rds")
  
  future::plan("multisession", workers = n_sessions)
  options(future.globals.maxSize = 1000 * 1024^2, future.seed = TRUE)
  
  ## Fit the models
  
  message("Cross-validation (takes forever)")
  
  tmp <- pbapply::pblapply(1:nrow(model_list), function(i) {
    
    x <- dt %>% 
      filter(commonname == model_list$commonname[i]) %>%
      arrange(stationstartdate, year, missionid, serialnumber) %>% 
      droplevels()
    
    model_selection_mesh <- make_mesh(x, xy_cols = c("X", "Y"), cutoff = mesh_cutoff)
    clust <- sample(seq_len(10), size = nrow(x), replace = TRUE)
    
    print(i)
    term <- all.vars(model_list$formula[[i]])[-1]
    
    if(length(term) > 0) {
      continuous_predictor <- 
        sapply(term, function(j) {
          any(grepl(paste0("s\\(", j, "\\)"),
                    as.character(model_list$formula[[i]])))
        })
      
      continuous_columns <- names(continuous_predictor)[continuous_predictor]
      categorical_columns <- names(continuous_predictor)[!continuous_predictor]
      
      if(length(continuous_columns) > 0) {
        x <- x %>% 
          mutate(across(all_of(continuous_columns), as.numeric))
      }
      
      if(length(categorical_columns) > 0) {
        x <- x %>% 
          mutate(across(all_of(categorical_columns), as.factor))
      }
    }
    
    out <- try({
      sdmTMB_cv(
        model_list$formula[[i]],
        data = x,
        mesh = model_selection_mesh,
        fold_ids = clust,
        family = model_family,
        spatial = model_list$spatial[[i]],
        spatiotemporal = model_list$spatiotemporal[[i]],
        time = "year",
        control = sdmTMBcontrol(newton_loops = 2)
      )
    })
    
    if(inherits(out, "try-error")) {
      warning(paste(paste(model_list$commonname, 
                          model_list$name, sep = "_")[i], "crashed"))
    }
    
    # Save / return
    if(segmental_save) {
      saveRDS(
        list(mod = out, mesh = model_selection_mesh),
        file =
          paste0("data/model_output/model_selection/Model-selection-models-",
                 gsub("\\(.*\\)", "", model_family_name), "_",
                 paste(model_list$commonname, gsub(" ", "-", model_list$name), sep = "_")[i],
                 ".rds"),
        compress = "xz"
      )
    } else {
      return(list(mod = out, mesh = model_selection_mesh))
    }
  })
  
  # Load and make objects
  
  if(segmental_save) {
    model_selection_models <- 
      pbapply::pblapply(1:nrow(model_list), function(i) {
        readRDS(
          file =
            paste0("data/model_output/model_selection/Model-selection-models-",
                   gsub("\\(.*\\)", "", model_family_name), "_",
                   paste(model_list$commonname, 
                         gsub(" ", "-", model_list$name), sep = "_")[i],
                   ".rds")
        )$mod
      })
    
    model_selection_models <- 
      setNames(model_selection_models, 
               paste(model_list$name, 
                     model_list$commonname, sep = "_"))
    
    model_selection_mesh <- 
      pbapply::pblapply(
        unique(dt$commonname), function(k) {
          readRDS(
            dir("data/model_output/model_selection/", 
                pattern = paste0("Model-selection-models-",
                                 gsub("\\(.*\\)", "", model_family_name), "_", k),
                full.names = TRUE)[1])$mesh
          
          
        })
  } else {
    model_selection_models <- lapply(tmp, function(k) k$mod)
    model_selection_models <- 
      setNames(model_selection_models, 
               paste(model_list$name, 
                     model_list$commonname, sep = "_"))
    
    model_selection_mesh <- 
      lapply(c(1,10), function(i) tmp[[i]]$mesh)
    
    model_selection_mesh <- 
      setNames(model_selection_mesh, levels(dt$commonname))
    
    if(save_model_output) {
      saveRDS(model_selection_models, 
              file = paste0(
                "data/model_output/model_selection/Model-selection-models-",
                gsub("\\(.*\\)", "", model_family_name), ".rds"),
              compress = "xz"
      ) %>% system.time()
    }
    
    saveRDS(model_selection_mesh,
            file = paste0(
              "data/model_output/model_selection/Model-selection-meshes-",
              gsub("\\(.*\\)", "", model_family_name), ".rds"),
            compress = "xz"
    )  
  }
  
  ## Save required components
  model_selection_params <- 
    list(
      model_names = names(model_selection_models)
    )
  
  saveRDS(
    model_selection_params,
    file = 
      paste0(
        "data/model_output/model_selection/Model-selection-parameters-",
        gsub("\\(.*\\)", "", model_family_name), ".rds")
  )
  
} else {
  model_list <- 
    readRDS(
      file = "data/model_output/model_selection/model_selection_list.rds")
  
  model_selection_mesh <- 
    readRDS(paste0(
      "data/model_output/model_selection/Model-selection-meshes-",
      gsub("\\(.*\\)", "", model_family_name), ".rds")
    )
  
  model_selection_params <-
    readRDS(paste0(
      "data/model_output/model_selection/Model-selection-parameters-",
      gsub("\\(.*\\)", "", model_family_name), ".rds")
    )
}
```  

```{r}
#| label: tbl-model-selection
#| timeit: !expr if(recompile_models & model_selection) TRUE else NULL
#| tbl-cap: !expr paste0("Model selection for ", sp_name, " based on ten-fold cross-validation. The models contain spatial and spatiotemporal random fields, unless specified otherwise. Each model is compared to the best fitting single component model ($\\Delta$NLL and $\\Delta$NLL % columns). The negative log-likelihood value is given in the NLL column and AIC difference to the best fitting model in the $\\Delta$AIC column.")

if(recompile_models & model_selection) {
  
  modsel_tab <- 
    pbapply::pblapply(seq_along(model_selection_models), function(i) {
      if(inherits(model_selection_models[[i]], "try-error")) {
        tibble(
          species = case_when(
            model_list$commonname[i] == "snabeluer" ~ "Beaked redfish", 
            model_list$commonname[i] == "vanlig uer" ~ "Golden redfish"
          ),
          model = model_list$name[i],
          formula = paste(trimws(format(model_list$formula[[i]])), 
                          collapse = ""),
          spatial = model_list$spatial[[i]],
          spatiotemporal = model_list$spatiotemporal[[i]],
          crashed = TRUE,
          converged = FALSE,
          sanity = FALSE,
          loglik = NA,
          AIC = NA
        )
      } else {
        tibble(
          species = case_when(
            model_list$commonname[i] == "snabeluer" ~ "Beaked redfish", 
            model_list$commonname[i] == "vanlig uer" ~ "Golden redfish"
          ),
          model = model_list$name[i],
          formula = paste(trimws(format(
            model_selection_models[[i]]$models[[1]]$formula[[1]])), 
            collapse = ""),
          spatial = unique(model_selection_models[[i]]$models[[1]]$spatial),
          spatiotemporal = unique(
            model_selection_models[[i]]$models[[1]]$spatiotemporal),
          crashed = FALSE,
          converged = model_selection_models[[i]]$converged,
          sanity = all(unlist(sanity(model_selection_models[[i]]$models[[1]],
                                     silent = TRUE))),
          loglik = model_selection_models[[i]]$sum_loglik,
          AIC = median(sapply(model_selection_models[[i]]$models, AIC))
        )
      }
    }, cl = n_cores) %>% 
    bind_rows() %>% 
    group_by(species) %>% 
    arrange(-loglik) %>% 
    mutate(
      deltaNLL = loglik - max(loglik, na.rm = TRUE),
      deltaNLL_pr = 100*(1-loglik/max(loglik, na.rm = TRUE)),
      deltaAIC = AIC - min(AIC, na.rm = TRUE),
      deltaAIC_pr = 100*(AIC/min(AIC, na.rm = TRUE) - 1)
    ) %>% 
    ungroup()
  
  ## Save
  write_csv(
    modsel_tab,
    file =
      paste0("data/model_output/model_selection/Model-selection-table-", 
             gsub("\\(.*\\)", "", model_family_name),
             ".csv")
  )
} else {
  modsel_tab <- read_csv(
    paste0("data/model_output/model_selection/Model-selection-table-", 
           gsub("\\(.*\\)", "", model_family_name),
           ".csv")
  )
}

## Print the table
modsel_tab %>% 
  relocate(formula, .after = last_col()) %>%
  dplyr::select(-spatial, -spatiotemporal, -AIC, -deltaAIC_pr) %>% 
  mutate(loglik = abs(loglik), 
         deltaNLL = abs(deltaNLL),
         deltaNLL_pr = abs(deltaNLL_pr)) %>% 
  rename(
    "Species" = "species", "Model" = "model", "Formula" = "formula", 
    "Crashed" = "crashed", "Converged" = "converged", "Sanity test" = "sanity", 
    "NLL" = "loglik", "ΔNLL" = "deltaNLL", "ΔNLL %" =  "deltaNLL_pr", 
    "ΔAIC" = "deltaAIC"
  ) %>% 
  make_table(digits = c(0,0,0,0,0,0,0,2,0,0))

# More compact alternative:
# modsel_tab %>% 
#   mutate(deltaNLL = abs(deltaNLL)) %>% 
#   dplyr::select(species, model, formula, sanity, deltaNLL) %>%
#   ungroup() %>% 
#   pivot_wider(id_cols = c(model, formula, sanity), 
#               names_from = species,
#               values_from = deltaNLL) %>% 
#   filter(!model %in% "Full model s(logdepth)") %>% 
#   rename("Model" = "model", "Formula" = "formula", 
#          "Sanity test" = "sanity") %>%
#   make_table(digits = 2)

```

```{r}
#| label: fig-model-selection
#| fig-cap: !expr paste("Model selection for", sp_name, "based on ten-fold cross-validation. The y-axis lists the model. The x-axis shows the negative log-likelihood for each model, where higher values indicate better model fit. Symbol color and shape reflect model diagnostics; black circles denote models without issues, while grey symbols indicate models that failed the sanity check or did not converge.")

tmp <- modsel_tab %>%
  group_by(species) %>% 
  mutate(status = case_when(
    !sanity & !converged ~ "Nonconvergent &\nsanity test issues",
    !sanity ~ "Sanity test issues",
    !converged ~ "Nonconvergent",
    .default = "All OK"
  ), .after = "sanity") %>% 
  mutate(deltaNLL = abs(loglik - max(loglik, na.rm = TRUE))) %>% 
  mutate(deltaNLL = abs(ifelse(crashed, Inf, deltaNLL))) %>% 
  rename("value" = "deltaNLL") %>% 
  ungroup()

tmp <- tmp %>% 
  mutate(
    model = factor(
      model,
      levels = tmp %>% 
        group_by(model) %>% reframe(value = mean(value)) %>% 
        arrange(value) %>% pull(model) %>% rev())
  )

ggplot(tmp, aes(y = model, x = value, color = status, shape = status)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c(16,15,17,18)) +
  scale_color_manual(values = c("black", rep("grey", 3))) +
  facet_grid(~species, scales = "free") +
  # scale_y_continuous(transform = "sqrt") + 
  labs(x = "Negative log likelihood difference",
       y = "Model", shape = "Model diagnostics", color = "Model diagnostics") +
  theme(legend.position = "bottom",
        panel.grid.major = element_line(color = "grey", linewidth = LS(0.5)))
```

## Survey indices

```{r mod-sel-survey-indices}
#| timeit: !expr if(recalculate_index & model_selection) TRUE else NULL

if(run_index) {
  if(recalculate_index & model_selection) {
    
    proj_dt <- read_rds("data/model_output/prediction-grid-data.rds") %>% 
      distinct(X, Y, depth, year) 
    
    ## Prediction grid cell side length
    grid_side_size <- attributes(proj_dt)$grid_side_size # kilometers due to UTM projection and division by 1000
    
    if(development_mode) {
      proj_dt <- proj_dt %>% 
        filter(year > 2003) %>% 
        mutate(sample = rep(1:4, length.out = nrow(.))) %>% 
        filter(sample == 1) %>% 
        dplyr::select(-sample)
    }
    
    models <- model_selection_models
    
    index <- pbapply::pblapply(seq_along(models), function(i) {
      
      # print(i)
      
      k <- models[[i]]
      if(inherits(k, "try-error")) return(NULL)
      k_name <- names(models)[i]
      
      terms <- trimws(unlist(strsplit(
        as.character(k$models[[1]]$formula[[1]])[3], "\\+")))
      terms <- terms[terms != "0"]
      terms <- terms[!grepl(":", terms)]
      
      pred_dt <- proj_dt
      
      if(grepl("Null model", k_name)) {
        pred_dt <- pred_dt %>% distinct(X, Y, year)
      } else {
        fyear <- any(grepl("fyear", terms))
        geartype <- any(grepl("geartype", terms))
        logdepth <- any(grepl("logdepth", terms))
        cruiseseries <- any(grepl("cruiseseries", terms))
        season <- any(grepl("season", terms))
        depth <- any(grepl("depth", terms))
        month <- any(grepl("month", terms))
        smonth <- any(grepl("s\\(month\\)", terms))
        yday <- any(grepl("yday", terms))
        distance <- any(grepl("distance", terms))
        commonname <- any(grepl("commonname", terms))
        temperature <- any(grepl("temperature", terms))
        
        if(!any(c(depth, temperature))) pred_dt <- pred_dt %>% distinct(X, Y, year)
        if(logdepth) pred_dt <- pred_dt %>% mutate(logdepth = log(depth))
        if(geartype) pred_dt <- pred_dt %>% mutate(geartype = "Shrimp")
        if(fyear) pred_dt <- pred_dt %>% mutate(fyear = as.factor(year))
        if(season) pred_dt <- pred_dt %>% mutate(season = "H2")
        if(month) {pred_dt <- pred_dt %>% 
          replicate_df("month", if(smonth) as.integer(8) else factor(8))}
        if(yday) pred_dt <- pred_dt %>% mutate(yday = 225)
        if(distance) pred_dt <- pred_dt %>% mutate(distance = 1) # assuming 3 knot trawling speed for 20 min
        
        if(include_temperature & temperature) {
          
          ### Temperature model ###
          load("data/model_output/temperature-model-fit.rda")
          
          ## Predict using the data
          
          pred_dt <- predict(
            fit[[1]]$mod,
            newdata = pred_dt %>% 
              mutate(fyear = as.factor(year),
                     month = as.integer(8)),
            type = "response",
            model = NA,
            return_tmb_object = FALSE) %>% 
            dplyr::select(-est_non_rf, -est_rf, -omega_s, -epsilon_st) %>% 
            rename("temperature" = "est") 
          
          if(!smonth) {
            pred_dt <- pred_dt %>% mutate(month = factor(month))
          }
        }
        
      }
      
      ## Slower more official way with CIs
      # lapply(seq_along(k$models), function(j) {
      #   predict(k$models[[j]], newdata = proj_dt, type = "response", model = NA,
      #           return_tmb_object = TRUE) %>% 
      #     get_index(
      #       bias_correct = FALSE,
      #       area = (grid_side_size/1.852)^2) %>% # grid cell area in square nautical miles
      #     mutate(est = est/1e6, # convert kg to kilotonnes
      #            lwr = lwr/1e6,
      #            upr = upr/1e6, 
      #            cv = sqrt(exp(se^2) - 1)) %>% 
      #     mutate(model = k_name, .before = 1) %>% 
      #     mutate(fold = j)
      # }) %>% 
      #   bind_rows()
      
      ## Quicker way
      mod_pred <- 
        lapply(seq_along(k$models), function(j) {
          predict(k$models[[j]], newdata = pred_dt, type = "response", 
                  model = NA, return_tmb_object = FALSE) %>%
            mutate(fold = j)
        })
      
      mod_pred %>%
        bind_rows() %>%
        group_by(fold, year, X, Y) %>%
        reframe(est = mean(est)) %>%
        group_by(fold, year) %>%
        reframe(
          # Y = weighted.mean(Y, est),
          # X = weighted.mean(X, est),
          # depth = weighted.mean(depth, est),
          est = (sum(est)*(grid_side_size/1.852)^2)/1e6
        ) %>%
        mutate(lwr = NA, upr = NA, cv = NA, se = NA) %>%
        mutate(commonname = unique(k$data$commonname),
               model = k_name, 
               .before = 1)
    }, cl = n_cores)
    
    names(index) <- names(models)
    
    write_rds(
      index,
      file =
        paste0("data/model_output/model_selection/Model-selection-indices-", 
               gsub("\\(.*\\)", "", model_family_name),
               ".rds"), 
      compress = "xz"
    )
    
    ms_indices <- index
    
  } else {
    
    proj_dt <- read_rds("data/model_output/prediction-grid-data.rds") %>% 
      distinct(X, Y, depth, year)
    
    ms_indices <- read_rds(
      paste0("data/model_output/model_selection/Model-selection-indices-", 
             gsub("\\(.*\\)", "", model_family_name),
             ".rds"), 
    ) 
  }
}
```

```{r}
#| code-fold: true
#| label: fig-mod-sel-index
#| fig-height: !expr ceiling(length(ms_indices)/3)*2.5
#| fig-cap: !expr paste0("Survey indices estimated using multiple-component models for ", sp_name, " using spatiotemporal generalized linear mixed effects models with the sdmTMB package (colored solid line). Black lines with shading indicate the full biomass models with multiple components; 0 + geartype + s(month) + s(depth) + s(temperature). ", value_text_name, " is used as the response variable, and the models use meshes with ", paste(sapply(model_selection_mesh, function(k) k$mesh$n), collapse = " and "), " vertices (knots), as well as ", model_family_name," model family. Predictions are made to a grid with ", proj_dt %>% distinct(X,Y) %>% nrow(), " cells with a side of ", attributes(proj_dt)$grid_side_size, " km. Each line represents one fold from the 10-fold cross-validation. See Table 1 for more detailed model specification.")

if(run_index) {
  tmp <- lapply(ms_indices, function(k) {
    if(inherits(k, "try-error")) return(NULL)
    if(any(!is.na(k$upr))) {
      if(max(k$upr, na.rm = TRUE) > 2*max(k$est, na.rm = TRUE)) {
        k[k$upr > 1.5*max(k$est, na.rm = TRUE),]$upr <- 
          ceiling(1.5*max(k$est, na.rm = TRUE))
      }}
    k
  }) %>%
    bind_rows() %>% 
    filter(!is.na(year)) %>% 
    mutate(model = gsub("_snabeluer|_vanlig uer", "", model)) %>% 
    # mutate(model = factor(model, levels = names(ms_indices))) %>% 
    mutate(
      species = case_when(
        commonname == "snabeluer" ~ "Beaked redfish", 
        commonname == "vanlig uer" ~ "Golden redfish"
      ),
      .after = "commonname")
  
  load("data/model_output/index-bm.rda")
  
  index <- index %>% 
    bind_rows() %>% 
    rename("species" = "length_group")
  
  cowplot::plot_grid(
    plotlist = 
      lapply(unique(tmp$species), function(k) {
        ggplot() +
          geom_ribbon(
            data = index %>% filter(species == k),
            aes(year, est, ymin = lwr, ymax = upr), alpha = 0.3, color = NA, 
            fill = "grey") +
          geom_path(
            data = index %>% filter(species == k),
            aes(year, est), alpha = 0.5) +
          geom_path(
            data = tmp %>% filter(species == k),
            aes(year, est, color = model, group = fold), alpha = 0.5) +
          facet_wrap(~model, scales = "free") + 
          coord_cartesian(expand = FALSE, ylim = c(0, NA)) +
          scale_x_continuous(
            breaks = seq(1980, 2030, 4),
            labels = substr(seq(1980, 2030, 4),3,4)
          ) +
          labs(
            x = 'Year',
            y = ifelse(model_type == "Biomass", 'Biomass (kt)', 
                       'Abundance (millions)')
          ) +
          theme_bw() +
          theme(legend.position = "none")
      }),
    labels = unique(tmp$species), ncol = 1
  )
}
```

# Family selection

The best fitting model 'No year' with formula rho_biomass ~ 0 + geartype + s(month) + s(depth) + s(temperature), and ar1 spatiotemporal random fields were used in testing the model family influence on survey indices. We tested [delta lognormal, delta gamma](https://pbs-assess.github.io/sdmTMB/articles/delta-models.html), [delta gamma poisson link](https://pbs-assess.github.io/sdmTMB/articles/poisson-link.html) [@thorson2018], and [tweedie](https://pbs-assess.github.io/sdmTMB/articles/model-description.html#tweedie) distribution families:

```{r family-selection}
#| timeit: !expr if(recompile_models & family_selection) TRUE else NULL

if(recompile_models & family_selection) {
  
  model_family_list <- list(
    "delta_lognormal" = sdmTMB::delta_lognormal(), 
    "delta_gamma" = sdmTMB::delta_gamma(),
    "poisson-link" = delta_gamma(type = "poisson-link"),
    "tweedie" = sdmTMB::tweedie()
  )
  
  future::plan("multisession", workers = n_sessions)
  options(future.globals.maxSize = 1000 * 1024^2, future.seed = TRUE)
  
  saveRDS(model_family_list,
          file = paste0(
            "data/model_output/model_selection/Model-family-comparison-list-",
             switch(response_variable, 
                    "catch biomass" = "cw", 
                    "biomass density" = "bm", 
                    "catch abundance" = "cc", 
                    "abundance density" = "ab"),
             ".rds")
          )
  
  ## Fit the models 
  
  message("Model fitting")
  
  model_family_models <- 
    pbapply::pblapply(unique(dt$commonname), function(k) {
      
      set.seed(7) # to make the mesh creation consequent
      
      x <- dt %>% 
        filter(commonname == k) %>% 
        droplevels()
      
      mesh <- make_mesh(x, xy_cols = c("X", "Y"), cutoff = mesh_cutoff)
      
      models <- pbapply::pblapply(
        seq_along(model_family_list), function(j) {
          out <- try({
            sdmTMB(
              value ~ 0 + geartype + s(month) + s(depth) + s(temperature),
              data = x,
              mesh = mesh,
              family = model_family_list[[j]],
              spatial = "on",
              spatiotemporal = 'ar1',
              time = "year",
              control = sdmTMBcontrol(newton_loops = 2)
            )
          })
          
          if(inherits(out, "try-error")) {
            warning(paste(paste(k, names(model_family_list[j]), 
                                sep = "_"), "crashed"))
          }
          
          # Save / return
          if(segmental_save) {
            saveRDS(
              out,
              file =
                paste0(
                  "data/model_output/model_selection/Family-selection-models-",
                  paste(k, 
                        gsub(" ", "-", names(model_family_list[j])),
                        sep = "_"),
                  ".rds"),
              compress = "xz"
            )
          } else {
            return(out)
          }   
        }, cl = length(model_family_list)
      )
      
      # Load and make objects
      
      if(segmental_save) {
        models <- 
          pbapply::pblapply(seq_along(model_family_list), function(j) {
            readRDS(
              file =
                paste0(
                  "data/model_output/model_selection/Family-selection-models-",
                  paste(k, 
                        gsub(" ", "-", names(model_family_list[j])),
                        sep = "_"),
                  ".rds")
            )
          })
      }
      
      models <- setNames(models, names(model_family_list))
      
      list(
        models = models, model_families = names(model_family_list),
        species = k, mesh = mesh, mesh_n = mesh$mesh$n, 
        mesh_cutoff = mesh_cutoff, crs = dt_crs)
    }) 
  
  model_family_models <- 
    setNames(
      model_family_models, 
      recode(unique(dt$commonname), 
             "snabeluer" = "Beaked redfish", "vanlig uer" = "Golden redfish")
    )
  
  attributes(model_family_models) <- 
    c(attributes(model_family_models),
      list(
        response_variable = response_variable, 
        value_name = value_name,
        value_text_name = value_text_name,
        value_text_unit = value_text_unit)
    )
  
  message("Save")
  
  if(save_model_output) {
    saveRDS(
      model_family_models, 
      file = 
        paste0("data/model_output/model_selection/Family-selection-models-",
             switch(response_variable, 
                    "catch biomass" = "cw", 
                    "biomass density" = "bm", 
                    "catch abundance" = "cc", 
                    "abundance density" = "ab"),
             ".rds"),
      compress = "xz"
    )
  }
  
  family_selection_params <- 
    list(
      formula = format(model_family_models[[1]]$models[[1]]$formula[[1]]),
      spatiotemporal = unique(model_family_models[[1]]$models[[1]]$spatiotemporal)
    )
  
  saveRDS(
    family_selection_params,
    file = 
      paste0("data/model_output/model_selection/Family-selection-parameters-",
             switch(response_variable, 
                    "catch biomass" = "cw", 
                    "biomass density" = "bm", 
                    "catch abundance" = "cc", 
                    "abundance density" = "ab"),
             ".rda")
  )
  
} else {
  family_selection_params <- 
    readRDS(
      paste0(
        "data/model_output/model_selection/Family-selection-parameters-",
             switch(response_variable, 
                    "catch biomass" = "cw", 
                    "biomass density" = "bm", 
                    "catch abundance" = "cc", 
                    "abundance density" = "ab"),
             ".rda")
    )
  
  model_family_list <-
    readRDS(
      paste0(
        "data/model_output/model_selection/Model-family-comparison-list-",
             switch(response_variable, 
                    "catch biomass" = "cw", 
                    "biomass density" = "bm", 
                    "catch abundance" = "cc", 
                    "abundance density" = "ab"),
             ".rds")
    )
}
```

```{r}
#| label: tbl-family-selection
#| timeit: !expr if(recompile_models & family_selection) TRUE else NULL
#| tbl-cap: !expr paste0("Comparison of model families using the ",  family_selection_params$formula, " formula and ", family_selection_params$spatiotemporal, " spatiotemporal random fields. The table presents the negative log-likelihood (NLL) and Akaike Information Criterion (AIC) values, along with their differences among model family alternatives for each species. The Sanity column reports the results of the sdmTMB::sanity_test() function, which ensures convergence, a positive definite Hessian matrix, and small parameter gradients.")

if(recompile_models & family_selection) {
  
  modtab <- lapply(seq_along(model_family_models), function(i) {
    lapply(seq_along(model_family_models[[i]]$models), function(j) {
      tibble(
        commonname = model_family_models[[i]]$species,
        family = names(model_family_models[[i]]$models[j]),
        formula = format(model_family_models[[i]]$models[[j]]$formula[[1]]),
        spatiotemporal = unique(model_family_models[[i]]$models[[j]]$spatiotemporal),
        converged = model_family_models[[i]]$models[[j]]$model$convergence == 0,
        sanity = all(unlist(sanity(model_family_models[[i]]$models[[j]], 
                                   silent = TRUE))),
        NLL = round(c(model_family_models[[i]]$models[[j]]$model$objective), 1),
        AIC = AIC(model_family_models[[i]]$models[[j]])
      )
    }) %>% 
      bind_rows() %>% 
      mutate(species = names(model_family_models[i]), .before = 1)
  }) %>% 
    bind_rows() %>% 
    group_by(species) %>% 
    mutate(deltaNLL = NLL - min(NLL),
           deltaAIC = AIC - min(AIC)) %>% 
    arrange(species, deltaNLL)
  
  write_delim(
    modtab,
    file = 
      paste0("data/model_output/model_selection/Family-selection-table-",
             switch(response_variable, 
                    "catch biomass" = "cw", 
                    "biomass density" = "bm", 
                    "catch abundance" = "cc", 
                    "abundance density" = "ab"),
             ".txt")
  )
} else {
  modtab <- read_delim(
    paste0("data/model_output/model_selection/Family-selection-table-",
           switch(response_variable, 
                  "catch biomass" = "cw", 
                  "biomass density" = "bm", 
                  "catch abundance" = "cc", 
                  "abundance density" = "ab"),
           ".txt")
  )
}

## Table

## Print
modtab %>% 
  dplyr::select(-formula, -spatiotemporal) %>% 
  rename_with(~sub("(^[[:alpha:]])", "\\U\\1", .x, perl = TRUE)) %>% 
  rename("ΔNLL" = "DeltaNLL",
         "ΔAIC" =  "DeltaAIC") %>% 
  knitr::kable(escape = FALSE, digits = 0) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "condensed"))
```

```{r}
#| label: fig-family-selection
#| fig-cap: "Negative log-likelihood difference among model families for each species from Table 1. The x-axis shows the negative log-likelihood for each model, where higher values indicate better model model_family_models. Symbol color and shape reflect model diagnostics: black circles denote models without issues, while grey symbols indicate models that failed the sanity check or did not converge."

modtab %>%
  mutate(family = factor(family, levels = rev(unique(.$family)))) %>% 
  mutate(status = case_when(
    !sanity & !converged ~ "Nonconvergent &\nsanity test issues",
    !sanity ~ "Sanity test issues",
    !converged ~ "Nonconvergent",
    .default = "All OK"
  ), .after = "sanity") %>% 
  ggplot(aes(family, deltaNLL, shape = status, color = status)) + 
  geom_point(size = 3) +
  coord_flip() + 
  scale_shape_manual(values = c(16,15,17)) +
  scale_color_manual(values = c("black", rep("grey", 3))) +
  facet_wrap(~ species, axes = "all", axis.labels = "margins") +
  labs(y = "Negative log likelihood difference",
       x = "Model family", shape = "Model diagnostics", 
       color = "Model diagnostics") +
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-resid-qq
#| fig-height: !expr 3.5*ceiling(length(model_family_list)/2)
#| timeit: !expr if(recompile_models & family_selection) TRUE else NULL
#| results: hide
#| fig-cap: Quantile–quantile plots for beaked (left) and golden redfish (right) model families. Model families are arranged in rows, with submodels displayed in columns.

if(recompile_models & family_selection) {
  set.seed(12)
  
  qq_dat <- 
    pbapply::pblapply(seq_along(model_family_models), function(j) {
      pbapply::pblapply(seq_along(model_family_models[[j]]$models), function(i) {
        lapply(seq_along(
          model_family_models[[j]]$models[[i]]$family$family), function(l) {
            rq_res <- residuals(model_family_models[[j]]$models[[i]], 
                                type = "mle-mvn", model = l)
            rq_res <- rq_res[is.finite(rq_res)] # some Inf
            tibble(
              species = names(model_family_models[j]),
              family = names(model_family_models[[j]]$models)[i],
              model = model_family_models[[j]]$models[[i]]$family$family[l],
              x = rq_res
            )
          }) %>% bind_rows()
      }, cl = length(model_family_models[[j]]$models)) %>% 
        bind_rows() %>% 
        mutate(family = factor(family, levels = unique(modtab$family)))
    }) 
  
  # Save
  
  saveRDS(
    qq_dat, 
    file = paste0(
      "data/model_output/model_selection/Family-selection-residuals-",
      switch(response_variable, 
             "catch biomass" = "cw", 
             "biomass density" = "bm", 
             "catch abundance" = "cc", 
             "abundance density" = "ab"),
      ".rds"),
    compress = "xz"
  )
} else {
  qq_dat <- readRDS(
    paste0(
      "data/model_output/model_selection/Family-selection-residuals-",
      switch(response_variable, 
             "catch biomass" = "cw", 
             "biomass density" = "bm", 
             "catch abundance" = "cc", 
             "abundance density" = "ab"),
      ".rds")
  )
}

cowplot::plot_grid(
  plotlist = 
    lapply(qq_dat, function(k) {
      k %>% 
        ggplot(aes(sample = x)) +
        geom_qq_line() +
        geom_qq(shape = 21, color = "grey60") +
        facet_wrap(~family + model, ncol = 2) +
        labs(
          title = unique(k$species),
          x = "Theoretical Quantiles",
          y = "Sample Quantiles"
        )
    }),
  ncol = 2
)
```

## Survey indices

```{r fam-sel-survey-indices}
#| timeit: !expr if(recalculate_index & family_selection) TRUE else NULL

if(run_index) {
  if(recalculate_index & family_selection) {
    
    proj_dt <- read_rds("data/model_output/prediction-grid-data.rds") %>% 
      distinct(X, Y, depth, year)
    
    grid_side_size <- attributes(proj_dt)$grid_side_size # kilometers due to UTM projection and division by 1000
    
    index <- pbapply::pblapply(seq_along(model_family_models), function(i) {
      print(i)
      pbapply::pblapply(seq_along(model_family_models[[i]]$models), function(j) {
        print(j)
        k <- model_family_models[[i]]$models[[j]]
        k_name <- names(model_family_models[[i]]$models[j])
        k_sp <- names(model_family_models[i])
        
        pred_dt <- proj_dt
        
        terms <- trimws(unlist(strsplit(as.character(k$formula[[1]])[3], "\\+")))
        terms <- terms[terms != "0"]
        terms <- terms[!grepl(":", terms)]
        
        fyear <- any(grepl("fyear", terms))
        logdepth <- any(grepl("logdepth", terms))
        cruiseseries <- any(grepl("cruiseseries", terms))
        geartype <- any(grepl("geartype", terms))
        season <- any(grepl("season", terms))
        depth <- any(grepl("depth", terms))
        month <- any(grepl("month", terms))
        smonth <- any(grepl("s\\(month\\)", terms))
        yday <- any(grepl("yday", terms))
        distance <- any(grepl("distance", terms))
        commonname <- any(grepl("commonname", terms))
        temperature <- any(grepl("temperature", terms))
        
        if(!any(c(depth, temperature))) pred_dt <- pred_dt %>% distinct(X, Y, year)
        if(logdepth) pred_dt <- pred_dt %>% mutate(logdepth = log(depth))
        if(geartype) pred_dt <- pred_dt %>% mutate(geartype = "Shrimp")
        if(fyear) pred_dt <- pred_dt %>% mutate(fyear = as.factor(year))
        if(season) pred_dt <- pred_dt %>% mutate(season = "H2")
        if(month) {pred_dt <- pred_dt %>% 
          replicate_df("month", if(smonth) as.integer(8) else factor(8))}
        if(yday) pred_dt <- pred_dt %>% mutate(yday = 225)
        if(distance) pred_dt <- pred_dt %>% mutate(distance = 1) # assuming 3 knot trawling speed for 20 min
        
        if(include_temperature & temperature) {
          
          ### Temperature model ###
          load("data/model_output/temperature-model-fit.rda")
          
          ## Predict using the data
          
          pred_dt <- predict(
            fit[[1]]$mod,
            newdata = pred_dt %>% 
              mutate(fyear = as.factor(year),
                     month = as.integer(8)),
            type = "response",
            model = NA,
            return_tmb_object = FALSE) %>% 
            dplyr::select(-est_non_rf, -est_rf, -omega_s, -epsilon_st) %>% 
            rename("temperature" = "est") 
          
          if(!smonth) {
            pred_dt <- pred_dt %>% mutate(month = factor(month))
          }
          
        }
        
        
        if(FALSE) {# Time consuming alternative. Takes hours to days, and tends to crash.
          mod_pred <- 
            predict(k, newdata = pred_dt, type = "response", model = NA,
                    return_tmb_object = TRUE) 
          
          get_index(
            obj = mod_pred,
            bias_correct = TRUE,
            area = (grid_side_size/1.852)^2) %>% # grid cell area in square nautical miles
            mutate(est = est/1e6, # convert kg to kilotonnes
                   lwr = lwr/1e6,
                   upr = upr/1e6, 
                   cv = sqrt(exp(se^2) - 1)) %>% 
            mutate(
              species = k_sp,
              family = k_name, 
              .before = 1)
        } else { # Quick alternative, no CIs
          
          mod_pred <- 
            predict(k, newdata = pred_dt, type = "response", model = NA,
                    return_tmb_object = FALSE) 
          
          mod_pred %>%
            bind_rows() %>%
            group_by(year, X, Y) %>%
            reframe(est = mean(est)) %>%
            group_by(year) %>%
            reframe(
              # Y = weighted.mean(Y, est),
              # X = weighted.mean(X, est),
              # depth = weighted.mean(depth, est),
              est = (sum(est)*(grid_side_size/1.852)^2)/1e6
            ) %>%
            mutate(lwr = NA, upr = NA, cv = NA, se = NA) %>%
            mutate(
              species = k_sp,
              family = k_name,
              .before = 1)
        }
      },
      cl = length(model_family_list)
      ) %>% bind_rows()
    }, cl = 2) 
    
    names(index) <- names(model_family_models)
    
    write_rds(
      index,
      file =
        paste0(
          "data/model_output/model_selection/Family-selection-indices-",
          switch(response_variable, 
                 "catch biomass" = "cw", 
                 "biomass density" = "bm", 
                 "catch abundance" = "cc", 
                 "abundance density" = "ab"),
          ".rds")
    )
    
    fs_indices <- index
  } else {
    proj_dt <- read_rds("data/model_output/prediction-grid-data.rds") %>% 
      distinct(X, Y, depth, year)
    
    fs_indices <- read_rds(
      paste0(
        "data/model_output/model_selection/Family-selection-indices-",
        switch(response_variable, 
               "catch biomass" = "cw", 
               "biomass density" = "bm", 
               "catch abundance" = "cc", 
               "abundance density" = "ab"),
        ".rds")
    )
  }
}
```

```{r}
#| label: fig-fam-sel-index
#| fig-height: 6
#| fig-cap: !expr paste0("Survey index for the two redfish species depending on selected model family. Colored lines indicate alternative models", ifelse(any(!is.na(fs_indices[[1]]$upr)), " and shading 95% confidence intervals.", ".")) 

if(run_index) {
  tmp <- lapply(fs_indices, function(k) {
    if(any(!is.na(k$upr))) {
      if(max(k$upr, na.rm = TRUE) > 2*max(k$est, na.rm = TRUE)) {
        k[k$upr > 1.5*max(k$est, na.rm = TRUE),]$upr <- 
          ceiling(1.5*max(k$est, na.rm = TRUE))
      }}
    k
  }) %>%
    bind_rows() 
  
  tmp %>% 
    ggplot(aes(year, est, color = family, fill = family)) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), color = NA, alpha = 0.5) +
    geom_path() +
    facet_wrap(~species, scale = "free_y", ncol = 1) +
    scale_x_continuous(
      breaks = seq(1980, 2030, 4),
      labels = substr(seq(1980, 2030, 4),3,4)
    ) +
    # coord_cartesian(expand = FALSE, ylim = c(0, 1.05*max(tmp$est, na.rm = TRUE))) +
    labs(
      x = 'Year',
      y = ifelse(model_type == "Biomass", 'Biomass (kt)', 
                 'Abundance (millions)'),
      color = "Model\nfamily", fill = "Model\nfamily"
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
}
```

# Spatiotemporal field selection

The best fitting model 'No year' with formula rho_biomass ~ 0 + geartype + s(month) + s(depth) + s(temperature), and delta lognormal model family were used to rank spatiotemporal random field structures. 

```{r spatiotemporal-selection}
#| code-fold: true
#| warning: false
#| message: false
#| results: hide
#| timeit: !expr if(recompile_models & spatiotemporal_selection) TRUE else NULL

if(recompile_models & spatiotemporal_selection) {
  
  spatiotemporal_list <- tibble(
    name = c(
      "Independent",
      "Autoregressive",
      "Random walk",
      "Off"
    ),
    spatiotemporal = c(
      "iid",
      "ar1",
      "rw",
      "off"
    )
  )
  
  spatiotemporal_list <- spatiotemporal_list %>% 
    replicate_df("commonname", levels(dt$commonname)) %>% 
    relocate("commonname", .before = 1)
  
  if(!dir.exists("data/model_output/model_selection")) {
    dir.create("data/model_output/model_selection")
  }
  
  saveRDS(spatiotemporal_list, 
          file = "data/model_output/model_selection/spatiotemporal_list.rds")
  
  future::plan("multisession", workers = n_sessions)
  options(future.globals.maxSize = 1000 * 1024^2, future.seed = TRUE)
  
  ## Fit the models 
  
  message("Cross-validation (takes an hour on a powerful computer)")
  
  tmp <- pbapply::pblapply(1:nrow(spatiotemporal_list), function(i) {
    
    x <- dt %>% 
      filter(commonname == spatiotemporal_list$commonname[i]) %>%
      arrange(stationstartdate, year, missionid, serialnumber) %>% 
      droplevels()
    
    spatiotemporal_mesh <- make_mesh(x, xy_cols = c("X", "Y"), 
                                     cutoff = mesh_cutoff)
    clust <- sample(seq_len(10), size = nrow(x), replace = TRUE)
    
    out <- try({
      sdmTMB_cv(
        formula(value ~ 0 + geartype + s(month) + s(depth) + s(temperature)),
        data = x,
        mesh = spatiotemporal_mesh,
        fold_ids = clust,
        family = model_family,
        spatial = "on",
        spatiotemporal = spatiotemporal_list$spatiotemporal[[i]],
        time = "year",
        control = sdmTMBcontrol(newton_loops = 2)
      )
    })
    
    if(inherits(out, "try-error")) {
      warning(paste(paste(spatiotemporal_list$commonname, 
                          spatiotemporal_list$name, sep = "_")[i], "crashed"))
    }
    
    # Save / return
    if(segmental_save) {
      saveRDS(
        list(mod = out, mesh = spatiotemporal_mesh),
        file =
          paste0(
          "data/model_output/model_selection/Spatiotemporal-selection-models-",
                 gsub("\\(.*\\)", "", model_family_name), "_",
                 paste(spatiotemporal_list$commonname, gsub(" ", "-", spatiotemporal_list$name), sep = "_")[i],
                 ".rds"),
        compress = "xz"
      )
    } else {
      return(list(mod = out, mesh = spatiotemporal_mesh))
    }
  })
  
  # Load and make objects
  
  if(segmental_save) {
    spatiotemporal_selection_models <- 
      pbapply::pblapply(1:nrow(spatiotemporal_list), function(i) {
        readRDS(
          file =
            paste0(
            "data/model_output/model_selection/Spatiotemporal-selection-models-",
                   gsub("\\(.*\\)", "", model_family_name), "_",
                   paste(spatiotemporal_list$commonname, 
                         gsub(" ", "-", spatiotemporal_list$name), sep = "_")[i],
                   ".rds")
        )$mod
      })
    
    spatiotemporal_selection_models <- 
      setNames(spatiotemporal_selection_models, 
               paste(spatiotemporal_list$name, 
                     spatiotemporal_list$commonname, sep = "_"))
    
    spatiotemporal_selection_mesh <- 
      pbapply::pblapply(
        unique(dt$commonname), function(k) {
          readRDS(
            dir("data/model_output/model_selection/", 
                pattern = paste0("Spatiotemporal-selection-models-",
                                 gsub("\\(.*\\)", "", model_family_name), "_", k),
                full.names = TRUE)[1])$mesh
        })
  } else {
    spatiotemporal_selection_models <- lapply(tmp, function(k) k$mod)
    spatiotemporal_selection_models <- 
      setNames(spatiotemporal_selection_models, 
               paste(spatiotemporal_list$name, 
                     spatiotemporal_list$commonname, sep = "_"))
    
    spatiotemporal_selection_mesh <- 
      lapply(c(1,5), function(i) tmp[[i]]$mesh)
    
    spatiotemporal_selection_mesh <- 
      setNames(spatiotemporal_selection_mesh, levels(dt$commonname))
    
    if(save_model_output) {
      saveRDS(spatiotemporal_selection_models, 
              file = paste0(
                "data/model_output/model_selection/Spatiotemporal-selection-models-",
                gsub("\\(.*\\)", "", model_family_name), ".rds"),
              compress = "xz"
      ) %>% system.time()
    }
    
    saveRDS(spatiotemporal_selection_mesh,
            file = paste0(
              "data/model_output/model_selection/Spatiotemporal-selection-meshes-",
              gsub("\\(.*\\)", "", model_family_name), ".rds"),
            compress = "xz"
    )  
  }
  
  ## Save required components
  spatiotemporal_selection_params <- 
    list(
      model_names = names(spatiotemporal_selection_models)
    )
  
  saveRDS(
    spatiotemporal_selection_params,
    file = 
      paste0(
        "data/model_output/model_selection/Spatiotemporal-selection-parameters-",
        gsub("\\(.*\\)", "", model_family_name), ".rds")
  )
  
} else {
  spatiotemporal_list <- 
    readRDS(
      file = "data/model_output/model_selection/spatiotemporal_list.rds")
  
  spatiotemporal_selection_mesh <- 
    readRDS(paste0(
      "data/model_output/model_selection/Spatiotemporal-selection-meshes-",
      gsub("\\(.*\\)", "", model_family_name), ".rds")
    )
  
  spatiotemporal_selection_params <-
    readRDS(paste0(
      "data/model_output/model_selection/Spatiotemporal-selection-parameters-",
      gsub("\\(.*\\)", "", model_family_name), ".rds")
    )
}
```  

```{r}
#| label: tbl-spatiotemporal-selection
#| timeit: !expr if(recompile_models & spatiotemporal_selection) TRUE else NULL
#| tbl-cap: !expr paste0("Ranking of spatiotemporal random fields for ", sp_name, " based on ten-fold cross-validation. The models contain spatial fields. Each model is compared to the best fitting single component model ($\\Delta$NLL and $\\Delta$NLL % columns). The negative log-likelihood value is given in the NLL column and AIC difference to the best fitting model in the $\\Delta$AIC column.")

## Table
if(recompile_models & spatiotemporal_selection) {
  spatiotemporal_selection_tab <- 
    pbapply::pblapply(seq_along(spatiotemporal_selection_models), function(i) {
      if(inherits(spatiotemporal_selection_models[[i]], "try-error")) {
        tibble(
          species = case_when(
            spatiotemporal_list$commonname[i] == "snabeluer" ~ "Beaked redfish", 
            spatiotemporal_list$commonname[i] == "vanlig uer" ~ "Golden redfish"
          ),
          model = spatiotemporal_list$name[i],
          formula = paste(trimws(format(spatiotemporal_list$formula[[i]])), 
                          collapse = ""),
          spatial = spatiotemporal_list$spatial[[i]],
          spatiotemporal = spatiotemporal_list$spatiotemporal[[i]],
          crashed = TRUE,
          converged = FALSE,
          sanity = FALSE,
          loglik = NA,
          AIC = NA
        )
      } else {
        tibble(
          species = case_when(
            spatiotemporal_list$commonname[i] == "snabeluer" ~ "Beaked redfish", 
            spatiotemporal_list$commonname[i] == "vanlig uer" ~ "Golden redfish"
          ),
          model = spatiotemporal_list$name[i],
          formula = paste(trimws(
            format(spatiotemporal_selection_models[[i]]$models[[1]]$formula[[1]])), 
            collapse = ""),
          spatial = unique(
            spatiotemporal_selection_models[[i]]$models[[1]]$spatial),
          spatiotemporal = unique(spatiotemporal_selection_models[[i]]$models[[1]]$spatiotemporal),
          crashed = FALSE,
          converged = spatiotemporal_selection_models[[i]]$converged,
          sanity = all(unlist(sanity(
            spatiotemporal_selection_models[[i]]$models[[1]],
            silent = TRUE))),
          loglik = spatiotemporal_selection_models[[i]]$sum_loglik,
          AIC = median(sapply(spatiotemporal_selection_models[[i]]$models, AIC))
        )
      }
    }, cl = n_cores) %>% 
    bind_rows() %>% 
    group_by(species) %>% 
    arrange(-loglik) %>% 
    mutate(
      deltaNLL = loglik - max(loglik, na.rm = TRUE),
      deltaNLL_pr = 100*(1-loglik/max(loglik, na.rm = TRUE)),
      deltaAIC = AIC - min(AIC, na.rm = TRUE),
      deltaAIC_pr = 100*(AIC/min(AIC, na.rm = TRUE) - 1)
    ) %>% 
    ungroup()
  
  ## Save
  write_csv(
    spatiotemporal_selection_tab,
    file =
      paste0("data/model_output/model_selection/Spatiotemporal-selection-table-", 
             gsub("\\(.*\\)", "", model_family_name),
             ".csv")
  )
} else {
  spatiotemporal_selection_tab <- read_csv(
    paste0("data/model_output/model_selection/Spatiotemporal-selection-table-", 
           gsub("\\(.*\\)", "", model_family_name),
           ".csv")
  )
}

## Print the table
spatiotemporal_selection_tab %>% 
  relocate(formula, .after = last_col()) %>%
  dplyr::select(-spatial, -spatiotemporal, -AIC, -deltaAIC_pr) %>% 
  mutate(loglik = abs(loglik), 
         deltaNLL = abs(deltaNLL),
         deltaNLL_pr = abs(deltaNLL_pr)) %>% 
  rename(
    "Species" = "species", "Model" = "model", "Formula" = "formula", 
    "Crashed" = "crashed", "Converged" = "converged", "Sanity test" = "sanity", 
    "NLL" = "loglik", "ΔNLL" = "deltaNLL", "ΔNLL %" =  "deltaNLL_pr", 
    "ΔAIC" = "deltaAIC"
  ) %>% 
  make_table(digits = c(0,0,0,0,0,0,0,1,0,0))

# More compact alternative:
# spatiotemporal_selection_tab %>% 
#   mutate(deltaNLL = abs(deltaNLL)) %>% 
#   dplyr::select(species, model, formula, sanity, deltaNLL) %>%
#   ungroup() %>% 
#   pivot_wider(id_cols = c(model, formula, sanity), 
#               names_from = species,
#               values_from = deltaNLL) %>% 
#   filter(!model %in% "Full model s(logdepth)") %>% 
#   rename("Model" = "model", "Formula" = "formula", 
#          "Sanity test" = "sanity") %>%
#   make_table(digits = 2)
```

```{r}
#| label: fig-spatiotemporal-selection
#| fig-cap: !expr paste("Ranking of spatiotemporal random fields", sp_name, "based on ten-fold cross-validation. The y-axis lists the model components. The x-axis shows the negative log-likelihood for each model, where higher values indicate better model fit. Symbol color and shape reflect model diagnostics; black circles denote models without issues, while grey symbols indicate models that failed the sanity check or did not converge.")

tmp <- spatiotemporal_selection_tab %>%
  group_by(species) %>% 
  mutate(status = case_when(
    !sanity & !converged ~ "Nonconvergent &\nsanity test issues",
    !sanity ~ "Sanity test issues",
    !converged ~ "Nonconvergent",
    .default = "All OK"
  ), .after = "sanity") %>% 
  mutate(deltaNLL = abs(loglik - max(loglik, na.rm = TRUE))) %>% 
  mutate(deltaNLL = abs(ifelse(crashed, Inf, deltaNLL))) %>% 
  rename("value" = "deltaNLL") %>% 
  ungroup()

tmp <- tmp %>% 
  mutate(
    model = factor(
      model,
      levels = tmp %>% group_by(model) %>% reframe(value = mean(value)) %>% 
        arrange(value) %>% pull(model) %>% rev())
  )

ggplot(tmp, aes(x = model, y = value, color = status, shape = status)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c(16,15,17,18)) +
  scale_color_manual(values = c("black", rep("grey", 3))) +
  facet_grid(~species, scales = "free") +
  coord_flip() + 
  # scale_y_continuous(transform = "sqrt") + 
  labs(y = "Negative log likelihood difference",
       x = "Variable", shape = "Model diagnostics", 
       color = "Model diagnostics") +
  theme(legend.position = "bottom",
        panel.grid.major = element_line(color = "grey", linewidth = LS(0.5)))
```

## Survey indices

```{r spat-sel-survey-indices}
#| timeit: !expr if(recalculate_index & spatiotemporal_selection) TRUE else NULL

if(run_index) {
  if(recalculate_index & spatiotemporal_selection) {
    
    proj_dt <- read_rds("data/model_output/prediction-grid-data.rds") %>% 
      distinct(X, Y, depth, year) 
    
    ## Prediction grid cell side length
    grid_side_size <- attributes(proj_dt)$grid_side_size # kilometers due to UTM projection and division by 1000
    
    if(development_mode) {
      proj_dt <- proj_dt %>% 
        filter(year > 2003) %>% 
        mutate(sample = rep(1:4, length.out = nrow(.))) %>% 
        filter(sample == 1) %>% 
        dplyr::select(-sample)
    }
    
    models <- spatiotemporal_selection_models
    
    index <- pbapply::pblapply(seq_along(models), function(i) {
      
      # print(i)
      
      k <- models[[i]]
      if(inherits(k, "try-error")) return(NULL)
      k_name <- names(models)[i]
      
      terms <- trimws(unlist(strsplit(
        as.character(k$models[[1]]$formula[[1]])[3], "\\+")))
      terms <- terms[terms != "0"]
      terms <- terms[!grepl(":", terms)]
      
      pred_dt <- proj_dt
      
      if(grepl("Null model", k_name)) {
        pred_dt <- pred_dt %>% distinct(X, Y, year)
      } else {
        fyear <- any(grepl("fyear", terms))
        geartype <- any(grepl("geartype", terms))
        logdepth <- any(grepl("logdepth", terms))
        cruiseseries <- any(grepl("cruiseseries", terms))
        season <- any(grepl("season", terms))
        depth <- any(grepl("depth", terms))
        month <- any(grepl("month", terms))
        smonth <- any(grepl("s\\(month\\)", terms))
        yday <- any(grepl("yday", terms))
        distance <- any(grepl("distance", terms))
        commonname <- any(grepl("commonname", terms))
        temperature <- any(grepl("temperature", terms))
        
        if(!any(c(depth, temperature))) {
          pred_dt <- pred_dt %>% distinct(X, Y, year)}
        if(logdepth) pred_dt <- pred_dt %>% mutate(logdepth = log(depth))
        if(geartype) pred_dt <- pred_dt %>% mutate(geartype = "Shrimp")
        if(fyear) pred_dt <- pred_dt %>% mutate(fyear = as.factor(year))
        if(season) pred_dt <- pred_dt %>% mutate(season = "H2")
        if(month) {pred_dt <- pred_dt %>% 
          replicate_df("month", if(smonth) as.integer(8) else factor(8))}
        if(yday) pred_dt <- pred_dt %>% mutate(yday = 225)
        if(distance) pred_dt <- pred_dt %>% mutate(distance = 1) # assuming 3 knot trawling speed for 20 min
        
        if(include_temperature & temperature) {
          
          ### Temperature model ###
          load("data/model_output/temperature-model-fit.rda")
          
          ## Predict using the data
          
          pred_dt <- predict(
            fit[[1]]$mod,
            newdata = pred_dt %>% 
              mutate(fyear = as.factor(year),
                     month = as.integer(8)),
            type = "response",
            model = NA,
            return_tmb_object = FALSE) %>% 
            dplyr::select(-est_non_rf, -est_rf, -omega_s, -epsilon_st) %>% 
            rename("temperature" = "est") 
          
          if(!smonth) {
            pred_dt <- pred_dt %>% mutate(month = factor(month))
          }
        }
        
      }
      
      ## Slower more official way with CIs
      # lapply(seq_along(k$models), function(j) {
      #   predict(k$models[[j]], newdata = proj_dt, type = "response", model = NA,
      #           return_tmb_object = TRUE) %>% 
      #     get_index(
      #       bias_correct = FALSE,
      #       area = (grid_side_size/1.852)^2) %>% # grid cell area in square nautical miles
      #     mutate(est = est/1e6, # convert kg to kilotonnes
      #            lwr = lwr/1e6,
      #            upr = upr/1e6, 
      #            cv = sqrt(exp(se^2) - 1)) %>% 
      #     mutate(model = k_name, .before = 1) %>% 
      #     mutate(fold = j)
      # }) %>% 
      #   bind_rows()
      
      ## Quicker way
      mod_pred <- 
        lapply(seq_along(k$models), function(j) {
          predict(k$models[[j]], newdata = pred_dt, type = "response", 
                  model = NA,
                  return_tmb_object = FALSE) %>%
            mutate(fold = j)
        })
      
      mod_pred %>%
        bind_rows() %>%
        group_by(fold, year, X, Y) %>%
        reframe(est = mean(est)) %>%
        group_by(fold, year) %>%
        reframe(
          # Y = weighted.mean(Y, est),
          # X = weighted.mean(X, est),
          # depth = weighted.mean(depth, est),
          est = (sum(est)*(grid_side_size/1.852)^2)/1e6
        ) %>%
        mutate(lwr = NA, upr = NA, cv = NA, se = NA) %>%
        mutate(commonname = unique(k$data$commonname),
               model = k_name, 
               .before = 1)
    }, cl = n_cores)
    
    names(index) <- names(models)
    
    write_rds(
      index,
      file =
        paste0(
          "data/model_output/model_selection/Spatiotemporal-selection-indices-", 
               gsub("\\(.*\\)", "", model_family_name),
               ".rds"), 
      compress = "xz"
    )
    
    ss_indices <- index
    
  } else {
    
    proj_dt <- read_rds("data/model_output/prediction-grid-data.rds") %>% 
      distinct(X, Y, depth, year)
    
    ss_indices <- read_rds(
      paste0(
        "data/model_output/model_selection/Spatiotemporal-selection-indices-", 
             gsub("\\(.*\\)", "", model_family_name),
             ".rds"), 
    ) 
  }
}
```

```{r}
#| code-fold: true
#| label: fig-spat-sel-index
#| fig-height: !expr ceiling(length(ss_indices)/3)*2.5
#| fig-cap: !expr paste0("Survey indices estimated using different spatiotemporal random field structures ", sp_name, " using spatiotemporal generalized linear mixed effects models with the sdmTMB package (colored solid line). Black lines with shading indicate the full biomass models with multiple components; 0 + geartype + s(month) + s(depth) + s(temperature). ", value_text_name, " is used as the response variable, and the models use meshes with ", paste(sapply(spatiotemporal_selection_mesh, function(k) k$mesh$n), collapse = " and "), " vertices (knots), as well as ", model_family_name," model family. Predictions are made to a grid with ", proj_dt %>% distinct(X,Y) %>% nrow(), " cells with a side of ", attributes(proj_dt)$grid_side_size, " km. Each line represents one fold from the 10-fold cross-validation. See Table 4 for more detailed model specification.")

if(run_index) {
  tmp <- lapply(ss_indices, function(k) {
    if(inherits(k, "try-error")) return(NULL)
    if(any(!is.na(k$upr))) {
      if(max(k$upr, na.rm = TRUE) > 2*max(k$est, na.rm = TRUE)) {
        k[k$upr > 1.5*max(k$est, na.rm = TRUE),]$upr <- 
          ceiling(1.5*max(k$est, na.rm = TRUE))
      }}
    k
  }) %>%
    bind_rows() %>% 
    filter(!is.na(year)) %>% 
    mutate(model = gsub("_snabeluer|_vanlig uer", "", model)) %>% 
    # mutate(model = factor(model, levels = names(ss_indices))) %>% 
    mutate(
      species = case_when(
        commonname == "snabeluer" ~ "Beaked redfish", 
        commonname == "vanlig uer" ~ "Golden redfish"
      ),
      .after = "commonname")
  
  load("data/model_output/index-bm.rda")
  
  index <- index %>% 
    bind_rows() %>% 
    rename("species" = "length_group")
  
  cowplot::plot_grid(
    plotlist = 
      lapply(unique(tmp$species), function(k) {
        ggplot() +
          geom_ribbon(
            data = index %>% filter(species == k),
            aes(year, est, ymin = lwr, ymax = upr), alpha = 0.3, color = NA, 
            fill = "grey") +
          geom_path(
            data = index %>% filter(species == k),
            aes(year, est), alpha = 0.5) +
          geom_path(
            data = tmp %>% filter(species == k),
            aes(year, est, color = model, group = fold), alpha = 0.5) +
          facet_wrap(~model, scales = "free") + 
          coord_cartesian(expand = FALSE, ylim = c(0, NA)) +
          scale_x_continuous(
            breaks = seq(1980, 2030, 4),
            labels = substr(seq(1980, 2030, 4),3,4)
          ) +
          labs(
            x = 'Year',
            y = ifelse(model_type == "Biomass", 'Biomass (kt)',
                       'Abundance (millions)')
          ) +
          theme_bw() +
          theme(legend.position = "none")
      }),
    labels = unique(tmp$species), ncol = 1
  )
}
```

