---
title: "SI Text 4: length based models for redfish"
subtitle: "Rho abundance; delta_lognormal"
# author:
#    - name: Mikko Vihtakari
#      affiliations:
#       - ref: imr
# affiliations: 
#    - id: imr 
#      name: Institute of Marine Research, Norway
date: today
published-title: "Version"
bibliography: "../manuscript/zotero_library.bib"
csl: "../manuscript/ices_jms.csl"
format:
  html:
    self-contained: true
    fig-width: 9
    fig-height: 4
    code-fold: true
    output-file: SI-Text-4-length-based-models
    toc: true
execute:
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

```{r setup, eval = TRUE, include = FALSE}

knitr::opts_knit$set(root.dir = '../../') 

knitr::knit_hooks$set(timeit = local({
  now = NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res = round(difftime(Sys.time(), now, units = "mins"), 1)
      now <<- NULL
      # use options$label if you want the chunk label as well
      print(paste('Time:', as.character(res), "min"))
    }
  }})
)

local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

```{r packages}
# Clear the workspace for manual code runs

rm(list = ls()) 

############## ####
#### Libraries ----

## Package names
packages <- c("tidyverse", "sf", "ggOceanMaps", "ggspatial", "sdmTMB", 
              "ggeffects", "scales", "concaveman", "plotly", "data.table",
              "dtplyr", "remotes", "sdmTMBextra", "lubridate", "inlabru")

## Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())

if("remotes" %in% packages[!installed_packages]) {
  install.packages("remotes")
}

if (any(installed_packages == FALSE)) {
  
  if("sdmTMBextra" %in% packages[!installed_packages]) {
    remotes::install_github("pbs-assess/sdmTMBextra", dependencies = TRUE,
                            upgrade = "never")
  }
  
  installed_packages <- packages %in% rownames(installed.packages())
  
  install.packages(packages[!installed_packages])
}

## Load packages
invisible(lapply(packages, function(x) {
  suppressPackageStartupMessages(library(x, character.only = TRUE))}))

# Backend for parallel

if(Sys.info()["sysname"] == "Linux") {
  library(flexiblas)
  flexiblas_switch(flexiblas_load_backend("OPENBLAS-THREADS"))
}

################### ###
# Custom functions ----

source("R/legacy/make_table.R") # a function to make tables
source("R/legacy/bubble_map.R")

detect_survey <- function(y) {
  dplyr::case_when(
    length(unique(y)) > 5 ~ "all-data",
    all(c("EggaN", "EggaS", "WinterS", "EcoS", "CoastalS") %in% y) ~ "all-surveys",
    all(c("EggaN", "EggaS") %in% y) ~ "egga-surveys",
    "CoastalS" %in% y ~ "coastal-survey",
    "EggaN" %in% y ~ "eggan-survey",
    "EggaS" %in% y ~ "eggas-survey",
    "WinterS" %in% y ~ "winter-survey",
    "EcoS" %in% y ~ "ecosystem-survey"
  )
}

#' @title Return the same column types for prediction dataset as used to formulate sdmTMB model
#' @param mod_data Data from an sdmTMB model, typically fit$data, where fit is the sdmTMB object
#' @param pred_data Prediction dataset whose column types should be converted
#' @return Returns \code{pred_data} with matching common column types to \code{mod_data}.
match_col_types <- function(mod_data, pred_data) {
  
  common_cols <- intersect(colnames(mod_data), colnames(pred_data))
  
  mod_classes <- sapply(mod_data[,common_cols], class)
  pred_classes <- sapply(pred_data[,common_cols], class)
  
  common_cols <- common_cols[mod_classes != pred_classes]
  
  pred_data[common_cols] <- lapply(common_cols, function(x) {
    class_fun <- match.fun(paste0("as.", class(mod_data[[x]])))
    if(inherits(pred_data[[x]], "factor")) {
      class_fun(as.character(pred_data[[x]]))
    } else {
      class_fun(pred_data[[x]])
    }
  })
  
  pred_data
}

#' @title List values in text
list_values <- function(x) {
  if(length(x) == 1) {
    paste(x)
  } else if(length(x) == 2) {
    paste(x, collapse = " and ")
  } else (
    paste0(paste(x[1:(length(x)-1)], collapse = ", "), ", and ", x[length(x)])
  )
}


################ ##
# ggplot theme ----

theme_cust <- theme_classic(base_size = 10) %+replace%
  theme(strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank(),
        legend.background = element_blank(),
        legend.box.background = element_blank(),
        plot.margin = margin(5.5, 10, 5.5, 5.5))

theme_set(theme_cust) # Set default theme globally for the entire project
```

<details>
<summary>Unfold to see R package versions and system information used to generate the document</summary>
```{r}
#| label: sys-info
sessionInfo()
```
</details>

Supplementary information file for @vihtakari providing reproducible documentation on calculation of the length-based survey indices using generalized linear mixed effects models fitted with the [sdmTMB package](https://CRAN.R-project.org/package=sdmTMB) [@sdmTMB] for R [@rcoreteam]. The document is written using [Quarto](https://quarto.org/) and can be compiled within the [GitHub repository](https://github.com/DeepWaterIMR/ref-index-paper) for the project. HTML version of this document contains foldable code fields showing the code used to compile the model. 

# Data

The dataset has been described in the main text and in SI Text 1. Here we read and compile data to a format required by sdmTMB:

```{r data}
# Settings

recompile_model <- FALSE # switch to compile the model. If FALSE, the model is loaded from the data/model_output folder.
recalculate_fixed_effects <- FALSE # switch to calculate fixed effects. If FALSE, loaded from a file.
recalculate_index <- FALSE # switch to recalculate the survey indices

overview_plots <- TRUE # switch to include overview plots. Increases file size
run_fixef <- TRUE # switch to make fixed effect plots. 
run_regional <- TRUE # switch to compute regional indices.

bias_correct_index <- TRUE # switch to turn index bias-correction off. TMB commonly crashes and takes long time. Set to FALSE to avoid memory related crashes. Should be on in the final version. 

length_based <- TRUE # switch between station-level (FALSE) and length-based (TRUE) data. Legacy. Do not change.
model_name <- "Length-based"

remove_zero_catch_surveys <- TRUE # switch to remove surveys with zero catch of a given species. Should be TRUE unless testing. 

marginal_effects <- FALSE # switch to plot marginal fixed effects through ggeffect (quicker) instead of conditional effects (slower). Marginal effects do not work for hurdle/delta models. 

## Response variable
response_variable <- "abundance density" # options: "catch biomass" meaning trawl catch weight of the target species (biomass column in dt). Log transformed swept area by each trawl will be used as offset argument. "biomass density" referring to the rho_biomass column giving readily calculated biomass density (biomass/swept area). "catch abundance" and "abundance density" for similar variables using abundance/count data. Note that the width of the trawl influences the relative difference between gear types, which is included to the model, and only affect the absolute level of the index estimate in the end. 

## Model family (use this to set the family)
model_family <- delta_lognormal()
model_family_name <- ifelse(is.null(model_family$clean_name), model_family$family, model_family$clean_name)

### Link function inversion function
if(length(model_family$family) > 1) {
  linkinv <- model_family[[2]]$linkinv
} else {
  lininv <- model_family$linkinv
}

## Threshold for automatic outlier removal
outlier_threshold <- NA # 0.9999 # Proportion (as used in the quantile() function). Set to NA to disable outlier removal. 

## Method for calculating Q-Q plots
residual_method <- "mle-mvn" # "mle-mvn" for fixed effect at their Maximum Likelihood Estimates assuming multivariate normal distribution (quick); "mle-mcmc" for fixed effects at their MLEs random effects with MCMC (very slow); "simulate" for simulation-based randomized quantile residuals (semi-quick). See https://pbs-assess.github.io/sdmTMB/articles/residual-checking.html

## Mesh cutoff argument (smaller values create meshes with more knots)
mesh_cutoff <- 100 # Minimum distance between knots before a new mesh vertex is added in kilometers
max_data_depth <- 1400 # Max depth for data in meters to be used in figures

## Prediction grid cell side length
grid_side_size <- 20 # kilometers due to UTM projection and division by 1000
max_prediction_depth <- 1400 # Max depth for predictions

####################
# Read the data ----

load("data/out/Compiled_redfish_survey_data.rda")

if(!length_based) dt <- stn 

dt <- lapply(c("snabeluer", "vanlig uer"), function(filter_sp) {
  out <- dt %>%
    distinct() %>% 
    filter(commonname %in% filter_sp) %>% 
    rename(lon = longitudestart,
           lat = latitudestart,
           year = startyear,
           depth = bottomdepthstart) %>% 
    mutate(gear_width = 25) %>% 
    mutate(cruiseseries = factor(cruiseseries),
           fyear = factor(year),
           id = paste(cruiseseries, year, platformname, serialnumber, sep = "-"),
           swept_area = (gear_width/1852)*distance) %>% 
    mutate(yday = lubridate::yday(stationstartdate),
           .after = season) %>% 
    # mutate(month = as.integer(month)) %>% 
    mutate(month = as.factor(month)) %>% 
    droplevels() %>% 
    mutate(
      rho_biomass = if(!length_based) biomass/swept_area else weight/swept_area,
      rho_abundance = if(!length_based) count/swept_area else n/swept_area,
      value = rho_biomass) %>% # this defines the response type
    filter(!is.na(value))
  
  ## Zero catch survey removal
  if(remove_zero_catch_surveys) {
    tmp <- out %>% 
      group_by(cruiseseries, cruise, year, platformname) %>% 
      reframe(sum = sum(value)) %>% 
      filter(sum == 0) %>% 
      mutate(tmpid = paste(cruise, year, platformname))
    
    out <- out %>% 
      mutate(tmpid = paste(cruise, year, platformname)) %>% 
      filter(!tmpid %in% tmp$tmpid) %>% 
      dplyr::select(-tmpid)
  }
  
  ## Length groups
  if(length_based) {
    if(filter_sp == "snabeluer") {
      
      out <- out %>% 
        mutate(
          length_group = 
            recode(
              length_group, 
              "(0,5]" = "(0,10]", "(5,10]" = "(0,10]", "(40,45]" = "(40,Inf]", 
              "(45,50]" = "(40,Inf]", "(50,55]" = "(40,Inf]", "(55,60]" = "(40,Inf]",
              "(60,65]" = "(40,Inf]", "(65,70]" = "(40,Inf]", "(70,75]" = "(40,Inf]",
              "(75,80]" = "(40,Inf]", "(80,85]" = "(40,Inf]", "(85,90]" = "(40,Inf]",
              "(90,95]" = "(40,Inf]"
            )
        ) %>% 
        group_by(across(c(-n, -weight, -rho_biomass, -rho_abundance, -value))) %>% 
        reframe(
          n = sum(n, na.rm = TRUE), weight = sum(weight, na.rm = TRUE), 
          rho_abundance = sum(rho_abundance, na.rm = TRUE), 
          rho_biomass = sum(rho_biomass, na.rm = TRUE), 
          value = sum(value, na.rm = TRUE))
      
    } else if(filter_sp == "vanlig uer") {
      
      out <- out %>% 
        mutate(
          length_group = 
            recode(
              length_group, 
              "(0,5]" = "(0,15]", "(5,10]" = "(0,15]", "(10,15]" = "(0,15]",
              "(40,45]" = "(40,55]", "(45,50]" = "(40,55]", "(50,55]" = "(40,55]", 
              "(55,60]" = "(55,Inf]", "(60,65]" = "(55,Inf]", "(65,70]" = "(55,Inf]",
              "(70,75]" = "(55,Inf]", "(75,80]" = "(55,Inf]", "(80,85]" = "(55,Inf]",
              "(85,90]" = "(55,Inf]", "(90,95]" = "(55,Inf]"
            )
        ) %>% 
        group_by(across(c(-n, -weight, -rho_biomass, -rho_abundance, -value))) %>% 
        reframe(
          n = sum(n, na.rm = TRUE), weight = sum(weight, na.rm = TRUE), 
          rho_abundance = sum(rho_abundance, na.rm = TRUE), 
          rho_biomass = sum(rho_biomass, na.rm = TRUE), 
          value = sum(value, na.rm = TRUE))
    }
  }
  
  ## Outlier removal
  
  if(!is.na(outlier_threshold)) {
    out <- out %>% 
      {if(length_based) group_by(length_group) else .} %>% 
      mutate(outlier = value > quantile(value, outlier_threshold), .after = "value")
  } else {
    out <- out %>% mutate(outlier = FALSE, .after = "value")
  }
  
  return(out)
}) %>% 
  bind_rows()

if(length_based) {
  dt <- dt %>% 
    mutate(
      length_group = factor(
        length_group, levels = sort(as.character(unique(dt$length_group))))
    )
}

## CRS

dt_crs <- get_crs(dt, ll_names = c("lon", "lat"))
dt <- sdmTMB::add_utm_columns(dt, ll_names = c("lon", "lat"), utm_crs = dt_crs)

## Response variable

if(response_variable == "biomass density") {
  dt <- dt %>% mutate(value = rho_biomass, .after = rho_abundance) %>% 
    filter(!is.na(value))
} else if(response_variable == "abundance density") {
  dt <- dt %>% mutate(value = rho_abundance, .after = rho_abundance) %>% 
    filter(!is.na(value))
} else if(response_variable == "catch biomass") {
  dt <- dt %>% mutate(value = biomass, .after = rho_abundance) %>% 
    filter(!is.na(value))
} else if(response_variable == "catch abundance") {
  dt <- dt %>% mutate(value = count, .after = rho_abundance) %>% 
    filter(!is.na(value))
} else {
  stop("The response_variable object has to be 'biomass density', 'abundance density', 'catch biomass', or 'catch abundance'")
}

# Response variable type definitions for automatic text

model_type <- ifelse(response_variable %in% c("biomass density", "catch biomass"), "Biomass", "Abundance")
offset <- response_variable %in% c("catch biomass", "catch count")

## Response variables and units for figure labels
if(response_variable == "biomass density") {
  value_name <- expression(paste("Density (kg ", nmi^-2, ")"))
} else if(response_variable == "catch biomass") {
  value_name <- "Catch weight (kg)"
} else if(response_variable == "abundance density") {
  value_name <- expression(paste("Density (n ", nmi^-2, ")"))
} else {
  value_name <- "Catch count (n)"
}

if(response_variable == "biomass density") {
  value_unit <- quote(kg~nmi^-2)
} else if(response_variable == "catch biomass") {
  value_unit <- "kg"
} else if(response_variable == "abundance density") {
  value_unit <- quote(n~nmi^-2)
} else {
  value_unit <- "n"
}

## Response variables and units for text
if(response_variable == "biomass density") {
  value_text_name <- "Swept area biomass density (kg nmi$^{-2}$)"
} else if(response_variable == "catch biomass") {
  value_text_name <- "Catch weight (kg)"
} else if(response_variable == "abundance density") {
  value_text_name <- "Swept area abundance density (n nmi$^{-2}$)"
} else {
  value_text_name <- "Catch count (n)"
}

if(response_variable == "biomass density") {
  value_text_unit <- "kg nmi$^{-2}$"
} else if(response_variable == "catch biomass") {
  value_text_unit <- "kg"
} else if(response_variable == "abundance density") {
  value_text_unit <- "n nmi$^{-2}$"
} else {
  value_text_unit <- "n"
}

map_ncol <- 8
survey <- detect_survey(unique(dt$cruiseseries))

## Species name, legacy; should be "both"

filter_sp <- "both"
sp_name <- case_when(filter_sp == "snabeluer" ~ "beaked redfish", 
                     filter_sp == "vanlig uer" ~ "golden redfish", 
                     filter_sp == "both" ~ "beaked and golden redfish")
Sp_name <- case_when(filter_sp == "snabeluer" ~ "Beaked redfish", 
                     filter_sp == "vanlig uer" ~ "Golden redfish", 
                     filter_sp == "both" ~ "Beaked and golden redfish")

dt <- dt %>% 
  mutate(
    species = case_when(
      commonname == "snabeluer" ~ "Beaked redfish", 
      commonname == "vanlig uer" ~ "Golden redfish"
    ),
    .after = "commonname")
```

The cleaned dataset comprised `r dt %>% filter(!outlier) %>% distinct(year, platformname, serialnumber) %>% nrow()` bottom trawl hauls collected over `r dt %>% filter(!outlier) %>% distinct(year) %>% nrow()` years, from `r min(dt$year)` to `r max(dt$year)`, during `r dt %>% distinct(year, cruise, platformname) %>% nrow()` surveys. Evaluated from `r ifelse(model_type == "Biomass", "weight", "count")` information, `r dt %>% filter(value > 0, !is.na(value), !outlier) %>% distinct(year, platformname, serialnumber) %>% nrow()` of these stations contained `r sp_name`, and `r dt %>% filter(value == 0, !is.na(value), !outlier) %>% distinct(year, platformname, serialnumber) %>% nrow()` stations did not contain the species. The analysis `r ifelse(length_based, paste("was allocated to", length(unique(dt$length_group)), "length groups for", length(unique(dt$commonname)), "species"), "was not allocated to length groups")`. 

`r value_text_name` was used as the response variable. `r ifelse(offset, "Logarithm transformed swept area (distance times arbitrary trawl width) was used as the offset argument in the sdmTMB model.", "")`

## Outlier removal

`r ifelse(is.na(outlier_threshold), "Outlier threshold was set to NA meaning that no automatic outlier removal was conducted", paste0("Outlier threshold was set to ", outlier_threshold, " sample quantile probability leading to removal of ", sum(dt$outlier), " values above ", round(quantile(dt$value, outlier_threshold), 0), " ", value_text_unit))`. `r ifelse(!is.na(outlier_threshold), paste0("These outliers had a catch weight between ", dt %>% filter(outlier) %>% pull(biomass) %>% min() %>% round(), " and ", dt %>% filter(outlier) %>% pull(biomass) %>% max() %>% round(), " kg ", sp_name, ", and trawling distance between ", dt %>% filter(outlier) %>% pull(distance) %>% min() %>% round(1), " and ", dt %>% filter(outlier) %>% pull(distance) %>% max() %>% round(1), " nmi."), "")`

```{r}
#| label: fig-outlier
#| fig-height: !expr ifelse(length_based, 8, 4)
#| fig-cap: !expr paste("Logarithmic distribution of the response variable values", ifelse(length_based, "allocated to length groups.", "."), ifelse(!is.na(outlier_threshold), "Red bars indicate values removed from the analyses as outliers.", "No values were removed as outliers."))

if(overview_plots) {
  if(length_based) {
    ggplot(dt, aes(value, fill = outlier)) +
      geom_vline(xintercept = quantile(dt$value, outlier_threshold), 
                 color = "grey") +
      geom_histogram() +
      scale_y_continuous(
        trans = "log1p",
        labels = scales::label_log(), 
        breaks = 10^seq(-1,6),
        expand = c(0,0)) +
      scale_x_continuous(
        trans = "log1p", 
        labels = scales::label_log(),
        breaks = c(0,1, 10, 100, 1e3, 1e4, 1e5, 1e6, 1e7),
        expand = c(0,0)
      ) +
      coord_cartesian(xlim = c(NA, 10^ceiling(max(log10(dt$value))))) +
      labs(y = "Count", x = value_name, fill = paste0("Outlier\n(", outlier_threshold*100, " percentile)")) +
      scale_fill_manual(values = c(`TRUE` = "#832424", `FALSE` = "grey50")) +
      facet_wrap(~species+length_group)
  } else {
    ggplot(dt, aes(value, fill = outlier)) +
      geom_vline(xintercept = quantile(dt$value, outlier_threshold), 
                 color = "grey") +
      geom_histogram() +
      scale_y_continuous(
        trans = "log1p",
        labels = scales::label_log(), 
        breaks = 10^seq(-1,6),
        expand = c(0,0)) +
      scale_x_continuous(
        trans = "log1p", 
        labels = scales::label_log(),
        breaks = c(0,1, 10, 100, 1e3, 1e4, 1e5, 1e6, 1e7),
        expand = c(0,0)
      ) +
      coord_cartesian(xlim = c(NA, 10^ceiling(max(log10(dt$value))))) +
      labs(y = "Count", x = value_name, fill = paste0("Outlier\n(", outlier_threshold*100, " percentile)")) +
      scale_fill_manual(values = c(`TRUE` = "#832424", `FALSE` = "grey50")) +
      facet_wrap(~species)
  }
}
```

## Annual averages

```{r}
#| label: fig-n-stn
#| fig-height: !expr ifelse(length_based, 8, 4)
#| fig-cap: !expr paste("Number of stations with and without", sp_name, "by year", ifelse(length_based, "allocated to length groups.", "."))

if(overview_plots) {
  if(!length_based) {
    dt %>%
      filter(!outlier) %>% 
      group_by(year, species) %>%
      reframe(yes = sum(value > 0, na.rm = TRUE),
              no = sum(value == 0, na.rm = TRUE)) %>%
      pivot_longer(cols = c(yes, no)) %>%
      mutate(name = factor(name, c("no", "yes"))) %>%
      ggplot(aes(x = year, y = value, fill = name)) +
      geom_col() +
      labs(x = "Year", y = "Number of stations", fill = "Contains\ntarget taxa") +
      scale_fill_manual(values = c("yes" = "#66a182", "no" = "grey")) +
      scale_y_continuous(expand = c(0.01,0)) +
      scale_x_continuous(breaks = seq(1980,2030,5),
                         labels = substr(seq(1980,2030,5),3,4),
                         expand = c(0.01,0)) +
      theme(legend.position = "bottom") +
      facet_wrap(~species)
  } else {
    dt %>%
      filter(!outlier) %>% 
      group_by(year, species, length_group) %>%
      reframe(yes = sum(value > 0, na.rm = TRUE),
              no = sum(value == 0, na.rm = TRUE)) %>%
      pivot_longer(cols = c(yes, no)) %>%
      mutate(name = factor(name, c("no", "yes"))) %>%
      ggplot(aes(x = year, y = value, fill = name)) +
      geom_col() +
      labs(x = "Year", y = "Number of stations", fill = "Contains\ntarget taxa") +
      facet_wrap(~species + length_group) +
      scale_fill_manual(values = c("yes" = "#66a182", "no" = "grey")) +
      scale_y_continuous(expand = c(0.01,0)) +
      scale_x_continuous(breaks = seq(1980,2030,5),
                         labels = substr(seq(1980,2030,5),3,4),
                         expand = c(0.01,0)) +
      theme(legend.position = "bottom")
  }
}
```

```{r}
#| label: fig-year-dist
#| fig-height: !expr ifelse(length_based, 8, 4)
#| fig-cap: !expr paste0(trimws(gsub("\\(.*\\)", "", value_text_name)), " of each trawl haul (grey dots) by year", ifelse(length_based, " allocated to length groups.", "."), " Crossbars indicate mean values and 95% confidence intervals. Blue cross bars are for all data, red for only positive data. The y-axis is log + 1 transformed.", ifelse(!is.na(outlier_threshold), " Dark red dots indicate trawl hauls that were deemed outliers by the set criterion and removed from further analyses.", "")) 

if(overview_plots) {
  dt %>%
    ggplot(aes(x = year, y = value)) +
    geom_jitter(aes(color = outlier), size = 0.3) +
    stat_summary(
      data = dt %>% filter(value > 0, !outlier),
      fun.data = mean_cl_normal, geom = "crossbar", width = .5,
      color = "red") +
    stat_summary(
      data = dt %>% filter(!outlier),
      fun.data = mean_cl_normal, geom = "crossbar", width = .5,
      color = "blue") +
    labs(x = "Year", y = value_name) +
    {if(length_based) facet_wrap(~species + length_group)} +
    {if(!length_based) facet_wrap(~species)} +
    scale_color_manual(values = c(`TRUE` = "#832424", `FALSE` = "grey70")) +
    scale_x_continuous(
      breaks = dt %>% pull(year) %>% unique,
      labels = substr(dt %>% pull(year) %>% unique, 3,4),
      expand = c(0.01,0)
    ) +
    scale_y_continuous(
      trans = "log1p",
      breaks = c(0,1,10,100,1e3,5e3,1e4,2.5e4,5e4,1e5,2e5,3e5,5e5,1e6),
      expand = c(0.01,0)
    ) +
    # coord_trans(y = "log1p") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          legend.position = "none")
}
```

```{r}
#| label: fig-lg-year
#| fig-height: 10
#| fig-cap: !expr paste0("Average ", tolower(trimws(gsub("\\(.*\\)", "", value_text_name))), " of ", sp_name, " (+ 95% confidence intervals) by year, species and length group.")

if(overview_plots) {
  if(length_based) {
    dt %>%
      droplevels() %>% 
      group_by(year, species, length_group) %>%
      reframe(
        mean = mean(value, na.rm = TRUE),
        se = sd(value, na.rm = TRUE)/sqrt(length(!is.na(value)))
      ) %>%
      ggplot(aes(x = length_group, y = mean, ymin = mean, ymax = mean + 1.96*se)) +
      geom_col(color = "black", linewidth = 0.3, position = "dodge",
               fill = "grey") +
      geom_errorbar(width = 0.2, linewidth = 0.3) +
      facet_wrap(~species + year, ncol = map_ncol,  dir = "v", scales = "free_y") +
      labs(y = value_name,
           x = "Length group (cm)") +
      scale_x_discrete(guide = guide_axis(angle = 90)) +
      coord_cartesian(expand = FALSE) +
      theme_classic(base_size = 8) + 
      theme(
        legend.position = "bottom",
        strip.background = element_blank(),
        strip.text = element_text(size = 10, face = 2)
      )
  }
}
```

## Depth dependency

```{r}
#| label: fig-depth-dist
#| fig-cap: !expr paste0("Bottom depth related to log+1 transformed ", trimws(tolower(gsub("\\(.*\\)", "", value_text_name))), " of each trawl haul (grey dots)", ifelse(length_based, "allocated to length groups.", "."), " The blue and red line and shading represent loess smoother and standard error for all data and only positive values, respectively. The y-axis is log + 1 transformed.")
#| fig-height: !expr ifelse(length_based, 8, 5)

if(overview_plots) {
  ggplot(data = dt %>% filter(!outlier), aes(x = depth, y = value)) +
    geom_point(color = "grey", size = 0.3) +
    geom_smooth(fill = "blue", color = "blue") +
    geom_smooth(data = dt %>% filter(value > 0, !outlier), 
                fill = "red", color = "red") +
    {if(length_based) facet_wrap(~species + length_group)} +
    {if(!length_based) facet_wrap(~species)} +
    labs(x = "Bottom depth (m)", y = value_name) +
    scale_y_continuous(
      trans = "log1p",
      breaks = c(0,1,10,100,1e3,5e3,1e4,2.5e4,5e4,1e5,2e5,3e5,5e5,1e6),
      expand = c(0.01,0)) +
    coord_cartesian(ylim = c(0,NA)) +
    theme_bw(base_size = 10) +
    scale_x_continuous(breaks = seq(0,2000,200),
                       expand = c(0.01,0))
}
```

## Seasonal dependency 

```{r}
#| label: fig-seasonal-dist
#| fig-cap: !expr paste0("Ordinal date related to log+1 transformed ", trimws(tolower(gsub("\\(.*\\)", "", value_text_name))), " of each trawl haul (grey dots)", ifelse(length_based, "allocated to length groups.", "."), " The blue and red line and shading represent loess smoother and standard error for all data and only positive values, respectively. The y-axis is log + 1 transformed.")
#| fig-height: !expr ifelse(length_based, 8, 5)

if(overview_plots) {
  ggplot(data = dt %>% filter(!outlier), 
         aes(x = as.Date(yday, "2023-01-01"), y = value)) +
    geom_point(color = "grey", size = 0.3) +
    geom_smooth(fill = "blue", color = "blue") +
    geom_smooth(data = dt %>% filter(value > 0, !outlier), 
                fill = "red", color = "red") +
    {if(length_based) facet_wrap(~species + length_group)} +
    {if(!length_based) facet_wrap(~species)} +
    labs(x = "Month", y = value_name) +
    scale_y_continuous(
      trans = "log1p",
      breaks = c(0,1,10,100,1e3,5e3,1e4,2.5e4,5e4,1e5,2e5,3e5,5e5,1e6),
      expand = c(0.01,0)) +
    coord_cartesian(ylim = c(0,NA), 
                    xlim = as.Date(c("2023-01-01", "2023-12-31"))) +
    theme_bw(base_size = 10) +
    scale_x_date(date_labels = "%b", 
                 date_breaks = "1 months",
                 expand = c(0,0))
}
```

## Spatial sampling distribution

```{r}
#| label: fig-sampling-distribution
#| fig-height: 9
#| fig-cap: Sampling station distribution (grey dots) by year together with density contours showing the regions with most dense sampling.

if(overview_plots) {
  tmp <- dt %>% 
    filter(!outlier) %>% 
    distinct(year, cruiseseries, cruise, platformname, serialnumber,
             lon, lat, X, Y) %>% 
    mutate(X = X*1000, Y = Y*1000)
  
  basemap(tmp, crs = dt_crs) +
    geom_point(data = tmp, aes(x = X, y = Y), size = 0.2, alpha = 0.3) +
    geom_density_2d(data = tmp, aes(x = X, y = Y)) + 
    facet_wrap(~year, ncol = map_ncol) +
    theme_cust +
    theme(axis.title = element_blank())
}
```

```{r}
#| label: fig-cog-sampling
#| fig-height: 6
#| fig-cap: Centers of sampling gravity (average longitude and latitude) by year. Labels indicate the last digits for year. Color is scaled to year with older estimates being blue and newer estimates yellow.

if(overview_plots) {
  tmp2 <- tmp %>% 
    group_by(year) %>% 
    reframe(lon = mean(lon), lat = mean(lat), X = mean(X),Y  = mean(Y))
  
  basemap(tmp2, crs = dt_crs, bathy.style = "rcg", legends = FALSE, base_size = 8,
          expand.factor = 2) +
    ggspatial::geom_spatial_text(
      data = tmp2,
      aes(X, Y, label = substr(year,3,4), color=year),
      fontface = 2,
      crs = dt_crs
    ) +
    scale_color_viridis_c() + 
    theme(legend.position = "none")
}
```

## Distribution maps

```{r}
#| label: fig-dist-map
#| fig-cap: !expr if(length_based) paste0("Distribution of ", sp_name, " length group ", tolower(model_type), " density scaled to symbol color and size.") else paste0("Annual distribution of ", sp_name, " ", tolower(model_type), " density scaled to symbol color and size.")
#| fig-height: 12

if(overview_plots) {
  if(length_based) {
    
    length_groups <- dt %>%
      filter(!outlier) %>% 
      filter(value > 0) %>%
      group_by(length_group) %>%
      count() %>%
      filter(n > 10) %>%
      pull(length_group) %>%
      droplevels() %>%
      levels()
    
    dt %>% 
      filter(length_group %in% length_groups) %>% 
      droplevels() %>% 
      bubble_map(label = value_name,  by_year = FALSE) +
      facet_wrap(~species + length_group, ncol = 4)
    
    # dt %>%
    #   filter(!outlier) %>% 
    #   filter(length_group %in% length_groups) %>%
    #   filter(!is.na(value)) %>%
    #   droplevels() %>%
    #   mutate(value = value) %>%
    #   bubble_map(label = value_name, by_year = FALSE) +
    #   facet_grid(year~length_group)
    
  } else {
    dt %>%
      filter(!outlier) %>% 
      bubble_map(label = value_name, ncol = map_ncol) +
      facet_wrap(~species)
  }
}
```

# Model fitting

```{r fit}
#| code-fold: true
#| warning: true
#| message: true
#| timeit: !expr if(recompile_model) TRUE else NULL

length_groups <- dt %>%
  filter(!outlier) %>% 
  filter(value > 0) %>%
  group_by(length_group, commonname) %>%
  count() %>%
  filter(n > 10) %>%
  mutate(lg = paste(commonname, length_group)) %>% 
  pull(lg) %>%
  sort()

if(recompile_model) {
  
  if(!dir.exists("data/model_output/length_based/")) {
    dir.create("data/model_output/length_based/")
  }
  
  pbapply::pblapply(length_groups, function(k) {
    
    print(k)
    
    x <- dt %>%
      mutate(lg = paste(commonname, length_group)) %>% 
      filter(lg == k) %>%
      droplevels()
    
    mesh <- make_mesh(x, xy_cols = c("X", "Y"), cutoff = mesh_cutoff)
    
    ## Add land barriers to the mesh
    land_shape <- basemap_data(data = x, crs = dt_crs)$shapefiles$land
    mesh <-
      sdmTMBextra::add_barrier_mesh(mesh, land_shape, proj_scaling = 1000)
    
    ## alternative penalties example:
    ## apply a wide normal density penalty (i.e., prior) on the fixed effects
    ## so the model is identifiable even in years with no non-zero observations
    mm <- model.matrix(~ 0 + month + geartype + poly(log(depth), 2), data = x)
    # colnames(mm)
    sds <- rep(25, ncol(mm)) # pick a large penalty SD; applied to N(0, SD^2) on the fixed effects
    ## skipping the polynomials because their scale is dramatically different:
    sds[seq(length(sds) - 1, length(sds))] <- NA # skip on the polynomials
    # sds
    means <- rep(0, ncol(mm))
    means[seq(length(means) - 1, length(means))] <- NA # skip on the polynomials
    
    # Fit the model:
    mod <- try({sdmTMB(
      value ~ 0 + month + geartype + poly(log(depth), 2),
      data = x,
      mesh = mesh,
      control = sdmTMBcontrol(
        profile = c("b_j", "b_j2"), # faster
        multiphase = FALSE, # fine either way
        # start = list(b_j = b_start, b_j2 = b_start2),
        # map = list(b_j = b_map, b_j2 = b_map)
        newton_loops = 2
      ),
      priors = sdmTMBpriors(b = normal(means, sds)),
      # family = delta_lognormal(),
      family = model_family,
      spatial = "on",
      # spatiotemporal = 'ar1',
      time = "year",
      ## Below should possibly compile easier:
      share_range = TRUE,
      time_varying = ~ 1,
      time_varying_type = "ar1", # or "rw"
      spatiotemporal = "iid"
    )})
    
    # mod <- try({sdmTMB(
    #   value ~ 0 + month + geartype + s(depth),
    #   data = x,
    #   mesh = mesh,
    #   family = model_family,
    #   spatial = "on",
    #   spatiotemporal = 'ar1',
    #   time = "year",
    #   control = sdmTMBcontrol(newton_loops = 2)
    # )})
    
    if(inherits(mod, "try-error")) return(paste(k, "crashed"))
    
    out <- list(
      mod = mod, mesh = mesh, mesh_n = mesh$mesh$n, 
      mesh_cutoff = mesh_cutoff, crs = dt_crs,
      length_group = k
    )
    
    attributes(out) <- 
      c(attributes(out),
        list(
          grid_side_size = grid_side_size,
          response_variable = response_variable, 
          remove_zero_catch_surveys = remove_zero_catch_surveys,
          outlier_threshold = outlier_threshold,
          value_name = value_name,
          value_unit = value_unit,
          value_text_name = value_text_name,
          value_text_unit = value_text_unit)
      )
    
    saveRDS(
      out,
      file = 
        paste0("data/model_output/length_based/", model_name, "-", gsub(" ", "_", k), "-fit-",
               switch(response_variable, 
                      "catch biomass" = "cw", 
                      "biomass density" = "bm", 
                      "catch abundance" = "cc", 
                      "abundance density" = "ab"), "-",
               gsub("\\(.*\\)", "", model_family_name),
               ".rds"),
      compress = "xz"
    )
    
    return(paste(k, "saved to data/model_output/"))
  }, cl = length(length_groups))
}
```

## Load the models

```{r load-fit}
#| timeit: !expr if(recompile_model) TRUE else NULL

fit <- pbapply::pblapply(length_groups, function(k) {
  
  file <- 
    paste0("data/model_output/length_based/", model_name, "-", gsub(" ", "_", k), "-fit-",
           switch(response_variable, 
                  "catch biomass" = "cw", 
                  "biomass density" = "bm", 
                  "catch abundance" = "cc", 
                  "abundance density" = "ab"), "-",
           gsub("\\(.*\\)", "", model_family_name),
           ".rds")
  
  if(!file.exists(file)) return(NULL)
  
  readRDS(file)
}, cl = length(length_groups))

names(fit) <- length_groups

# switch to make shifting between tweedie and hurdle easier
hurdle_model <- length(fit[[1]]$mod$family$family) > 1
# switch to make including factor year easier
factor_year <- any(grepl("fyear", as.character(fit[[1]]$mod$formula[[1]])))
cruiseseries <- any(grepl("cruiseseries", as.character(fit[[1]]$mod$formula[[1]])))
geartype <- any(grepl("geartype", as.character(fit[[1]]$mod$formula[[1]])))
season <- any(grepl("season", as.character(fit[[1]]$mod$formula[[1]])))
month <- any(grepl("month", as.character(fit[[1]]$mod$formula[[1]])))
smonth <- any(grepl("s\\(month\\)", as.character(fit[[1]]$mod$formula[[1]])))
yday <- any(grepl("yday", as.character(fit[[1]]$mod$formula[[1]])))
commonname <- any(grepl("commonname", as.character(fit[[1]]$mod$formula[[1]])))
length_group <- any(grepl("length_group", as.character(fit[[1]]$mod$formula[[1]])))

terms <- all.vars(fit[[1]]$mod$formula[[1]])[-1]
terms <- terms[!grepl(":", terms)]
```

# Model specifications

- Type: `r model_type`
- Response variable: `r value_text_name`
- Formula: `r format(fit[[1]]$mod$formula[[1]])`
- Swept area as offset: `r offset`
- Model family: `r model_family_name`
- Hurdle model: `r hurdle_model`
- Spatiotemporal random fields: `r fit[[1]]$mod$spatiotemporal`
- Year in formula: `r factor_year`
- Species: `r list_values(unique(trimws(gsub("\\(.*\\]", "", names(fit)))))`
- Length group models: `r paste(names(fit), collapse = ", ")`
- Number of mesh vertices: `r paste(sapply(fit, function(k) k$mesh_n), collapse = ", ")`
- Projection: `r as.integer(dt_crs)`
- Number of years: `r length(unique(fit[[1]]$mod$data$year))`
- Number of stations (by species): `r list_values(Filter(Negate(is.null), unique(sapply(fit, function(k) nrow(k$mod$data)))))`
- Number of zero stations (by length group): `r list_values(sapply(fit, function(k) if(is.null(k)) NULL else nrow(k$mod$data %>% filter(value == 0))))`
- Number of non-zero stations (by length group): `r list_values(sapply(fit, function(k) if(is.null(k)) NULL else nrow(k$mod$data %>% filter(value != 0))))`

```{r}
#| label: tbl-sanity
#| tbl-cap: "Sanity check results for the models using the sdmTMB::sanity() function."
#| timeit: !expr if(recompile_model) TRUE else NULL

if(recompile_model) {
  fit_sanity <- 
    pbapply::pblapply(fit, function(k) {
      if(is.null(k)) return(NULL)
      sanity(k$mod, silent = TRUE)
    }, cl = length(fit))
  
  fit_sanity <- lapply(seq_along(fit_sanity), function(i) {
    
    cn <- unique(trimws(gsub("\\(.*\\]", "", names(fit_sanity[i]))))
    lg <- unique(trimws(gsub(cn, "", names(fit_sanity[i]))))
    
    if(is.null(fit_sanity[[i]])) {
      return({
        tibble(
          "Species" = case_when(
            cn == "snabeluer" ~ "Beaked redfish", 
            cn == "vanlig uer" ~ "Golden redfish"
          ),
          "LG" = lg, 
          "hessian_ok" = NA,
          "eigen_values_ok" = NA, 
          "nlminb_ok" = NA,
          "range_ok" = NA,
          "gradients_ok" = NA,
          "se_magnitude_ok" = NA,
          "se_na_ok" = NA,
          "sigmas_ok" = NA,
          "all_ok" = NA
        )
      })
    }
    
    data.frame(fit_sanity[[i]]) %>% 
      mutate(
        Species = case_when(
          cn == "snabeluer" ~ "Beaked redfish", 
          cn == "vanlig uer" ~ "Golden redfish"
        ),
        LG = lg,
        .before = 1
      )
  }) %>% 
    bind_rows() 
  
  write_csv(
    fit_sanity,
    paste0("data/model_output/length_based/", model_name, "-sanity-test-",
           switch(response_variable, 
                  "catch biomass" = "cw", 
                  "biomass density" = "bm", 
                  "catch abundance" = "cc", 
                  "abundance density" = "ab"), "-",
           gsub("\\(.*\\)", "", model_family_name),
           ".csv")
  )
} else {
  fit_sanity <- 
    read_csv(
      paste0("data/model_output/length_based/", model_name, "-sanity-test-",
             switch(response_variable, 
                    "catch biomass" = "cw", 
                    "biomass density" = "bm", 
                    "catch abundance" = "cc", 
                    "abundance density" = "ab"), "-",
             gsub("\\(.*\\)", "", model_family_name),
             ".csv")
    )
}

fit_sanity %>% 
  setNames(gsub('_ok', '', names(.))) %>% 
  setNames(gsub('_', '<br>', names(.))) %>% 
  kbl(escape = FALSE) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## Mesh

```{r}
#| label: fig-mesh
#| fig-height: 6
#| fig-cap: !expr paste0("Mesh for the sdmTMB model (black lines) with red dots showing the trawl stations. Size of the dots is scaled to ", ifelse(length_based, "the first length group ", "total "), sp_name, " catch. Triangles with X are on land and have been modelled as correlation barriers.")

if(overview_plots) {
  i = 1
  mesh <- fit[[i]]$mesh
  
  mesh_crs <- sf::st_crs(dt_crs)
  mesh_crs <- st_crs(sub("units=m", "units=km", mesh_crs$proj4string))
  
  mesh_df_water <- mesh$mesh_sf[mesh$normal_triangles, ]
  mesh_df_land <- mesh$mesh_sf[mesh$barrier_triangles, ]
  
  basemap(
    mesh$mesh$loc %>% as.data.frame() %>%
      mutate(V1 = 1000*V1, V2 = 1000*V2) %>% 
      st_as_sf(coords = c("V1", "V2"), crs = dt_crs), 
    crs = mesh_crs,
    land.col = "#eeeac4",
    lon.interval = 5) +
    geom_point(
      data = fit[[i]]$mod$data %>%  
        group_by(year, platformname, geartype, serialnumber, X, Y) %>% 
        reframe(value = sum(value)),
      aes(X, Y, size = value), alpha = 0.3, color = "red", shape = 21) +
    inlabru::gg(mesh$mesh, edge.color = "black") +
    geom_sf(
      data = mesh_df_land, shape = "X", size = FS(6)
    ) +
    coord_sf(datum = st_crs("+proj=utm +zone=33 +datum=WGS84 +units=km +no_defs"),
             expand = FALSE) +
    scale_y_continuous(
      "Latitude (UTM, km from the equator)",
      breaks = seq(0,1e4,1e3)
    ) +
    scale_x_continuous(
      "Longitude (UTM, km)",
      breaks = seq(-1e3,1e4,5e2)
    ) + 
    scale_size_area() +
    theme(legend.position = "none")
}
```

## Model output

```{r model-output, collapse=TRUE}
if(FALSE) lapply(fit, function(k) k$mod)
```

# Residuals

```{r}
#| label: fig-resid-qq
#| fig-height: !expr ceiling(length(fit)/2)
#| results: hide
#| fig-cap: !expr paste0("Quantile-quantile plots for the ", ifelse(hurdle_model & residual_method == "mle-mvn", "binomial (left) and positive (right) models.", "models."), ifelse(length_based, " Length group models are shown side-by-side.", ""))
#| timeit: !expr if(recompile_model) TRUE else NULL

if(TRUE) {
  set.seed(12)
  
  qq_dat <- 
    pbapply::pblapply(seq_along(fit), function(i) {
      
      cn <- unique(trimws(gsub("\\(.*\\]", "", names(fit[i]))))
      lg <- unique(trimws(gsub(cn, "", names(fit[i]))))
      
      if(is.null(fit[[i]])) {
        return({
          tibble(
            species = case_when(
              cn == "snabeluer" ~ "Beaked redfish", 
              cn == "vanlig uer" ~ "Golden redfish"
            ),
            length_group = lg,
            model = "crashed",
            x = NA
          )
        })
      }
      
      error_flag <- TRUE
      lapply(seq_along(fit[[i]]$mod$family$family), function(j) {
        rq_res <- try({residuals(fit[[i]]$mod, type = "mle-mvn", model = j)}, 
                      silent = TRUE)
        if(inherits(rq_res, "try-error")) {
          if(error_flag) message("mle-mvn type residuals did not work, shifting to mle-eb")
          rq_res <- try({residuals(fit[[i]]$mod, type = "mle-eb", model = j)}, 
                        silent = TRUE)
          error_flag = FALSE
          
          if(inherits(rq_res, "try-error")) {
            message("mle-eb type residuals did not work either, returning NA")
            return({tibble(
              species = case_when(
                cn == "snabeluer" ~ "Beaked redfish", 
                cn == "vanlig uer" ~ "Golden redfish"
              ),
              length_group = lg,
              model = fit[[i]]$mod$family$family[j],
              x = NA
            )
            })
          }
        }
        rq_res <- rq_res[is.finite(rq_res)] # some Inf
        tibble(
          species = case_when(
            cn == "snabeluer" ~ "Beaked redfish", 
            cn == "vanlig uer" ~ "Golden redfish"
          ),
          length_group = lg,
          model = fit[[i]]$mod$family$family[j],
          x = rq_res
        )
      }) %>% bind_rows()
    }, cl = length(fit)
    ) 
  
  cowplot::plot_grid(
    plotlist = 
      lapply(qq_dat, function(k) {
        if(any(k$model == "crashed")) {
          tibble(x = seq(-3,3,0.1)) %>%   
            ggplot(aes(x,x)) + 
            geom_blank() +
            annotate("text", x = 0, y = 0, label = "Model crashed") + 
            labs(
              title = paste(unique(k$species), unique(k$length_group)),
              x = "Theoretical Quantiles",
              y = "Sample Quantiles"
            )
        } else {
          k %>% 
            ggplot(aes(sample = x)) +
            geom_qq_line() +
            geom_qq(shape = 21, color = "grey60") +
            {if(length(unique(k$model) > 1)) facet_wrap(~ model, ncol = 2)} +
            labs(
              title = paste(unique(k$species), unique(k$length_group)),
              x = "Theoretical Quantiles",
              y = "Sample Quantiles"
            )
        }
      })
  )
}   
```

# Prediction settings

Prediction grid was created for the model to generate estimates for the study area.

## Prediction grid

Grid cells were modeled as points but converted to squares with `r grid_side_size` km sides according to the `r as.integer(dt_crs)` projection given as an EPSG code corresponding to the zone `r paste0(substr(as.integer(dt_crs),4,5), "N")` in the Universal Transverse Mercator (UTM) coordinate system.

```{r prediction-grid}
#| warning: true
#| message: false
#| results: hide
#| timeit: !expr if(recompile_model) TRUE else NULL

if(recompile_model) { # switch to recompile the prediction grid. If FALSE, data are loaded from a file.
  
  tmp <- sf::st_read("data/gis/prediction_region.gpkg", quiet = TRUE)
  
  proj_dt_sf <- tmp %>%
    st_transform(dt_crs) %>%
    st_make_grid(cellsize = rep(grid_side_size*1000,2),
                 square = TRUE, what = "centers") %>%
    st_sf() %>%
    st_filter(tmp)
  
  proj_dt <- proj_dt_sf %>%
    st_coordinates() %>%
    as_tibble() %>%
    mutate(X = X/1000, Y = Y/1000)
  
  # ru (raster user) uses the GEBCO 15 arc-second grid: https://www.gebco.net/data_and_products/gridded_bathymetry_data/
  # and seems to cause trouble with some systems.
  # rc uses a downsampled ETOPO 1 degree grid. Works on more systems but less
  # resolution than ru. Set the switch to FALSE if ru does not work.
  
  if(TRUE) { # switch to use 15-arcsecond GEBCO grid to acquire depths
    
    if(is.null(getOption("ggOceanMaps.userpath"))) { # set the path to GEBCO grid
      options(ggOceanMaps.userpath = "/gpfs/gpfs0/felles/Dyphavsarter/Data fellesressurser/Gebco/gebco_2023/GEBCO_2023.nc")
    }
    
    proj_dt$depth <-
      proj_dt %>%
      mutate(X = X*1000, Y = Y*1000) %>%
      st_as_sf(coords = c("X", "Y"), crs = dt_crs) %>%
      ggOceanMaps::get_depth(
        bathy.style = "ru",
        bind = FALSE)
  } else { # Downsampled ETOPO 1 degree grid
    
    proj_dt$depth <-
      proj_dt %>%
      mutate(X = X*1000, Y = Y*1000) %>%
      st_as_sf(coords = c("X", "Y"), crs = dt_crs) %>%
      ggOceanMaps::get_depth(
        bathy.style = "rc",
        bind = FALSE)
  }
  
  proj_dt <- proj_dt %>%
    filter(!is.na(depth)) %>%
    filter(depth >= 1 & depth <= max_prediction_depth) %>%
    replicate_df("year", unique(dt$year)) %>%
    replicate_df("geartype", "Shrimp") %>%
    mutate(fyear = as.factor(year))
  
  if(season) proj_dt <- proj_dt %>% replicate_df("season", "H2")
  if(month) proj_dt <- proj_dt %>% replicate_df("month", if(smonth) as.integer(9) else factor(9))
  if(yday) proj_dt <- proj_dt %>% replicate_df("yday", 245)
  if(commonname) proj_dt <- proj_dt %>% replicate_df("commonname", unique(dt$commonname))
  
  attributes(proj_dt)$grid_side_size <- grid_side_size
  
  saveRDS(proj_dt, file = "data/model_output/length_based/lbm-prediction-grid-data.rds", compress = "xz")
} else {
  proj_dt <- read_rds("data/model_output/length_based/lbm-prediction-grid-data.rds")
}
```

The resulting grid contained `r proj_dt %>% dplyr::select(X, Y) %>% distinct() %>% nrow()` points (@fig-pred-grid). Depths for these points were obtained from the [GEBCO 15 arcsecond bathymetry grid](https://www.gebco.net/data_and_products/gridded_bathymetry_data/) with the help of the [ggOceanMaps](https://mikkovihtakari.github.io/ggOceanMaps/reference/get_depth.html) package.

```{r}
#| label: fig-pred-grid
#| fig-height: 8
#| fig-cap: "Predition grid for the sdmTMB model. The actual UTM coordinates for the model grid have been divided by 1000 to help the model converge. Color of the points refers to the bottom depth category obtained from the ETOPO 15 arcsecond bathymetry grid. Actual bottom depth estimates are continuous."

breaks = seq(round_any(min(proj_dt$depth), 10, floor),
             round_any(max(proj_dt$depth), 10, ceiling),
             100)

basemap(dt %>% filter(!outlier)) + # land.col = "#eeeac4", land.border.col = NA
  geom_sf(
    data =
      proj_dt %>%
      dplyr::select(X, Y, depth) %>%
      distinct() %>%
      mutate(X = X*1e3, Y = Y*1e3) %>%
      st_as_sf(
        coords = c("X", "Y"),
        crs = suppressWarnings(dt_crs)
      ),
    aes(color = depth),
    shape = 15, size = 0.08*grid_side_size) +
  labs(color = "Depth (m)") +
  scale_size(range = c(6,6)) +
  binned_scale(
    aesthetics = "color",
    scale_name = "stepsn",
    palette = function(x) viridis::viridis_pal(option = "turbo")(length(breaks)+1),
    limits = c(min(breaks), max(breaks)),
    # guide = guide_bins(size = 10),
    breaks = breaks,
    show.limits = TRUE
  ) # +
# theme(legend.key.size = unit(20, "mm"))
# +
# guides(color = guide_legend(override.aes = list(size = 10)))
```

## Predict

```{r predict}
#| timeit: !expr if(recompile_model) TRUE else NULL

if(recompile_model) {
  mod_pred <- 
    pbapply::pblapply(fit, function(k) {
      predict(
        k$mod,
        newdata = proj_dt,
        type = "response",
        model = NA,
        return_tmb_object = TRUE)
    }, cl = length(fit))
  
  save(
    mod_pred,
    file = 
      paste0("data/model_output/length_based/", model_name, "-predictions-",
             switch(response_variable, 
                    "catch biomass" = "cw", 
                    "biomass density" = "bm", 
                    "catch abundance" = "cc", 
                    "abundance density" = "ab"), "-",
             gsub("\\(.*\\)", "", model_family_name),
             ".rda"),
    compress = "xz"
  )
} else {
  load(
    paste0("data/model_output/length_based/", model_name, "-predictions-",
           switch(response_variable, 
                  "catch biomass" = "cw", 
                  "biomass density" = "bm", 
                  "catch abundance" = "cc", 
                  "abundance density" = "ab"), "-",
           gsub("\\(.*\\)", "", model_family_name),
           ".rda")
  )
}
```

```{r, include = FALSE}
#| label: fig-pred-scatter
#| fig-height: !expr 1.5*ceiling(length(fit)/2)
#| timeit: !expr if(recompile_model) TRUE else NULL
#| fig-cap: "Observed swept area density estimates compared to averaged model simulations over 100 repetitions using the model for positions with data. A) Regression between observed and average simulated response variables. Dashed and blue lines indicate the line of equality and an average regression model, respectively. Average regression parameters, together with their 95% confidence intervals (CIs), are given under the length group title. '0 prob' means the proportion of simulations where a zero observation was zero. B) Difference between simulated and observed values for trawl stations: black dots indicate average values, error bars represent 95% CIs, and grey horizontal lines indicate zero. If the error bars do not cross the zero line, the difference is significant. C) Proportion of trawl stations with no target species. Average values simulated from the models are indicated as black dots, with 95% CIs as error bars. Grey horizontal lines indicate the observed values with all years combined."

dat_pred <- pbapply::pblapply(fit, function(k) {
  tmp <- simulate(k$mod, nsim = 100, model = NA)
  
  out <- tibble(
    example_sim = tmp[,1],
    mean_sim = rowMeans(tmp),
    sd_sim = apply(tmp,1, sd),
    pr_zero = apply(tmp,1, function(g) sum(g == 0))/ncol(tmp)
  ) %>%
    mutate(cv_sim = sd_sim/mean_sim, .after = "sd_sim") %>%
    mutate(value = k$mod$data$value, .before = 1)
  
  cf <- lapply(1:ncol(tmp), function(i) {
    moo <- summary(lm(tmp[,i] ~ k$mod$data$value))
    tibble(intercept = coef(moo)[1,1],
           slope = coef(moo)[2,1],
           r2 = moo$r.squared)
  }) %>%
    bind_rows() %>%
    reframe(
      int_mean = mean(intercept),
      int_sd = sd(intercept),
      int_min = quantile(intercept, 0.025),
      int_max = quantile(intercept, 0.975),
      slp_mean = mean(slope),
      slp_sd = sd(slope),
      slp_min = quantile(slope, 0.025),
      slp_max = quantile(slope, 0.975),
      r2_mean = mean(r2),
      r2_sd = sd(r2),
      r2_min = quantile(r2, 0.025),
      r2_max = quantile(r2, 0.975)
    ) %>%
    bind_cols(
      out %>%
        filter(value == 0) %>%
        reframe(
          zero_mean = mean(pr_zero),
          zero_min = quantile(pr_zero, 0.025),
          zero_max = quantile(pr_zero, 0.975)
        )
    ) %>%
    mutate(zero_obs = mean(k$mod$data$value == 0)) %>%
    bind_cols(
      tibble(
        sim_mean = mean(colMeans(tmp)),
        sim_sd = sd(colMeans(tmp)),
        sim_min = quantile(colMeans(tmp), 0.025),
        sim_max = quantile(colMeans(tmp), 0.975),
        obs_mean = mean(out$value),
        res_mean = mean(colMeans(tmp - k$mod$data$value)),
        res_min = quantile(colMeans(tmp - k$mod$data$value), 0.025),
        res_max = quantile(colMeans(tmp - k$mod$data$value), 0.975),
      )
    )
  
  list(dt = out, coef = cf)
}, cl = length(fit))

if(FALSE) {
  cowplot::plot_grid(
    plotlist =
      lapply(seq_along(dat_pred), function(i) {
        cowplot::plot_grid(
          ggplot(dat_pred[[i]]$dt, aes(x = value, y = mean_sim)) +
            geom_abline(intercept = 0, slope = 1, linetype = 2) +
            geom_point(shape = 21, alpha = 0.3) +
            geom_abline(
              intercept = dat_pred[[i]]$coef$int_mean,
              slope = dat_pred[[i]]$coef$slp_mean,
              color = "blue"
            ) +
            annotate(
              "text", x = -Inf, y = Inf,
              label =
                paste0(
                  "R2 = ", round(dat_pred[[i]]$coef$r2_mean, 2), " (",
                  round(dat_pred[[i]]$coef$r2_min, 2), "-",
                  round(dat_pred[[i]]$coef$r2_max, 2), ")\nslope = ",
                  round(dat_pred[[i]]$coef$slp_mean, 2), " (",
                  round(dat_pred[[i]]$coef$slp_min, 2), "-",
                  round(dat_pred[[i]]$coef$slp_max, 2), ")\n0 prob = ",
                  round(dat_pred[[i]]$coef$zero_mean, 2), " (",
                  round(dat_pred[[i]]$coef$zero_min, 2), "-",
                  round(dat_pred[[i]]$coef$zero_max, 2), ")"
                ),
              vjust = 1, hjust = -0.1
            ) +
            labs(x = bquote('Observed response ('*.(value_unit)*')'), 
                 y = bquote('Average simulated response ('*.(value_unit)*')'),
                 title = names(dat_pred[i])),
          
          ggplot(dat_pred[[i]]$coef,
                 aes(x=1, y = res_mean, ymin = res_min, ymax = res_max)) +
            geom_hline(yintercept = 0, color = "grey") +
            geom_errorbar(width = 0.2) +
            geom_point() +
            labs(y = bquote('Mean simulated - observed (± 95% CIs; '*.(value_unit)*')')) +
            scale_x_continuous(expand = c(1,0)) +
            theme(
              axis.title.x = element_blank(),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank()),
          
          ggplot(data = dat_pred[[i]]$coef,
                 aes(x=1, y = zero_mean, ymin = zero_min, ymax = zero_max)) +
            geom_hline(aes(yintercept = zero_obs), color = "grey") +
            geom_errorbar(width = 0.2) +
            geom_point() +
            labs(y = "Proportion of zero observations (± 95% CIs)") +
            scale_x_continuous(expand = c(1,0)) +
            scale_y_continuous(limits = c(0,1), expand = c(0,0)) +
            theme(
              axis.title.x = element_blank(),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank()),
          
          rel_widths = c(6,2,2), align = "h", nrow = 1, labels = "AUTO"
        )
      }),
    ncol = ifelse(length(fit) == 1, 1, 2)
  )
}
```

```{r}
#| fig-height: !expr ifelse(length(fit) == 1, 6, 3.5*ceiling(length(fit)/2))
#| label: fig-pred-hist
#| timeit: !expr if(recompile_model) TRUE else NULL
#| fig-cap: "Histograms showing distributions of the log + 1 transformed observed density (A), average simulated density for positions with data (B), and predicted density throughout the predicted model domain (C). Blue lines indicate mean response variable values."

if(TRUE) {
  dat_pred <- pbapply::pblapply(fit, function(k) {
    tmp <- simulate(k$mod, nsim = 100, model = NA)
    
    out <- tibble(
      example_sim = tmp[,1],
      mean_sim = rowMeans(tmp),
      sd_sim = apply(tmp,1, sd),
      pr_zero = apply(tmp,1, function(g) sum(g == 0))/ncol(tmp)
    ) %>%
      mutate(cv_sim = sd_sim/mean_sim, .after = "sd_sim") %>%
      mutate(value = k$mod$data$value, .before = 1)
    
    cf <- lapply(1:ncol(tmp), function(i) {
      moo <- summary(lm(tmp[,i] ~ k$mod$data$value))
      tibble(intercept = coef(moo)[1,1],
             slope = coef(moo)[2,1],
             r2 = moo$r.squared)
    }) %>%
      bind_rows() %>%
      reframe(
        int_mean = mean(intercept),
        int_sd = sd(intercept),
        int_min = quantile(intercept, 0.025),
        int_max = quantile(intercept, 0.975),
        slp_mean = mean(slope),
        slp_sd = sd(slope),
        slp_min = quantile(slope, 0.025),
        slp_max = quantile(slope, 0.975),
        r2_mean = mean(r2),
        r2_sd = sd(r2),
        r2_min = quantile(r2, 0.025),
        r2_max = quantile(r2, 0.975)
      ) %>%
      bind_cols(
        out %>%
          filter(value == 0) %>%
          reframe(
            zero_mean = mean(pr_zero),
            zero_min = quantile(pr_zero, 0.025),
            zero_max = quantile(pr_zero, 0.975)
          )
      ) %>%
      mutate(zero_obs = mean(k$mod$data$value == 0)) %>%
      bind_cols(
        tibble(
          sim_mean = mean(colMeans(tmp)),
          sim_sd = sd(colMeans(tmp)),
          sim_min = quantile(colMeans(tmp), 0.025),
          sim_max = quantile(colMeans(tmp), 0.975),
          obs_mean = mean(out$value),
          res_mean = mean(colMeans(tmp - k$mod$data$value)),
          res_min = quantile(colMeans(tmp - k$mod$data$value), 0.025),
          res_max = quantile(colMeans(tmp - k$mod$data$value), 0.975),
        )
      )
    
    list(dt = out, coef = cf)
  }, cl = length(fit))
  
  cowplot::plot_grid(
    plotlist =
      lapply(seq_along(mod_pred), function(i) {
        
        tmp <- fit[[i]]$mod$data
        
        cowplot::plot_grid(
          
          ggplot(tmp, aes(x = value)) +
            geom_histogram() +
            geom_vline(xintercept = mean(tmp$value), color = "blue") +
            scale_y_continuous(expand = c(0.01,0)) +
            scale_x_continuous(
              trans = "log1p",
              breaks = c(1, 10, 100, 1e3, 3e3, 5e3, 1e4, 5e4, 1e5, 1e6, 1e7),
              expand = expansion(mult = 0.01)
            ) +
            expand_limits(
              x = c(0, ceiling(max(c(tmp$value, dat_pred[[i]]$dt$mean_sim,
                                     mod_pred[[i]]$data$est))))
            ) +
            labs(x = bquote('log(observed response + 1) ('*.(value_unit)*')')),
          
          ggplot(dat_pred[[i]]$dt, aes(x = mean_sim)) +
            geom_histogram() +
            geom_vline(xintercept = mean(dat_pred[[i]]$dt$mean_sim),
                       color = "blue") +
            scale_y_continuous(expand = c(0.01,0)) +
            scale_x_continuous(
              trans = "log1p",
              breaks = c(1, 10, 100, 1e3, 3e3, 5e3, 1e4, 5e4, 1e5, 1e6, 1e7),
              expand = expansion(mult = 0.01)
            ) +
            expand_limits(
              x = c(0, ceiling(max(c(tmp$value, dat_pred[[i]]$dt$mean_sim,
                                     mod_pred[[i]]$data$est))))
            ) +
            labs(x = bquote('log(mean simulated response for observations + 1) ('*.(value_unit)*')')),
          
          ggplot(mod_pred[[i]]$data, aes(x = est)) +
            geom_histogram() +
            geom_vline(xintercept = mean(mod_pred[[i]]$data$est), color = "blue") +
            scale_y_continuous(expand = c(0.01,0)) +
            scale_x_continuous(
              trans = "log1p",
              breaks = c(1, 10, 100, 1e3, 3e3, 5e3, 1e4, 5e4, 1e5, 1e6, 1e7),
              expand = expansion(mult = 0.01)
            ) +
            expand_limits(
              x = c(0, ceiling(max(c(tmp$value, dat_pred[[i]]$pred$est, mod_pred[[i]]$data$est))))
            ) +
            labs(x = bquote('log(predicted response for model domain + 1) ('*.(value_unit)*')')),
          
          ncol = 1,
          labels = "AUTO"
        )
      }),
    labels = names(mod_pred), ncol = ifelse(length(fit) == 1, 1, 2)
  )
}
```

# Model effects

## Fixed effects

```{r, timeit = recalculate_fixed_effects}
#| label: fig-fixed-effects
#| fig-height: !expr floor(length(fit)/1.1)
#| timeit: !expr if(recalculate_fixed_effects) TRUE else NULL
#| fig-cap: Conditional main (fixed) effects for the combined sdmMTB models predicted over a range of gear types (A, D), months (B, E), and bottom depths (C, F) for beaked (A-C) and golden (D-F) redfish. Colors refer to gear type and the grey vertical line in C to the reference depth used to estimate other fixed effects. September 2024 was used as a temporal reference point. Error bars and shading indicate 95% confidence intervals. Spatial and spatiotemporal effects are not included to the output.

if(recalculate_fixed_effects) { 
  fixef_dt <- pbapply::pblapply(seq_along(fit), function(i) {
    
    ### Reference depth
    
    depth_dt <- predict(
      fit[[i]]$mod, 
      tibble(depth = 1:max_prediction_depth) %>% 
        replicate_df(., "year", max(fit[[i]]$mod$data$year)) %>%
        {if(geartype) replicate_df(., "geartype", "Shrimp") else .} %>% 
        {if(month) replicate_df(., "month", if(smonth) as.integer(9) else as.factor(9)) else .} %>%  
        {if(factor_year) mutate(., fyear = factor(year)) else .},
      se_fit = TRUE, re_form = NA, model = NA
    ) %>%
      mutate(commonname = unique(fit[[i]]$mod$data$commonname)) %>%
      mutate(
        lwr = exp(est - 1.96*est_se), 
        upr = exp(est + 1.96*est_se), 
        est = exp(est)
      ) %>% 
      mutate(geartype = recode(geartype, "Cod" = "Cod trawls",
                               "Shrimp" = "Shrimp trawls"))
    
    ref_depth <- depth_dt %>% 
      slice(which.max(est)) %>%
      pull(depth)
    
    other_dt <- predict(
      fit[[i]]$mod,
      tibble(depth = ref_depth) %>%
        {if(factor_year) 
          replicate_df(., "year", unique(fit[[i]]$mod$data$year)) else 
            replicate_df(., "year", max(fit[[i]]$mod$data$year))} %>% 
        {if(geartype) 
          replicate_df(., "geartype", unique(fit[[i]]$mod$data$geartype)) else .} %>% 
        {if(month) replicate_df(., "month", if(smonth) 1:12 else factor(1:12)) else .} %>%  
        {if(factor_year) mutate(., fyear = factor(year)) else .},
      se_fit = TRUE, re_form = NA, model = NA
    ) %>%
      mutate(commonname = unique(fit[[i]]$mod$data$commonname)) %>%
      mutate(
        lwr = exp(est - 1.96*est_se), 
        upr = exp(est + 1.96*est_se), 
        est = exp(est)
      ) %>% 
      mutate(geartype = recode(geartype, "Cod" = "Cod trawls",
                               "Shrimp" = "Shrimp trawls"))
    
    return(list(depth = depth_dt, other = other_dt, ref_depth = ref_depth, lg = names(fit)[i]))
    
  }, cl = length(fit)) 
  
  fixef_dt <- setNames(fixef_dt, names(fit))
  
  save(fixef_dt, file = "data/model_output/length_based/fixed_effect_data.rda", compress = "xz")
} else {
  if(run_fixef) load("data/model_output/length_based/fixed_effect_data.rda")
}

if(run_fixef) {
  gear_cols <- c("Cod trawls" = "#449BCF", "Shrimp trawls" =  "#82C893", "Both" = "#FF5F68")
  
  cowplot::plot_grid(
    plotlist = lapply(seq_along(fixef_dt), function(i) {
      
      ymax <- ifelse(max(fixef_dt[[i]]$depth$upr) > 1.1*max(fixef_dt[[i]]$depth$est), 1.1*max(fixef_dt[[i]]$depth$est), ceiling(max(fixef_dt[[i]]$depth$upr)))
      
      p_depth <- fixef_dt[[i]]$depth %>%
        ggplot(aes(x = depth, y = est, ymin = lwr, ymax = upr,
                   color = geartype, fill = geartype)) +
        geom_vline(xintercept = fixef_dt[[i]]$ref_depth) +
        geom_ribbon(alpha = 0.3, color = NA) +
        geom_path() +
        scale_color_manual(values = gear_cols) +
        scale_fill_manual(values = gear_cols) +
        coord_cartesian(expand = FALSE, ylim = c(0,ymax)) +
        labs(y = value_name, x = "Depth (m)", color = "Gear type", fill = "Gear type") +
        theme_bw(base_size = 10) +
        theme(legend.position = "none",
              strip.background = element_blank(),
              panel.background = element_blank(),
              plot.background = element_blank(),
              legend.background = element_blank(),
              legend.box.background = element_blank(),
              plot.margin = margin(5.5, 10, 5.5, 5.5))
      
      ### Year
      
      if(factor_year) {
        p_year <- fixef_dt[[i]]$other %>%
          {if(geartype) filter(.,geartype == "Shrimp trawls") else .} %>%
          {if(month) filter(., month == 9) else .} %>% 
          ggplot(
            aes(x = year, y = est, ymin = lwr, ymax = upr, color = geartype)
          ) +
          geom_path() +
          geom_errorbar(width = 0.2) +
          geom_point() +
          # scale_y_log10(n.breaks = 8) +
          scale_x_continuous(
            breaks = seq(1980, 2030, 4),
            labels = substr(seq(1980, 2030, 4),3,4)
          ) +
          scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                             limits = c(0, NA)) +
          coord_cartesian(expand = FALSE) +
          scale_color_manual(values = gear_cols) +
          theme_bw(base_size = 10) +
          # {if(length_based) facet_wrap(~length_group) } +
          labs(x = "Year", y = value_name) +
          theme(legend.position = "none")
      }
      
      ### Gear type
      
      if(geartype) {
        p_gear <- fixef_dt[[i]]$other %>%
          filter(year == max(year)) %>%
          {if(month) filter(., month == 9) else .} %>% 
          ggplot(aes(x = geartype, y = est, ymin = lwr, ymax = upr, 
                     color = geartype)) +
          geom_errorbar(width = 0.2) +
          geom_point() +
          scale_color_manual(values = gear_cols) +
          scale_y_continuous(
            transform = "log1p",
            n.breaks = 8,
            expand = expansion(mult = c(0, 0.1)),
            limits = c(0, NA)) +
          labs(y = value_name, x = "Gear type") +
          theme_bw(base_size = 10) +
          theme(legend.position = "none",
                strip.background = element_blank(),
                panel.background = element_blank(),
                plot.background = element_blank(),
                legend.background = element_blank(),
                legend.box.background = element_blank(),
                plot.margin = margin(5.5, 10, 5.5, 5.5))
      }
      
      ### Month
      
      if(month) {
        if(inherits(fit[[i]]$mod$data$month, c("numeric", "integer"))) {
          p_month <- fixef_dt[[i]]$other %>%
            filter(year == max(year)) %>%
            {if(geartype) filter(.,geartype == "Shrimp trawls") else .} %>%
            ggplot(
              aes(x = month, y = est, ymin = lwr, ymax = upr, 
                  color = geartype, group = geartype, fill = geartype)
            ) +
            geom_ribbon(alpha = 0.5, color = NA) +
            geom_path() +
            geom_point() +
            scale_x_continuous(breaks = 1:12, 
                               expand = expansion(mult = c(0.05, 0.05))) +
            scale_y_continuous(
              transform = "log1p",
              n.breaks = 3,
              expand = expansion(mult = c(0, 0.1)),
              limits = c(0, NA)) +
            scale_color_manual(values = gear_cols) +
            scale_fill_manual(values = gear_cols) +
            labs(x = "Month", y = value_name) +
            theme_bw(base_size = 10) +
            theme(legend.position = "none",
                  strip.background = element_blank(),
                  panel.background = element_blank(),
                  plot.background = element_blank(),
                  legend.background = element_blank(),
                  legend.box.background = element_blank(),
                  panel.grid.minor = element_blank(),
                  plot.margin = margin(5.5, 10, 5.5, 5.5)
            )
        } else {
          p_month <- fixef_dt[[i]]$other %>%
            filter(year == max(year)) %>%
            {if(geartype) filter(.,geartype == "Shrimp trawls") else .} %>%
            ggplot(
              aes(x = month, y = est, ymin = lwr, ymax = upr, 
                  color = geartype, group = geartype)
            ) +
            geom_path() +
            geom_errorbar(width = 0.2) +
            geom_point() +
            # scale_y_log10(n.breaks = 8) +
            scale_x_discrete(breaks = 1:12, expand = expansion(mult = c(0.05, 0.05))) +
            scale_y_continuous(
              transform = "log1p",
              n.breaks = 3,
              expand = expansion(mult = c(0, 0.1)),
              limits = c(0, NA)) +
            scale_color_manual(values = gear_cols) +
            # theme_bw(base_size = 10) +
            # {if(length_based) facet_wrap(~length_group) } +
            labs(x = "Month", y = value_name) +
            theme_bw(base_size = 10) +
            theme(legend.position = "none",
                  strip.background = element_blank(),
                  panel.background = element_blank(),
                  plot.background = element_blank(),
                  legend.background = element_blank(),
                  legend.box.background = element_blank(),
                  plot.margin = margin(5.5, 10, 5.5, 5.5)
            )
        }
      }
      
      title <- cowplot::ggdraw() + 
        cowplot::draw_label(fixef_dt[[i]]$lg)
      
      if(month & factor_year) {
        return({
          cowplot::plot_grid(
            title, 
            cowplot::plot_grid(
              p_gear, p_year + ylab(NULL), rel_widths = c(2,6), ncol = 2, 
              labels = c("A", "B"), align = "v"),
            cowplot::plot_grid(
              p_month, p_depth + ylab(NULL), ncol = 2, rel_widths = c(3,6),
              labels = c("C", "D"), align = "v"),
            rel_heights = c(1,9,9), ncol = 1, align = "v"
          )
        })
      } 
      
      if(month & !factor_year) {
        return({
          cowplot::plot_grid(
            title, 
            cowplot::plot_grid(
              p_gear, p_month + ylab(NULL), p_depth + ylab(NULL), 
              rel_widths = c(3,3,4), nrow = 1
            ),
            rel_heights = c(1,9), ncol = 1
          )
        })
      } else {
        return({
          cowplot::plot_grid(
            p_gear, p_depth + ylab(NULL), 
            ncol = 2, labels = c("A", "B"), rel_widths = c(2,2)
          )
        })
      }
    }),
    ncol = 2
  )
}
```

```{r}
#| label: tbl-conditional-values
#| tbl-cap: "Conditional values used for fixed effects in model predictions and fixed effect plots."

if(run_fixef) {
  lapply(fixef_dt, function(k) {
    tibble(
      `length group` = k$lg, 
      depth = k$ref_depth,
      year = unique(k$depth$year),
      geartype = unique(k$depth$geartype),
      month = unique(k$depth$month),
    )
  }) %>% 
    bind_rows() %>% 
    make_table()
}
```

```{r}
#| label: fig-simulated-effects
#| fig-height: !expr floor(length(fit)/1.1)
#| timeit: !expr if(recalculate_fixed_effects) TRUE else NULL
#| fig-cap: !expr paste0("Average values by fixed effect (", paste(terms, collapse = ", "), ") for observed (yellow) and simulated (grey) data. Data type is indicated as line type and shape, and gear type as color. Shading and error bars indicate confidence intervals. Simulated data are a result of 100 repetitions. Spatial and spatiotemporal effects are included to the results.")

if(run_fixef) { # turn off the plot using this switch
  
  cowplot::plot_grid(
    plotlist = 
      pbapply::pblapply(seq_along(fit), function(i) {
        k = fit[[i]]
        
        tmp <- simulate(k$mod, nsim = 100, model = NA) %>%
          as.data.frame()
        
        colnames(tmp) <- paste0("sim", 1:ncol(tmp))
        
        pp <- cowplot::plot_grid(
          plotlist =
            
            lapply(terms, function(j) {
              tmp <- bind_cols(
                k$mod$data[j],
                tmp
              ) %>% 
                pivot_longer(cols = c(-all_of(j))) %>% 
                group_by(across(all_of(j)), name) %>%
                reframe(value = mean(value)) %>%
                group_by(across(all_of(j))) %>%
                reframe(mean = mean(value),
                        sd = sd(value),
                        n = n(),
                        lwr = mean - 1.96*sd, # quantile(value,0.025),
                        upr = mean + 1.96*sd #quantile(value,0.975)
                ) %>%
                mutate(type = "Simulated") %>%
                bind_rows(
                  k$mod$data %>%
                    group_by(across(all_of(j))) %>%
                    reframe(
                      mean = mean(value),
                      sd = sd(value),
                      n = n(),
                      lwr = mean - 1.96*(sd/sqrt(n)),
                      upr = mean + 1.96*(sd/sqrt(n))
                    ) %>%
                    mutate(type = "Observed")
                ) 
              
              if(inherits(tmp[[j]], c("numeric", "integer"))) {
                tmp %>% 
                  ggplot(aes(x = get(j), y = mean, ymin = lwr, ymax = upr,
                             shape = type, color = type, fill = type)) +
                  geom_smooth() + 
                  # geom_ribbon(color = NA, alpha = 0.3) +
                  # geom_path(linewidth = 1) +
                  scale_color_manual(values = c("#f7cb15", "#878e88")) + 
                  scale_fill_manual(values = c("#f7cb15", "#878e88")) +
                  labs(x = j, y = if(which(terms %in% j) == 1) value_name else NULL,
                       color = "Data", fill = "Data") +
                  coord_cartesian(ylim = c(0, NA), expand = FALSE) +
                  theme_bw() +
                  theme(
                    legend.position = "none", 
                    # if(which(terms %in% j) == length(terms)) c(0.9,0.9) else "none",
                    strip.background = element_blank(),
                    panel.background = element_blank(),
                    plot.background = element_blank(),
                    legend.background = element_blank(),
                    legend.box.background = element_blank() #,
                    # plot.margin = margin(5.5, 10, 5.5, 5.5)
                  )
              } else {
                tmp %>% 
                  ggplot(aes(x = get(j), y = mean, ymin = lwr, ymax = upr,
                             color = type, fill = type,
                             group = type)) +
                  geom_errorbar(width = 0.2, position = position_dodge(0.5)) +
                  geom_point(position = position_dodge(0.5)) +
                  # geom_ribbon(color = NA, alpha = 0.3) +
                  # geom_path(linewidth = 1) +
                  scale_color_manual(values = c("#f7cb15", "#878e88")) + 
                  scale_fill_manual(values = c("#f7cb15", "#878e88")) +
                  labs(x = j, y = if(which(terms %in% j) == 1) value_name else NULL,
                       color = "Data", fill = "Data") +
                  #coord_cartesian(ylim = c(0, NA), expand = FALSE) +
                  theme_bw() +
                  theme(
                    legend.position = "none", 
                    # if(which(terms %in% j) == length(terms)) c(0.9,0.9) else "none",
                    strip.background = element_blank(),
                    panel.background = element_blank(),
                    plot.background = element_blank(),
                    legend.background = element_blank(),
                    legend.box.background = element_blank() #,
                    # plot.margin = margin(5.5, 10, 5.5, 5.5)
                  )
              } 
            }), 
          nrow = 1
        )
        
        title <- cowplot::ggdraw() + 
          cowplot::draw_label(names(fit)[i])
        
        cowplot::plot_grid(
          title, pp,rel_heights = c(1,9), ncol = 1
        )
      }, cl = length(fit)),
    ncol = 2
  )
}
```

## Random effects

```{=html}
<details>
<summary>Click to see the random effect table</summary>
```

```{r, max.height='100px'}
#| label: tbl-random-effects
#| tbl-cap: "Random effects (estimate column) and their 95% confidence intervals for each length group. Model family and estimate space are given in the family and link columns. Range, sigma_O, sigma_E and phi in the term column mean Matérn range, marginal standard deviation of spatial and spatiotemporal random fields as well as the observation error scale parameter, respectively."

pbapply::pblapply(seq_along(fit), function(j) {
  
  lapply(seq_along(fit[[j]]$mod$family$family), function(i) {
    tidy(fit[[j]]$mod, "ran_pars", conf.int = TRUE, model = i) %>%
      dplyr::select(-std.error) %>%
      mutate(family = fit[[j]]$mod$family$family[i],
             link = fit[[j]]$mod$family$link[i],
             .before = 1)
  }) %>%
    bind_rows() %>%
    mutate(`length group` = names(fit)[j], .before = 1)
}, cl = length(fit)) %>%
  bind_rows() %>%
  mutate(across(where(is.numeric), ~ ifelse(.x > 1e6, Inf, .x))) %>% 
  mutate(family = stringr::str_to_title(family)) %>% 
  rename_with(function(k) gsub("_|\\.", " ", k)) %>% 
  rename_with(stringr::str_to_title) %>% 
  make_table(digits = 2)
```

```{=html}
</details>
```

## Spatial random fields

```{r, timeit = recompile_model}
#| label: fig-spatial-effect-map
#| fig-height: !expr length(fit)*1.5
#| timeit: !expr if(recompile_model) TRUE else NULL
#| fig-cap: !expr paste0("Spatial random fields (the omega parameter) for the ", ifelse(hurdle_model, "models. Left panels represent the binomial model and right panels the positive component model.", "model."))

cowplot::plot_grid(
  plotlist =
    pbapply::pblapply(seq_along(mod_pred), function(i) {
      
      tmp <- mod_pred[[i]]$data %>%
        filter(year == max(year)) 
      
      if(!any(grepl("omega", colnames(tmp)))) {
        return({
          cowplot::plot_grid(
            plotlist = lapply(fit[[i]]$mod$family$family, function(k) {
              ggplot() + 
                geom_blank() + 
                annotate("text", x = 0, y = 0, 
                         label = "No spatial\nrandom fields", size = FS(10)) + 
                theme_void()
            })
          )
        })
      }
      
      tmp <- tmp%>%
        dplyr::select(matches(c("X", "Y", "omega"), ignore.case = FALSE)) %>%
        distinct() %>%
        pivot_longer(c(-X, -Y)) %>%
        mutate(X = 1000*X, Y = 1000*Y) %>%
        st_as_sf(
          coords = c("X", "Y"),
          crs = suppressWarnings(dt_crs),
          remove = FALSE
        ) %>%
        mutate(name = factor(name))
      
      levels(tmp$name) <- fit[[i]]$mod$family$family
      
      cowplot::plot_grid(
        plotlist = lapply(levels(tmp$name), function(k) {
          
          if(nrow(tmp %>% filter(name == k)) == 0) {
            return({
              ggplot() + 
                geom_blank() + 
                annotate("text", x = 0, y = 0, 
                         label = "No spatial\nrandom fields", size = FS(10)) + 
                theme_void()
            })
          }
          
          basemap(tmp %>% filter(name == k), crs = dt_crs) +
            geom_raster(
              data = tmp %>% filter(name == k) %>% st_drop_geometry(),
              aes(x = X, y = Y, fill = value)) +
            scale_fill_gradient2() +
            labs(subtitle = k) +
            theme(legend.position = "bottom",
                  axis.title = element_blank(),
                  axis.text = element_blank())
        })
      )
    }),
  labels = names(mod_pred), ncol = ifelse(length(fit) > 2,2,1)
)
```

# Model results

```{r}
#| label: fig-result-map-detail
#| fig-height: !expr floor(length(fit)*0.9)
#| timeit: !expr if(recompile_model) TRUE else NULL
#| fig-cap: Spatial overall density averaged over years and fixed effects. The color scale has been square-root transformed to make low values more visible.

cowplot::plot_grid(
  plotlist =
    pbapply::pblapply(seq_along(mod_pred), function(i) {
      
      tmp <- mod_pred[[i]]$data %>%
        group_by(X, Y) %>%
        reframe(est = mean(est)) %>%
        mutate(X = 1000*X, Y = 1000*Y) %>%
        st_as_sf(
          coords = c("X", "Y"),
          crs = suppressWarnings(dt_crs),
          remove = FALSE
        )
      
      basemap(tmp, crs = dt_crs, base_size = 10) +
        geom_raster(
          data = tmp,
          aes(x = X, y = Y, fill = est)) +
        scale_fill_viridis_c(
          trans = "sqrt",
          breaks = trans_breaks("sqrt", function(x) x^2)
        ) +
        labs(x = "Latitude", y = "Longitude", fill = NULL) +
        theme(
          legend.position = "bottom",
          strip.background = element_blank(),
          panel.background = element_blank(),
          plot.background = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          legend.background = element_blank(),
          legend.box.background = element_blank(),
          plot.margin = margin(5.5, 10, 5.5, 5.5)
        )
    }),
  labels = names(mod_pred), hjust = 0
)
```

```{r}
#| label: fig-result-map-year
#| fig-height: !expr length(fit)*3
#| timeit: !expr if(recompile_model) TRUE else NULL
#| fig-cap: !expr paste(gsub("(^[[:alpha:]])", "\\U\\1", response_variable, perl=TRUE), "anomaly by", ifelse(length_based, "decade for each length group.", "year."))


tmp <- pbapply::pblapply(seq_along(mod_pred), function(i) { 
  
  decadal_means <- mod_pred[[i]]$data %>%
    {if(length_based) {
      mutate(., time = cut(
        year, seq(1980,2030,10), 
        labels = paste(seq(1980,2020,10), seq(1990,2030,10), sep = "-")))
    } else {mutate(., time = year)}} %>% 
    group_by(time, X, Y) %>%
    reframe(est = mean(est)) %>%
    mutate(X = 1000*X, Y = 1000*Y, length_group = names(mod_pred)[i])
  
  overall_means <- decadal_means %>%
    group_by(length_group, X, Y) %>% 
    reframe(mean = mean(est))
  
  total_means <- overall_means %>% 
    group_by(length_group) %>% 
    reframe(total = mean(mean)) %>% 
    pull(total)
  
  overall_max <- overall_means %>% 
    group_by(length_group) %>% 
    reframe(max = max(mean)) %>% 
    pull(max)
  
  full_join(decadal_means, overall_means, 
            by = join_by(length_group, X, Y)) %>%
    mutate(log_anomaly = log10(est) - log10(mean),
           anomaly = est - mean,
           pr_anomaly = 100*(est - mean)/mean,
           total_pr_anomaly = 100*(est - mean)/total_means,
           max_pr_anomaly = 100*(est - mean)/overall_max) %>%  
    st_as_sf(
      coords = c("X", "Y"), 
      crs = suppressWarnings(dt_crs),
      remove = FALSE
    )
}) 

## Plot ####
cowplot::plot_grid(
  plotlist = 
    lapply(seq_along(tmp), function(i) {
      
      response <- "max_pr_anomaly"
      limit_rounding <- 10 # rounding for limits
      break_rounding <- 5 # rounding for breaks
      
      limits <- c(floor(quantile(tmp[[i]][[response]], c(0))/limit_rounding)*limit_rounding, 
                  ceiling(quantile(tmp[[i]][[response]], c(1))/limit_rounding)*limit_rounding)
      
      breaks <- seq(sign(limits[1])*sqrt(abs(limits[1])), sqrt(limits[2]), 
                    length.out = 6)
      breaks <- round(sign(breaks)*breaks^2/break_rounding)*break_rounding
      
      basemap(tmp[[i]], crs = dt_crs, base_size = 10) +
        geom_raster(
          data = tmp[[i]], 
          aes(x = X, y = Y, fill = get(response))) +
        scale_fill_gradient2(
          trans = scales::trans_new(
            name = "negsqrt",
            transform = function(x) {
              if_else(x < 0, -1*sqrt(abs(x)), sqrt(abs(x)))
            },
            inverse = function(x) {
              if_else(x < 0, -1*x^2, x^2)
            }
          ),
          breaks = breaks,
          midpoint = 0,
          limits = limits
        ) +
        labs(
          x = "Latitude", y = "Longitude", 
          title = unique(tmp[[i]]$length_group),
          fill = ifelse(
            response %in% c("pr_anomaly", "total_pr_anomaly", "max_pr_anomaly"), 
            paste0(gsub("(^[[:alpha:]])", "\\U\\1", response_variable, perl=TRUE), "\npercent anomaly"),
            expression(atop(Biomass~density, paste("anomaly (kg ", nmi^{-2}, ")")))
          )
          # 
          # expression(paste("Biomass density\nanomaly (kg ", nmi^{-2}, ")"))
          # expression(atop(Log[10]~density, paste("anomaly (kg ", nmi^{-2}, ")")))
        ) +
        facet_wrap(~time, ncol = ifelse(length_based, 5, map_ncol-2)) +
        theme(
          legend.position = "bottom",
          axis.text = element_blank(),
          axis.title = element_blank(),
          axis.ticks = element_blank(),
          strip.background = element_blank(),
          panel.background = element_blank(),
          plot.background = element_blank(),
          legend.background = element_blank(),
          legend.box.background = element_blank()
        )
    }), ncol = 1
)
```

# Indices

```{r index-calculus}
#| warning: false
#| timeit: !expr if(recalculate_index) TRUE else NULL

if(recalculate_index) {
  
  pbapply::pblapply(seq_along(mod_pred), function(i) {
    print(i)
    out <- try({get_index(
      obj = mod_pred[[i]],
      bias_correct = bias_correct_index,
      area = (grid_side_size/1.852)^2) %>% # grid cell area in square nautical miles
        mutate(est = est/1e6,
               lwr = lwr/1e6,
               upr = upr/1e6,
               cv = sqrt(exp(se^2) - 1)) %>% # convert kg to kilotonnes
        mutate(length_group = names(mod_pred[i]), .before = 1) %>% 
        mutate(species = trimws(gsub("\\(.*\\]", "", length_group)), 
               .before = 1) %>% 
        mutate(
          species = case_when(
            species == "snabeluer" ~ "Beaked redfish", 
            species == "vanlig uer" ~ "Golden redfish"
          )
        ) %>% 
        mutate(length_group = gsub("snabeluer |vanlig uer ", "", length_group))
    })
    
    if(inherits(out, "try-error")) return(paste(names(mod_pred[i]), "crashed"))
    
    write_csv(
      out,
      file = 
        paste0("data/model_output/length_based/", model_name, "-", 
               gsub(" ", "_", names(mod_pred[i])), "-index-",
               switch(response_variable, 
                      "catch biomass" = "cw", 
                      "biomass density" = "bm", 
                      "catch abundance" = "cc", 
                      "abundance density" = "ab"), "-",
               gsub("\\(.*\\)", "", model_family_name),
               ".csv")
    )
    
    return(paste(names(mod_pred[i]), "saved to data/model_output/"))
    
  })
}

index <- pbapply::pblapply(names(fit), function(k) {
  
  file <- 
    paste0("data/model_output/length_based/", model_name, "-", 
           gsub(" ", "_", k), "-index-",
           switch(response_variable, 
                  "catch biomass" = "cw", 
                  "biomass density" = "bm", 
                  "catch abundance" = "cc", 
                  "abundance density" = "ab"), "-",
           gsub("\\(.*\\)", "", model_family_name),
           ".csv")
  
  if(!file.exists(file)) return(NULL)
  
  read_csv(file) 
}, cl = length(fit))

names(index) <- names(fit)
```

## Overall indices

```{r}
#| label: fig-index
#| fig-height: 8
#| fig-cap: !expr paste0("Survey index for ", sp_name, if_else(length_based, " length groups", ""), " estimated using spatiotemporal generalized linear mixed effects models with the sdmTMB package (solid line). ", value_text_name, " is used as the response variable", ifelse(offset, " with swept area as the offset argument.", "."), " The model has ", paste(terms, collapse = ", "), " as fixed effects, uses a mesh with ", fit[[1]]$mesh_n, " vertices (knots), ", model_family_name," model family, and ", fit[[1]]$mod$spatiotemporal, " spatiotemporal random fields. Predictions are made to a grid with ", proj_dt %>% distinct(X,Y) %>% nrow(), " cells with a side of ", grid_side_size, " km. Grey shading indicates 95% confidence intervals. Dots indicate years with survey data.")

# g = unique(trimws(gsub("\\(.*\\]", "", names(index))))[1]
tmp <- 
  lapply(unique(trimws(gsub("\\(.*\\]", "", names(index)))), function(g) {
    
    lapply(index[grep(g, names(index))], function(k) {
      if(any(!is.na(k$upr))) {
        if(max(k$upr, na.rm = TRUE) > 2*max(k$est, na.rm = TRUE)) {
          k[k$upr > 2*max(k$est, na.rm = TRUE),]$upr <- ceiling(2*max(k$est, na.rm = TRUE))
        }}
      k
    }) %>%
      bind_rows() 
  })

cowplot::plot_grid(
  plotlist = lapply(tmp, function(tmp) {
    tmp %>%
      ggplot(aes(year, est)) +
      {if(any(!is.na(tmp$lwr))) geom_ribbon(
        aes(ymin = lwr, ymax = upr), alpha = 0.4, color = NA,
        fill = "grey")}  +
      geom_path() +
      coord_cartesian(expand = FALSE) +
      scale_x_continuous(
        breaks = seq(1980, 2030, 6),
        labels = substr(seq(1980, 2030, 6),3,4)
      ) +
      facet_wrap(~length_group, scales = "free_y", ncol = 4) +
      expand_limits(y = 0) +
      labs(
        x = 'Year',
        y = ifelse(model_type == "Biomass", 'Biomass (kt)', 'Abundance (millions)'),
        title = unique(tmp$species)
      ) +
      theme_bw() 
  }), ncol = 1
)
```

```{r}
#| label: fig-index-2
#| fig-height: 6
#| fig-cap: Length based survey indices for beaked and golden redfish shown in an alternative way. The figure shows years from 1988.

cowplot::plot_grid(
  plotlist = 
    lapply(tmp, function(k) {
      k %>% 
        filter(year >= 1988) %>%  
        ggplot(aes(x = year, y = est)) + 
        geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.4, color = NA,
                    fill = "grey") +
        geom_path() +
        # expand_limits(y = 0) +
        labs(
          x = 'Year',
          y = ifelse(model_type == "Biomass", 'Biomass (kt)', 'Abundance (millions)')
        ) +
        scale_x_continuous(
          breaks = seq(1980, 2030, 4),
          labels = substr(seq(1980, 2030, 4),3,4)
        ) +
        facet_grid(length_group ~ species, scales = "free_y") +
        coord_cartesian(ylim = c(0, NA)) + 
        scale_y_continuous(
          # trans = "log1p",
          # limits = function(k) round(c(min(k)*0.98, max(k)*1.02), 1),
          breaks = function(k) {
            if(any(range(k) > 10)) {
              c(0, floor(k))
            } else {
              c(0, round(k, 1))
            }
          }
          # n.breaks = 3
        ) +
        theme(plot.margin = margin(0,0,0,0),
              panel.spacing = unit(-0.5, "cm"))
    })
)
```


```{r}
#| code-fold: true
#| label: fig-combined-index
#| fig-height: 10
#| fig-cap: !expr paste("Length based survey indices for", sp_name, "estimated using sdmTMB. Red solid line and grey shading indicates the total summed index and 95% confidence intervals (CIs), respectively. Colored fill indicates the average contribution of each length group included to the index. Dots indicate years with survey data. The figure assumes similar gear selectivity among all length groups.")

cowplot::plot_grid(
  plotlist = 
    lapply(tmp, function(k) {
      
      lg_cols <- setNames(pal_viridis()(length(unique(k$length_group))), 
                          unique(k$length_group))
      
      total_index <- k %>%
        group_by(year) %>%
        reframe(
          est = log(sum(est, na.rm = TRUE)*1e6),
          upr = est + 1.96*sqrt(sum(se^2, na.rm = TRUE)),
          lwr = est - 1.96*sqrt(sum(se^2, na.rm = TRUE))
        ) %>%
        mutate(across(est:lwr, function(k) exp(k)/1e6))
      
      ymax <-
        ifelse(all(is.na(total_index$upr)),
               ceiling(1.05*max(total_index$est, na.rm = TRUE)),
               ifelse(max(total_index$upr, na.rm = TRUE) >
                        1.2*max(total_index$est, na.rm = TRUE),
                      1.2*max(total_index$est, na.rm = TRUE),
                      ceiling(max(total_index$upr, na.rm = TRUE))
               )
        )
      
      ggplot() +
        {if(any(!is.na(total_index$upr)))
          geom_ribbon(
            data = total_index,
            aes(x = year, ymin = lwr, ymax = upr),
            alpha = 0.4, color = NA, fill = "grey")
        } +
        geom_area(
          data = k,
          aes(year, est, fill = length_group)) +
        # geom_path(
        #   data = k,
        #   aes(year, est, color = length_group)) +
        geom_line(data = total_index, aes(year, est, color = "Total"), color = "red") +
        geom_point(data = total_index, aes(year, est), shape = 1, color = "red") +
        coord_cartesian(
          expand = FALSE,
          ylim = c(0, ymax)
        ) +
        scale_x_continuous(
          breaks = seq(1980, 2030, 4),
          labels = substr(seq(1980, 2030, 4),3,4)
        ) +
        scale_fill_manual(values = lg_cols) +
        # scale_color_manual(values = "black") +
        # scale_color_manual(values = lg_cols) +
        labs(
          x = 'Year',
          y = ifelse(model_type == "Biomass", 'Biomass (kt)', 'Abundance (millions)'),
          fill = "Length group",
          # color = "Length group",
          title = unique(k$species)
        ) +
        theme_bw() +
        theme(legend.position = "bottom")
    }), 
  ncol = 1
)
```

```{r}
#| label: fig-lg-index-year
#| fig-height: 8
#| fig-cap: !expr paste0("Estimated total ", tolower(trimws(gsub("\\(.*\\)", "", model_type))), " of ", sp_name, " length groups (+ 95% confidence intervals) by year.")

cowplot::plot_grid(
  plotlist = 
    lapply(bind_rows(index) %>% pull(species) %>% unique(), function(g) {
      index %>%
        bind_rows() %>% 
        filter(species == g) %>% 
        ggplot(aes(x = length_group, y = est, ymin = est, ymax = upr)) +
        geom_col(color = "black", linewidth = 0.3, position = "dodge",
                 fill = "grey") +
        geom_errorbar(width = 0.2, linewidth = 0.3) +
        facet_wrap(~year, ncol = 5, dir = "v") +
        labs(y = ifelse(
          model_type == "Biomass", 'Biomass (kt)', 'Abundance (millions)'),
          x = "Length group (cm)", title = g) +
        scale_x_discrete(guide = guide_axis(angle = 90)) +
        coord_cartesian(expand = FALSE, ylim = c(0, ifelse(g == "Beaked redfish", 1000, 100))) +
        theme(
          legend.position = "bottom",
          strip.background = element_blank(),
          strip.text = element_text(size = 10, face = 2)
        )
    })
)
```

```{r}
#| label: fig-lg-average
#| fig-cap: !expr paste0(model_type, " of ", sp_name, " length groups (+ 95% confidence intervals) averaged over years.")

cowplot::plot_grid(
  plotlist = 
    lapply(bind_rows(index) %>% pull(species) %>% unique(), function(g) {
      index %>% 
        bind_rows() %>% 
        filter(species == g) %>% 
        group_by(length_group) %>% 
        reframe(mean = mean(est), se = sd(est)/sqrt(n())) %>% 
        ggplot(aes(x = length_group, y = mean, ymin = mean, ymax = mean + 1.96*se)) + 
        geom_col(color = "black", linewidth = 0.3, position = "dodge",
                 fill = "grey") +
        geom_errorbar(width = 0.2, linewidth = 0.3) +
        labs(y = ifelse(
          model_type == "Biomass", 'Biomass (kt)', 'Abundance (millions)'),
          x = "Length group (cm)", title = g) +
        scale_x_discrete(guide = guide_axis(angle = 90)) +
        coord_cartesian(expand = FALSE) +
        theme(
          strip.background = element_blank(),
          strip.text = element_text(size = 10, face = 2)
        )
    })
)
```

## Regional trends

```{r regional-indices}
#| timeit: !expr if(recalculate_index) TRUE else NULL
#| results: hide

if(run_regional) {
  if(recalculate_index) {
    
    regions <- st_read("data/gis/regions.gpkg", quiet = TRUE) %>% 
      st_transform(dt_crs) %>% 
      mutate(region = gsub(" \\(.*\\)", "", region))
    
    reg_proj_dt <- proj_dt %>% 
      mutate(x = 1000*X, y = 1000*Y) %>% 
      sf::st_as_sf(coords = c("x", "y"), crs = dt_crs, remove = FALSE) %>% 
      sf::st_join(regions) %>% 
      sf::st_drop_geometry() %>% 
      filter(!is.na(region)) 
    
    ## Index calculus (takes a long time)
    regional_indices <- 
      pbapply::pblapply(seq_along(fit), function(i) {
        
        lgi_out <- pbapply::pblapply(unique(reg_proj_dt$region), function(j) {
          
          pred_obj <- predict(
            fit[[i]]$mod,
            newdata = reg_proj_dt %>% 
              filter(region == j),
            type = "response",
            model = NA,
            return_tmb_object = TRUE)
          
          out <- try({get_index(
            obj = pred_obj,
            bias_correct = bias_correct_index,
            area = (grid_side_size/1.852)^2) %>% # grid cell area in square nautical miles
              mutate(est = est/1e6, # convert kg to kilotonnes
                     lwr = lwr/1e6,
                     upr = upr/1e6, 
                     cv = sqrt(exp(se^2) - 1)
              ) %>% 
              mutate(length_group = names(mod_pred[i]),
                     region = j, .before = 1) %>% 
              mutate(species = trimws(gsub("\\(.*\\]", "", length_group)),
                     .before = 1) %>% 
              mutate(
                species = case_when(
                  species == "snabeluer" ~ "Beaked redfish", 
                  species == "vanlig uer" ~ "Golden redfish"
                )
              ) %>% 
              mutate(length_group = gsub("snabeluer |vanlig uer ", "", length_group))
          })
          
          if(inherits(out, "try-error")) return(paste(names(mod_pred[i]), "crashed"))
          
          # write_csv(
          #   out,
          #   file = 
          #     paste0("data/model_output/length_based/", model_name, "-", 
          #            gsub(" ", "_", names(mod_pred[i])), "-regional-index-",
          #            switch(response_variable, 
          #                   "catch biomass" = "cw", 
          #                   "biomass density" = "bm", 
          #                   "catch abundance" = "cc", 
          #                   "abundance density" = "ab"), "-",
          #            gsub("\\(.*\\)", "", model_family_name),
          #            ".csv")
          # )
          # 
          # return(paste(names(mod_pred[i]), "saved to data/model_output/"))
          
          return(out)
        }
        ) %>% 
          bind_rows()
        
        return(lgi_out)
      }, cl = 2
      ) %>% 
      bind_rows() %>% 
      # filter(!is.na(region)) %>% 
      mutate(
        region = factor(
          region, 
          levels = c("North Sea", "Norwegian Coast S", "Norwegian Coast N", 
                     "Western Slope", "Svalbard Slope", "Northern Barents Sea", 
                     "Central Barents Sea", "Eastern Barents Sea")
        )
      )
    
    write_csv(
      regional_indices,
      file =
        paste0("data/model_output/length_based/", model_name,
               "-regional-indices-",
               switch(response_variable,
                      "catch biomass" = "cw",
                      "biomass density" = "bm",
                      "catch abundance" = "cc",
                      "abundance density" = "ab"), "-",
               gsub("\\(.*\\)", "", model_family_name),
               ".csv")
    )
  } else {
    regional_indices <- 
      read_csv(
        paste0("data/model_output/length_based/", model_name,
               "-regional-indices-",
               switch(response_variable,
                      "catch biomass" = "cw",
                      "biomass density" = "bm",
                      "catch abundance" = "cc",
                      "abundance density" = "ab"), "-",
               gsub("\\(.*\\)", "", model_family_name),
               ".csv")
      )
  }
}
```

```{r}
#| label: fig-region-length-index
#| fig-height: !expr length(fit)*1.3
#| fig-cap: Regional indices for redfish by species and length group (black lines). The y-axis scale differs among vertical panels. Grey shading indicates 95% confidence intervals and has been limited to 1.2 times the maximum index value for each species.

tmp <- 
  regional_indices %>% 
  mutate(lg = paste(sapply(strsplit(species, " "), "[", 1), length_group)) %>% 
  group_by(lg) %>% 
  mutate(upr = if_else(upr > 1.2*max(est), 1.2*max(est), upr)) %>% 
  ungroup()

levels(tmp$region) <- sub(" ", "\n", levels(tmp$region))

ggplot(tmp, aes(x = year, y = est)) +
  {if(any(!is.na(tmp$lwr))) geom_ribbon(
    aes(ymin = lwr, ymax = upr), alpha = 0.4, color = NA, 
    fill = "grey")}  +
  geom_path() +
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(
    breaks = seq(1980, 2030, 5), 
    labels = substr(seq(1980, 2030, 5),3,4)
  ) +
  facet_grid(lg~region, scales = "free_y") + 
  coord_cartesian(ylim = c(0, NA), expand = FALSE) +
  labs(
    x = 'Year', 
    y = ifelse(model_type == "Biomass", 'Biomass (kt)', 'Abundance (millions)')
  ) +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.grid.minor = element_blank())

```

```{r}
#| label: fig-region-index
#| fig-cap: !expr paste0("Trends for ", sp_name, " length groups allocated to regions. The figure assumes similar gear selectivity among all length groups. The figure shows years from 1988.")
#| fig-height: 14

if(run_regional) {
  
  cowplot::plot_grid(
    plotlist = 
      lapply(unique(regional_indices$species), function(k) {
        
        tmp <- 
          regional_indices %>% 
          filter(species == k) %>% 
          filter(year >= 1988)
        
        lg_cols <- setNames(pal_viridis()(length(unique(tmp$length_group))), 
                            unique(tmp$length_group))
        
        tmp %>%
          group_by(length_group, year, region) %>%
          reframe(est = sum(est)) %>%
          ggplot(aes(x = year, y = est, fill = length_group)) +
          geom_area(alpha = 0.7) +
          coord_cartesian(expand = FALSE) +
          scale_x_continuous(
            breaks = seq(1980, 2030, 5),
            labels = substr(seq(1980, 2030, 5),3,4)
          ) +
          facet_wrap(~region) +
          scale_fill_manual(values = lg_cols) +
          labs(
            x = 'Year', fill = "Length group (cm)",
            y = ifelse(model_type == "Biomass", 'Biomass (kt)', 'Abundance (millions)'),
            title = k
          ) +
          theme_bw() +
          theme(legend.position = "bottom")
      }),
    ncol = 1
  )
}
```

```{r}
#| label: fig-region-response-distribution
#| fig-cap: !expr paste(model_type, "distribution among regions", ifelse(length_based, "for each length group", ""), "averaged over years.")
#| fig-height: 12

if(run_regional) {
  
  cowplot::plot_grid(
    plotlist = 
      lapply(unique(regional_indices$species), function(k) {
        
        tmp <- 
          regional_indices %>% 
          filter(species == k) %>% 
          group_by(., length_group, region, year) %>%
          reframe(est = sum(est)) %>% 
          group_by(., length_group, region) %>% 
          reframe(mean = mean(est), sd = sd(est), n = n()) %>% 
          group_by(., length_group) %>% 
          mutate(sum = sum(mean)) %>% 
          mutate(prop = mean/sum, prop_sd = sd/sum) %>% 
          ungroup() %>% 
          mutate(prop_upr = prop + 1.96*prop_sd/sqrt(n)) %>% 
          mutate(prop_upr = if_else(prop_upr > 1, 1, prop_upr)) %>% 
          mutate(region = sapply(region, function(k) {
            paste(strwrap(k, width = 16), collapse = "\n")
          }))
        
        
        tmp %>% 
          ggplot(aes(x = region, y = prop, ymin = prop, 
                     ymax = prop_upr)) +
          geom_col(position = position_dodge(1), color = "black",
                   linewidth = LS(0.5), alpha = 0.5) +
          geom_errorbar(position = position_dodge(1), width = 0.2,
                        linewidth = LS(0.5)) +
          facet_wrap(~length_group, ncol = 4) +
          labs(x = "Region", y = paste("Proportion of total", tolower(model_type)),
               title = k) +
          guides(
            fill = guide_legend(
              keywidth=0.2, keyheight=0.2, default.unit="inch")
          ) +
          scale_y_continuous(expand = c(0,0)) +
          scale_x_discrete(guide = guide_axis(angle = 90)) +
          expand_limits(y = 0.5) +
          theme(legend.position = "right")
      }), ncol = 1
  )
}
```