---
title: "IMR data overview for compiled redfish survey index work"
subtitle: "NMD (BioticExplorer) database"
author: |
  Mikko Vihtakari <mikko.vihtakari@hi.no>
  Institute of Marine Research, Norway (IMR)
date: "Version: `r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    theme: lumen
    toc_float: true
    toc_depth: 3
    collapsed: true
    variant: markdown+simple_table
    fig_caption: true
    number_sections: true
    code_folding: hide
header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
 \usepackage[T1]{fontenc}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.pos = "H",
  fig.width = 9,
  fig.height = 5,
  tab.pos = "H",
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

options(knitr.kable.NA = "")
```

This document contains the scripts and overview of data extracted from the IMR database using BioticExplorerServer database set up locally on a computer. 

# Data export

Use this switch to update data. Data update will only work on Mikko's computers:

```{r definitions, class.source = 'fold-show', cache = FALSE}
update_data <- TRUE
filter_sp <- c("snabeluer", "vanlig uer", "uerslekten")
sp_name <- "redfish"
Sp_name <- "Redfish"
sp_names <- c("beaked redfish", "golden redfish", "uerslekten")
```

Load required packages:

```{r packages, cache=FALSE}
## Load packages

packages <- c("sf", "tidyverse", "ggOceanMaps", "ggspatial")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, function(x) {
  suppressPackageStartupMessages(library(x, character.only = TRUE))
}))

rm(packages, installed_packages)

## ---------------------------

## ggplot theme

theme_cust <- theme_classic(base_size = 8) %+replace%
  theme(
    strip.background = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.box.background = element_blank(),
    plot.margin = margin(5.5, 10, 5.5, 5.5)
  )

theme_set(theme_cust) # Set default theme globally for the entire project

## Custom functions

source("R/make_table.R")
source("R/bubble_map.R")

## Flatten list of ggplots
## https://stackoverflow.com/questions/55731855/how-to-extract-every-ggplot2-plot-from-a-nested-list

flattenlist <- function(x) {
  morelists <- sapply(x, function(xprime) {
    'list' %in% class(xprime) & !('gg' %in% class(xprime))
  })
  out <- c(x[!morelists], unlist(x[morelists], recursive = FALSE))
  if (sum(morelists)) {
    Recall(out)
  } else {
    return(out)
  }
}

## For printing of the ggFishPlots results

htmlcat <- function(text) {
  cat(gsub(pattern = "\n", replacement = "  \n", x = text))
}
```

## Connect to the database

This stuff works only on computers with local access to the BioticExplorerServer database:

```{r bes-database}
if (update_data) {
  library(duckdb)

  ## Connect to the database

  bes_database_path <- Sys.getenv("BES_DATABASE_PATH", "paste your BES database path here")

  con_db <- bes_database_path %>%
    normalizePath() %>%
    duckdb::duckdb(read_only = TRUE) %>%
    DBI::dbConnect()

  ## Create the data objects

  stnall <- dplyr::tbl(con_db, "stnall")
  indall <- dplyr::tbl(con_db, "indall")
  gearlist <- dplyr::tbl(con_db, "gearindex") %>% collect()
  mission <- dplyr::tbl(con_db, "mission")
  updated <- dplyr::tbl(con_db, "metadata") %>%
    dplyr::pull(timestart) %>%
    as.Date()

  # Save updated date

  writeLines(
    as.character(updated),
    "data/survey data/biotic_database_updated.txt"
  )

  ## The cruises to be read

  latest_file <- dir(
    "data/survey data",
    pattern = "redfish bottom trawl cruises"
  )
  latest_file <- latest_file[which.max(as.Date(sapply(
    strsplit(latest_file, " "),
    "[",
    1
  )))]

  cs_list <- read_csv(
    file.path("data/survey data", latest_file),
    show_col_types = FALSE
  ) %>%
    filter(!is.na(cruiseseries)) %>%
    mutate(cruise = as.character(cruise))
}
```

# Station level data

The script under calculates abundance and biomass swept area density assuming 25 m effective width for shrimp trawls (Campelen) and `r round(25*3.746)` m width for cod trawls (Alfredo). This trawl width comes from Greenland halibut work and needs to be adjusted for `r sp_name`.

```{r stn-data, cache = FALSE}
if (update_data) {
  # Extract relevant data
  stn <- stnall %>%
    filter(
      gearcategory == "Bottom trawls",
      missionid %in% local(cs_list$missionid),
      !is.na(distance),
      !is.na(latitudestart),
      !is.na(longitudestart)
    ) %>%
    collect() %>%
    select(where(~ !all(is.na(.x)))) %>%
    filter(
      startyear > 1980,
      startyear < lubridate::year(Sys.Date()), # filters out the current year by default
      !is.na(catchweight),
      # cruiseseriescode != "NA",
      gearcategory %in% c("Bottom trawls"),
      missiontypename %in% c("Forskningsfartøy", "Leiefartøy"),
      gearcondition %in% 1:2,
      latitudestart > 0,
      !stationtype %in% c("C", "2", "A", "E"),
      !is.na(distance),
      !nation %in% c("140", "90"), # Remove Russian data
      !gear %in% c(3136, 3134), # remove 3 3136 Bomtrål, 120 3134 Singeltrål
      distance <= 5,
      distance >= 0.3
    ) %>%
    filter(
      if_else(
        cruiseseriescode == "23",
        samplequality %in% 1:2, # add targetted seithe surveys which used acoustics to select stations, but not on redfish
        samplequality %in% 1
      )
    ) %>%
    mutate(id = paste(cruise, platformname, serialnumber, sep = "_"))

  # Fix the naming

  stn[
    stn$commonname %in% c("uerslekten", "uerfamilien"),
  ]$commonname <- "uerslekten"

  # Remove stations with unclarities in catchcount / weight
  tmp <- stn %>%
    filter(commonname %in% filter_sp) %>%
    group_by(cruiseseriescode, id, commonname) %>%
    reframe(
      catchweight = sum(catchweight, na.rm = TRUE),
      catchcount = sum(catchcount, na.rm = TRUE),
      lengthsampleweight = sum(lengthsampleweight, na.rm = TRUE),
      lengthsamplecount = sum(lengthsamplecount, na.rm = TRUE)
    ) %>%
    filter(
      catchcount < lengthsamplecount | catchweight < lengthsampleweight
    ) %>%
    pull(id)

  stn <- stn %>%
    filter(!id %in% tmp)

  # General manipulation

  stn$commonname[!stn$commonname %in% filter_sp] <- "other"

  stn <- stn %>%
    group_by(
      startyear,
      cruise,
      callsignal,
      platformname,
      missionid,
      stationstartdate,
      stationstopdate,
      bottomdepthstart,
      gear,
      gearname,
      trawldoorspread,
      verticaltrawlopening,
      wirelength,
      vesselspeed,
      serialnumber,
      longitudestart,
      latitudestart,
      distance,
      commonname
    ) %>%
    reframe(
      biomass = sum(catchweight, na.rm = TRUE),
      count = sum(catchcount, na.rm = TRUE)
    ) %>%
    mutate(commonname = factor(commonname)) %>%
    group_by(
      startyear,
      cruise,
      callsignal,
      platformname,
      missionid,
      stationstartdate,
      stationstopdate,
      bottomdepthstart,
      gear,
      gearname,
      trawldoorspread,
      verticaltrawlopening,
      wirelength,
      vesselspeed,
      serialnumber,
      longitudestart,
      latitudestart,
      distance
    ) %>%
    complete(
      commonname,
      fill = list(biomass = 0, count = 0),
      explicit = FALSE
    ) %>%
    filter(commonname != "other") %>%
    ungroup()
  # filter(!(biomass != 0 & count == 0)) # Inconsistent records

  ## Replace 0 counts with positive biomasses by NAs

  if (any(stn$biomass == 0 & stn$count != 0)) {
    stn[stn$biomass == 0 & stn$count != 0, ]$biomass <- NA
  }

  if (any(stn$biomass != 0 & stn$count == 0)) {
    stn[stn$biomass != 0 & stn$count == 0, ]$count <- NA
  }

  if (any(stn$biomass > 2e4)) {
    message(
      sum(stn$biomass > 2e4, na.rm = TRUE),
      " stations with over 20k catches removed."
    )

    stn[stn$biomass > 2e4 & !is.na(stn$biomass), ]$biomass <- NA
  }

  # Species

  stn <- droplevels(stn)
  stn <- stn %>%
    mutate(commonname = factor(as.character(commonname), levels = filter_sp))

  # Remove stations outside the study area

  roi <- sf::read_sf("data/gis/study area.gpkg")

  tmp <- stn %>%
    st_as_sf(
      coords = c("longitudestart", "latitudestart"),
      crs = 4326
    ) %>%
    st_intersects(roi) %>%
    as.integer() %>%
    is.na() %>%
    `!`

  stn <- stn[tmp, ]

  ## Remove unreasonably large count values from the dataset

  tmp <- stn$biomass / stn$count

  if (min(tmp, na.rm = TRUE) < 0.001) {
    # stop("Smallest average individual weights less than a gram. Something wrong with the count column?")
    message(
      sum(tmp < 0.001, na.rm = TRUE),
      " stations removed because average individual weight for each was less than a gram."
    )

    stn[tmp < 0.001 & !is.na(tmp), ]$count <- NA
  }

  if (max(tmp, na.rm = TRUE) > 10) {
    # stop("Largest average individual weights over 10 kg. Something wrong with the count column?")
    message(
      sum(tmp > 10, na.rm = TRUE),
      " stations removed because average individual weight for each was over 10 kg."
    )

    stn[tmp > 10 & !is.na(tmp), ]$count <- NA
  }

  ## Date format
  tmp <- strptime(
    stn$stationstartdate,
    format = "%Y-%m-%d %H:%M:%S",
    tz = "GMT"
  )
  stn$stationstartdate <-
    if_else(
      is.na(tmp),
      strptime(
        paste(stn$stationstartdate, "00:00:01"),
        format = "%Y-%m-%d %H:%M:%S",
        tz = "GMT"
      ),
      tmp
    )

  tmp <- strptime(stn$stationstopdate, format = "%Y-%m-%d %H:%M:%S", tz = "GMT")
  stn$stationstopdate <-
    if_else(
      is.na(tmp),
      strptime(
        paste(stn$stationstopdate, "00:00:01"),
        format = "%Y-%m-%d %H:%M:%S",
        tz = "GMT"
      ),
      tmp
    )

  ## Replace missing start dates with stop dates

  stn[
    is.na(stn$stationstartdate) & !is.na(stn$stationstopdate),
    "stationstartdate"
  ] <-
    stn[
      is.na(stn$stationstartdate) & !is.na(stn$stationstopdate),
      "stationstopdate"
    ]

  ## Trawling duration

  tmp <- difftime(stn$stationstopdate, stn$stationstartdate, units = 'mins')
  tmp[tmp <= 0 | tmp > 400] <- NA
  stn$duration <- tmp

  ### Assuming that >5 & <1.2 knot trawling speeds are errors in *duration* registration (distance not changed)

  tmp <- stn$distance / (as.numeric(stn$duration) / 60)
  stn[tmp > 5 & !is.na(tmp), ]$duration <- NA
  stn[tmp < 1.2 & !is.na(tmp), ]$duration <- NA

  ## Cruise series name

  stn <- stn %>%
    left_join(
      cs_list %>%
        select(
          startyear,
          cruise,
          callsignal,
          platformname,
          missionid,
          cruiseseries
        ),
      by = join_by(startyear, platformname, cruise, callsignal, missionid)
    ) %>%
    relocate(cruiseseries, .after = startyear)

  ### Remove EggaN 1994-1996 as redfish was not recorded correctly these years

  stn <- stn %>% filter(!(cruiseseries == "EggaN" & startyear < 1997))

  ## Gears
  stn <- stn %>%
    ungroup() %>%
    mutate(
      # gear = if_else(cruiseseries == "EggaN" & gear == "3181", "3178", gear), # what is this?
      gear = if_else(cruiseseries == "EggaN" & gear == "3118", "3181", gear),
      gear = if_else(
        cruiseseries == "EggaN" & (gear == "3100" | gear == "3110"),
        "3181",
        gear
      ),
      gear = if_else((gear == "3100"), "3271", gear),
      gear = if_else(gear == "3292", "3270", gear),
      gear = if_else(gear == "3293", "3271", gear)
    ) %>%
    dplyr::select(-gearname) %>%
    left_join(
      gearlist %>% dplyr::select(code, gearname) %>% rename("gear" = "code"),
      by = join_by(gear)
    ) %>%
    relocate(gearname, .after = "gear") %>%
    mutate(
      geartype = if_else(
        grepl("GOV-", gearname),
        "GOV",
        if_else(
          grepl("^32", gear),
          "Shrimp",
          if_else(grepl("^31", gear), "Cod", "Other")
        )
      )
    ) %>%
    mutate(
      gear_width = if_else(
        geartype == "GOV",
        25,
        if_else(
          geartype == "Shrimp",
          25,
          if_else(geartype == "Cod", 25 * 3.746, NA)
        )
      )
    ) %>%
    mutate(
      rho_biomass = biomass / (distance * (gear_width / 1852)),
      rho_abundance = count / (distance * (gear_width / 1852))
    ) %>%
    mutate(
      tmp = wirelength / trawldoorspread,
      trawldoorspread = ifelse(tmp < 2.4 | tmp > 40, NA, trawldoorspread),
      verticaltrawlopening = ifelse(
        verticaltrawlopening < 2 | verticaltrawlopening > 7,
        NA,
        verticaltrawlopening
      ),
      duration = ifelse(duration > 300, NA, duration),
    ) %>%
    dplyr::select(-tmp)

  ## Remove observations on land
  # stn <- ggOceanMaps::dist2land(stn, binary = TRUE, dist.col = "in_ocean",
  #                               bind = TRUE) %>%
  #   filter(in_ocean) %>%
  #   dplyr::select(-in_ocean)

  ## Fill in missing depths (and correct too deep stations)
  stn[
    !is.na(stn$bottomdepthstart) & stn$bottomdepthstart > 1400,
    "bottomdepthstart"
  ] <- NA

  stn[is.na(stn$bottomdepthstart), "bottomdepthstart"] <-
    ggOceanMaps::get_depth(
      stn[is.na(stn$bottomdepthstart), ],
      bathy.style = "rc",
      bind = FALSE
    )

  stn <- stn %>% filter(!is.na(bottomdepthstart))

  stn[stn$bottomdepthstart == 0, "bottomdepthstart"] <-
    ggOceanMaps::get_depth(
      stn[stn$bottomdepthstart == 0, ],
      bathy.style = "rc",
      bind = FALSE
    )

  gearList <- read_rds('data/definitions/imr_gearlist.rds')

  # stn %>%
  #   ungroup() %>%
  #   dplyr::select(longitudestart, latitudestart, startyear, cruiseseries) %>%
  #   distinct() %>%
  #   st_as_sf(coords = c("longitudestart", "latitudestart"), crs = 4326) %>%
  #   sf::write_sf("data/gis/station_positions.gpkg")

  stn <- ungroup(stn)

  ### Save ####

  write_rds(stn, 'data/survey data/stn_Database.rds', compress = "xz")
} else {
  stn <- read_rds('data/survey data/stn_Database.rds')
  gearList <- read_rds('data/definitions/imr_gearlist.rds')
  updated <- readLines("data/survey data/biotic_database_updated.txt") %>%
    as.Date()
}
```

## Station overview

The dataset contained `r stn %>% distinct(startyear, callsignal, serialnumber) %>% nrow()` trawl hauls from `r min(stn$startyear)` to `r max(stn$startyear)`. `r stn %>% filter(commonname == "snabeluer", rho_biomass > 0) %>% distinct(startyear, callsignal, serialnumber) %>% nrow()` of these stations contained beaked redfish, `r stn %>% filter(commonname == "vanlig uer", rho_biomass > 0) %>% distinct(startyear, callsignal, serialnumber) %>% nrow()` golden redfish, and `r stn %>% filter(commonname == "uerslekten", rho_biomass > 0) %>% distinct(startyear, callsignal, serialnumber) %>% nrow()` "uerslekten" (originally called "uerslekten", and "uerfamilien"). `r stn$stationstartdate %>% max` is the latest record in the dataset. The database was updated on `r format(updated, "%d %B, %Y")`.

```{r, fig.cap = paste("Overview of station positions over all years. Purple crosses indicate stations with no", sp_name, "occurrences."), fig.height = 7}
p <-
  basemap(stn, bathy.style = "rbg", legends = FALSE) +
  geom_spatial_point(
    data = stn %>% filter(rho_biomass == 0),
    aes(x = longitudestart, y = latitudestart),
    color = "purple",
    shape = "+",
    alpha = 0.5
  ) +
  geom_spatial_point(
    data = stn %>% filter(rho_biomass > 0 & !is.na(rho_biomass)),
    aes(
      x = longitudestart,
      y = latitudestart,
      size = rho_biomass,
      color = geartype
    ),
    shape = 21
  ) +
  facet_grid(commonname ~ cruiseseries) +
  scale_size_area() +
  labs(
    color = "Gear type",
    size = expression(paste("Biomass density (kg ", nmi^-2, ")"))
  ) +
  theme(legend.position = "bottom", axis.text = element_blank())

p
```

```{r, fig.cap = paste("Number of stations with and without", sp_name, "by year and survey series.")}
stn %>%
  group_by(cruiseseries, startyear, commonname) %>%
  reframe(
    yes = sum(rho_biomass > 0, na.rm = TRUE),
    no = sum(rho_biomass == 0, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = c(yes, no)) %>%
  mutate(name = factor(name, c("no", "yes"))) %>%
  ggplot(aes(x = startyear, y = value, fill = name)) +
  geom_col() +
  labs(x = "Year", y = "Number of stations", fill = "Contains\ntarget taxa") +
  facet_grid(commonname ~ cruiseseries) +
  scale_fill_manual(values = c("yes" = "#66a182", "no" = "grey")) +
  scale_y_continuous(expand = c(0.01, 0)) +
  scale_x_continuous(
    breaks = seq(1980, 2030, 4),
    labels = substr(seq(1980, 2030, 4), 3, 4),
    expand = c(0.01, 0)
  ) +
  theme(legend.position = "bottom")
```

```{r, fig.cap = paste("Average number of stations per year with and without", sp_name, "by survey series.")}
stn %>%
  group_by(cruiseseries, startyear, commonname) %>%
  reframe(
    yes = sum(rho_biomass > 0, na.rm = TRUE),
    no = sum(rho_biomass == 0, na.rm = TRUE)
  ) %>%
  group_by(cruiseseries, commonname) %>%
  reframe(yes = mean(yes, na.rm = TRUE), no = mean(no, na.rm = TRUE)) %>%
  pivot_longer(cols = c(yes, no)) %>%
  mutate(name = factor(name, c("no", "yes"))) %>%
  ggplot(aes(x = cruiseseries, y = value, fill = name)) +
  geom_col() +
  labs(
    x = "Year",
    y = "Average number of stations per year",
    fill = "Contains\nthe taxa"
  ) +
  facet_wrap(~commonname) +
  scale_fill_manual(values = c("yes" = "#66a182", "no" = "grey")) +
  scale_y_continuous(expand = c(0, 0)) +
  # scale_x_continuous(expand = c(0.01,0)) +
  theme(legend.position = "bottom")
```

```{r, fig.cap = paste("Number of stations with and without", sp_name, "by year.")}
stn %>%
  group_by(startyear, commonname) %>%
  reframe(
    yes = sum(rho_biomass > 0, na.rm = TRUE),
    no = sum(rho_biomass == 0, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = c(yes, no)) %>%
  mutate(name = factor(name, c("no", "yes"))) %>%
  ggplot(aes(x = startyear, y = value, fill = name)) +
  geom_col() +
  labs(x = "Year", y = "Number of stations", fill = "Contains\ntarget taxa") +
  facet_wrap(~commonname) +
  scale_fill_manual(values = c("yes" = "#66a182", "no" = "grey")) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(
    breaks = seq(1980, 2030, 4),
    labels = substr(seq(1980, 2030, 4), 3, 4),
    expand = c(0.01, 0)
  ) +
  theme(legend.position = "bottom")
```

```{r, fig.cap = paste("Total annual biomass and number as well as average abundance and biomass densities of", sp_name, "by year and cruise-series."), fig.height=6}
stn %>%
  group_by(cruiseseries, startyear, commonname) %>%
  reframe(
    biomass = sum(biomass, na.rm = TRUE),
    count = sum(count, na.rm = TRUE),
    rho_biomass = mean(rho_biomass, na.rm = TRUE),
    rho_abundance = mean(rho_abundance, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = c(biomass, count, rho_biomass, rho_abundance)) %>%
  group_by(cruiseseries, startyear, name) %>%
  mutate(total = sum(value, na.rm = TRUE), prop = value / total) %>%
  # filter(name == "rho_biomass") %>%
  ggplot(aes(x = startyear, y = value, fill = commonname)) +
  geom_col() +
  facet_grid(name ~ cruiseseries, scales = "free_y") +
  theme(legend.position = "bottom")
```

```{r, fig.cap = paste("Abundance and biomass proportions of", sp_name, " in survey catch by year and cruise-series.")}
stn %>%
  group_by(cruiseseries, startyear, commonname) %>%
  reframe(
    biomass = sum(biomass, na.rm = TRUE),
    count = sum(count, na.rm = TRUE),
    rho_biomass = mean(rho_biomass, na.rm = TRUE),
    rho_abundance = mean(rho_abundance, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = c(biomass, count, rho_biomass, rho_abundance)) %>%
  group_by(cruiseseries, startyear, name) %>%
  mutate(total = sum(value), prop = value / total) %>%
  filter(name %in% c("rho_abundance", "rho_biomass")) %>%
  ggplot(aes(x = startyear, y = prop, fill = commonname)) +
  geom_col() +
  facet_grid(name ~ cruiseseries, scales = "free_y")
```

```{r, fig.cap = paste("Annual average biomass and number as well as abundance and biomass densities of", sp_name, "by cruise-series.")}
stn %>%
  group_by(cruiseseries, startyear, commonname) %>%
  reframe(
    biomass = sum(biomass, na.rm = TRUE),
    count = sum(count, na.rm = TRUE),
    rho_biomass = mean(rho_biomass, na.rm = TRUE),
    rho_abundance = mean(rho_abundance, na.rm = TRUE)
  ) %>%
  group_by(cruiseseries, commonname) %>%
  reframe(across(c(biomass, count, rho_biomass, rho_abundance), mean)) %>%
  pivot_longer(cols = c(biomass, count, rho_biomass, rho_abundance)) %>%
  group_by(cruiseseries, name) %>%
  mutate(total = sum(value, na.rm = TRUE), prop = value / total) %>%
  # filter(name == "rho_biomass") %>%
  ggplot(aes(x = cruiseseries, y = value, fill = commonname)) +
  geom_col() +
  facet_wrap(~name, scales = "free_y")
```

## Gear

```{r, fig.cap = "Overview of used bottom trawls. Reketrål means Campelen-like shrimp trawls and other names Alfredo-like cod trawls. Number in parenthesis after the name indicates number of occurrences."}
ggplot(
  stn %>%
    # filter(rho_biomass > 0) %>%
    left_join(stn %>% group_by(gearname) %>% count) %>%
    mutate(gearname = paste0(gearname, " (", n, ")")),
  aes(x = gearname, y = log10(rho_biomass + 1), color = cruiseseries)
) +
  geom_boxplot() +
  theme_bw() +
  facet_grid(commonname ~ geartype, scales = "free") +
  labs(color = "Survey series") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

```{r gear-tab}
colfun <- function(x) {
  mn <- mean(x, na.rm = TRUE)
  ifelse(
    is.na(mn),
    "",
    paste(
      round(mn, 1),
      "\u00B1",
      paste(round(range(x, na.rm = TRUE), 1), collapse = "-")
    )
  )
}

stn %>%
  group_by(cruiseseries, gear) %>%
  mutate(
    verticaltrawlopening = ifelse(
      verticaltrawlopening < 0,
      NA,
      verticaltrawlopening
    )
  ) %>%
  reframe(
    n = n(),
    `Door spread (m)` = colfun(trawldoorspread),
    `Trawl opening height (m)` = colfun(verticaltrawlopening),
    `Trawling speed (kn)` = colfun(vesselspeed),
    `Tow distance (nmi)` = colfun(distance),
    `Tow duration (min)` = colfun(duration),
    `Beaked redfish density (kg/nmi2)` = colfun(rho_biomass[
      commonname == "snabeluer"
    ]),
    `Golden redfish density (kg/nmi2)` = colfun(rho_biomass[
      commonname == "vanlig uer"
    ])
  ) %>%
  left_join(
    gearList
  ) %>%
  #dplyr::select(-gear) %>%
  dplyr::relocate(gearname, .after = "cruiseseries") %>%
  rename(
    "Survey series" = "cruiseseries",
    "Gear name" = "gearname",
    "Gear code" = "gear",
    "Description" = "description"
  ) %>%
  make_table(
    caption = "Overview of trawls used in the included survey time-series."
  )
```

```{r, fig.cap = "Trawl statistics by survey time-series.", fig.height=8}
stn %>%
  distinct(
    startyear,
    serialnumber,
    geartype,
    duration,
    distance,
    cruise,
    cruiseseries,
    trawldoorspread,
    verticaltrawlopening,
    wirelength,
    bottomdepthstart
  ) %>%
  mutate(duration = as.numeric(duration)) %>%
  mutate(
    speed = distance / (as.numeric(duration) / 60),
    wirelengthratio = wirelength / bottomdepthstart
  ) %>%
  dplyr::select(
    cruiseseries,
    distance,
    duration,
    speed,
    trawldoorspread,
    verticaltrawlopening,
    wirelengthratio
  ) %>%
  pivot_longer(
    cols = c(
      distance,
      duration,
      speed,
      trawldoorspread,
      verticaltrawlopening,
      wirelengthratio
    ),
    values_drop_na = TRUE
  ) %>%
  mutate(
    name = recode(
      name,
      "distance" = "Trawled distance (nmi)",
      "duration" = "Trawl duration (min)",
      "speed" = "Trawl speed (knots)",
      "trawldoorspread" = "Trawl door spread (m)",
      "verticaltrawlopening" = "Vertical trawl opening (m)",
      "wirelengthratio" = "Trawl wire length / bottom depth"
    )
  ) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  labs(x = "Value", y = "Count") +
  facet_grid(cruiseseries ~ name, scales = "free") +
  theme_bw(base_size = 10)
```

```{r mean-gear-tab}
colfun <- function(x) {
  mn <- mean(x, na.rm = TRUE)
  ifelse(
    is.na(mn),
    "",
    paste(
      round(mn, 1),
      "\u00B1",
      paste(round(range(x, na.rm = TRUE), 1), collapse = "-")
    )
  )
}

stn %>%
  mutate(
    speed = distance / (as.numeric(duration) / 60),
    verticaltrawlopening = ifelse(
      verticaltrawlopening < 0,
      NA,
      verticaltrawlopening
    ),
    wirelengthratio = wirelength / bottomdepthstart
  ) %>%
  group_by(cruiseseries, geartype) %>%
  mutate() %>%
  summarise(
    n = n(),
    `Tow distance (nmi)` = colfun(distance),
    `Tow duration (min)` = colfun(duration),
    `Trawling speed (kn)` = colfun(speed),
    `Door spread (m)` = colfun(trawldoorspread),
    `Vertical trawl opening (m)` = colfun(verticaltrawlopening),
    `Wire length / bottom depth` = colfun(wirelengthratio),
    `Beaked redfish density (kg/nmi2)` = colfun(rho_biomass[
      commonname == "snabeluer"
    ]),
    `Golden redfish density (kg/nmi2)` = colfun(rho_biomass[
      commonname == "vanlig uer"
    ])
  ) %>%
  rename("Survey series" = "cruiseseries", "Trawl type" = "geartype") %>%
  make_table(
    caption = "Average trawl statistics (mean and range) by survey and trawl type."
  )
```


```{r, fig.cap="Trawl statistics by gear type."}
cowplot::plot_grid(
  ggplot(
    stn %>% distinct(startyear, serialnumber, geartype, duration, distance),
    aes(x = distance, fill = geartype)
  ) +
    geom_histogram() +
    labs(
      x = "Trawled distance (nmi)",
      y = "Number of stations",
      fill = "Survey"
    ) +
    # scale_y_continuous(expand = c(0,0)) +
    ggtitle("Distance") +
    theme(legend.position = "none"),
  ggplot(
    stn %>% distinct(startyear, serialnumber, geartype, duration, distance),
    aes(x = duration, fill = geartype)
  ) +
    geom_histogram() +
    labs(x = "Trawl duration (min)", y = "Number of stations") +
    # scale_y_continuous(expand = c(0,0)) +
    ggtitle("Duration") +
    theme(legend.position = c(0.8, 0.8)),
  ggplot(
    stn %>% distinct(startyear, serialnumber, geartype, duration, distance),
    aes(x = distance / (as.numeric(duration) / 60))
  ) +
    geom_histogram() +
    labs(x = "Trawl speed (knots; nmi/h)", y = "Number of stations") +
    # scale_y_continuous(expand = c(0,0)) +
    ggtitle("Speed"),
  ggplot(
    stn %>%
      distinct(
        startyear,
        serialnumber,
        geartype,
        duration,
        distance,
        trawldoorspread
      ),
    aes(x = trawldoorspread, fill = geartype)
  ) +
    geom_histogram() +
    labs(x = "Trawl door spread (m)", y = "Number of stations") +
    # scale_y_continuous(expand = c(0,0)) +
    ggtitle("Door spread") +
    theme(legend.position = "none"),
  ggplot(
    stn %>%
      distinct(
        startyear,
        serialnumber,
        geartype,
        duration,
        distance,
        trawldoorspread,
        verticaltrawlopening
      ),
    aes(x = verticaltrawlopening, fill = geartype)
  ) +
    geom_histogram() +
    labs(x = "Vertical trawl opening (m)", y = "Number of stations") +
    # scale_y_continuous(expand = c(0,0)) +
    ggtitle("Vertical trawl opening") +
    theme(legend.position = "none"),
  ggplot(
    stn %>%
      distinct(
        startyear,
        serialnumber,
        geartype,
        duration,
        distance,
        trawldoorspread,
        verticaltrawlopening,
        wirelength,
        bottomdepthstart
      ),
    aes(x = wirelength / bottomdepthstart)
  ) +
    geom_histogram() +
    labs(x = "Trawl wire length / bottom depth", y = "Number of stations") +
    # scale_y_continuous(expand = c(0,0)) +
    ggtitle("Wire length ratio"),
  ncol = 2,
  byrow = FALSE
)
```

## Effort

```{r, fig.cap = "Trawling duration by year and gear type."}
stn %>%
  distinct(
    startyear,
    serialnumber,
    platformname,
    geartype,
    duration,
    distance
  ) %>%
  filter(!is.na(duration)) %>%
  ggplot(aes(x = startyear, y = duration, group = startyear)) +
  geom_jitter(color = "grey", size = 0.3) +
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "crossbar",
    width = .5,
    color = "blue"
  ) +
  facet_wrap(~geartype) +
  scale_x_continuous(
    breaks = seq(1980, 2030, 4),
    labels = substr(seq(1980, 2030, 4), 3, 4)
  ) +
  coord_cartesian(ylim = c(0, NA)) +
  labs(x = "Year", y = "Trawl duration (h)") +
  scale_y_continuous(
    breaks = c(0, 15, 20, 30, 60, 90, 120, 200, 250),
    expand = c(0, 0)
  ) +
  theme_bw()
```

```{r, fig.cap = "Trawling distance by year and gear type."}
stn %>%
  distinct(startyear, serialnumber, platformname, geartype, distance) %>%
  ggplot(aes(x = startyear, y = distance, group = startyear)) +
  geom_jitter(color = "grey", size = 0.3) +
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "crossbar",
    width = .5,
    color = "blue"
  ) +
  facet_wrap(~geartype) +
  coord_cartesian(ylim = c(0, NA)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(
    breaks = seq(1980, 2030, 4),
    labels = substr(seq(1980, 2030, 4), 3, 4)
  ) +
  labs(x = "Year", y = "Trawl distance (nmi)") +
  theme_bw()
```

## Timing

```{r, fig-survey-time, fig.cap = "Overview of survey timing. Cruises have been separated.", fig.height=10}

tmp <- stn %>%
  filter(!is.na(stationstartdate)) %>%
  mutate(
    yday = as.Date(lubridate::yday(stationstartdate) - 1, origin = "2023-01-01")
  ) %>%
  group_by(startyear, cruise, cruiseseries, geartype) %>%
  reframe(min = min(yday), max = max(yday)) %>%
  mutate(
    cruise = factor(cruise, unique(.$cruise)),
    ypos = startyear +
      as.numeric(cruise) / (nlevels(cruise) * 2)
  )

tmp %>%
  ggplot(aes(
    x = min,
    xend = max,
    y = ypos,
    yend = ypos,
    color = cruiseseries,
    group = cruiseseries
  )) +
  geom_segment(linewidth = 2, alpha = 0.7) +
  scale_x_date("Month", date_labels = "%b", date_breaks = "1 months") +
  scale_y_continuous("Year", breaks = seq(1970, 2030, 1), expand = c(0.01, 0)) +
  coord_cartesian(ylim = c(NA, round_any(max(tmp$ypos), 1, ceiling))) +
  # scale_color_manual("Survey", values = survey_cols) +
  theme_bw() +
  theme(
    panel.grid.minor.y = element_blank()
  )
```

## Response distribution

```{r bm-density-comparison, fig.cap="Biomass density differences among cruise series (a-b) and gear categories (c-d). b and d have zero densities removed."}
cowplot::plot_grid(
  ggplot(stn, aes(x = cruiseseries, y = rho_biomass)) +
    geom_boxplot() +
    facet_wrap(~commonname) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    scale_y_continuous(
      trans = "log1p",
      breaks = c(0, 10, 1e2, 1e3, 1e4, 5e4, 1.5e5)
    ) +
    labs(
      x = "Cruise series",
      y = expression(paste("Density (kg ", nmi^-2, ")"))
    ),

  ggplot(
    stn %>%
      filter(rho_biomass > 0),
    aes(x = cruiseseries, y = rho_biomass)
  ) +
    geom_boxplot() +
    facet_wrap(~commonname) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    scale_y_continuous(
      trans = "log1p",
      breaks = c(0, 10, 1e2, 1e3, 1e4, 5e4, 1.5e5)
    ) +
    expand_limits(y = 0) +
    labs(
      x = "Cruise series",
      y = expression(paste("Density (kg ", nmi^-2, ")"))
    ) +
    theme(axis.title.y = element_blank()),

  ggplot(
    stn,
    aes(x = geartype, y = rho_biomass)
  ) +
    geom_boxplot() +
    facet_wrap(~commonname) +
    scale_y_continuous(
      trans = "log1p",
      breaks = c(0, 10, 1e2, 1e3, 1e4, 5e4, 1.5e5)
    ) +
    labs(x = "Trawl type", y = expression(paste("Density (kg ", nmi^-2, ")"))),

  ggplot(
    stn %>%
      filter(rho_biomass > 0),
    aes(x = geartype, y = rho_biomass)
  ) +
    geom_boxplot() +
    facet_wrap(~commonname) +
    scale_y_continuous(
      trans = "log1p",
      breaks = c(0, 10, 1e2, 1e3, 1e4, 5e4, 1.5e5)
    ) +
    expand_limits(y = 0) +
    labs(x = "Trawl type", y = expression(paste("Density (kg ", nmi^-2, ")"))) +
    theme(axis.title.y = element_blank()),

  labels = "auto"
)
```

```{r ab-density-comparison, fig.cap="Abundance density differences among cruise series (a-b) and gear categories (c-d). b and d have zero densities removed."}
cowplot::plot_grid(
  ggplot(stn, aes(x = cruiseseries, y = rho_abundance)) +
    geom_boxplot() +
    facet_wrap(~commonname) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    scale_y_continuous(
      trans = "log1p",
      breaks = c(0, 10, 1e2, 1e3, 1e4, 5e4, 3e5)
    ) +
    labs(
      x = "Cruise series",
      y = expression(paste("Density (n ", nmi^-2, ")"))
    ),

  ggplot(
    stn %>%
      filter(rho_biomass > 0),
    aes(x = cruiseseries, y = rho_abundance)
  ) +
    geom_boxplot() +
    facet_wrap(~commonname) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    scale_y_continuous(
      trans = "log1p",
      breaks = c(0, 10, 1e2, 1e3, 1e4, 5e4, 3e5)
    ) +
    expand_limits(y = 0) +
    labs(
      x = "Cruise series",
      y = expression(paste("Density (n ", nmi^-2, ")"))
    ) +
    theme(axis.title.y = element_blank()),

  ggplot(
    stn,
    aes(x = geartype, y = rho_abundance)
  ) +
    geom_boxplot() +
    facet_wrap(~commonname) +
    scale_y_continuous(
      trans = "log1p",
      breaks = c(0, 10, 1e2, 1e3, 1e4, 5e4, 3e5)
    ) +
    labs(x = "Trawl type", y = expression(paste("Density (n ", nmi^-2, ")"))),

  ggplot(
    stn %>%
      filter(rho_biomass > 0),
    aes(x = geartype, y = rho_abundance)
  ) +
    geom_boxplot() +
    facet_wrap(~commonname) +
    scale_y_continuous(
      trans = "log1p",
      breaks = c(0, 10, 1e2, 1e3, 1e4, 5e4, 3e5)
    ) +
    expand_limits(y = 0) +
    labs(x = "Trawl type", y = expression(paste("Density (n ", nmi^-2, ")"))) +
    theme(axis.title.y = element_blank()),

  labels = "auto"
)
```

```{r bm-density-depth, fig.cap = "Bottom depth versus swept area biomass density of each trawl haul (grey dots) allocated to species. The blue line and shading represents loess smoother and standard error. The y-axis is log + 1 transformed.", fig.height=7}
ggplot(stn, aes(x = bottomdepthstart, y = rho_biomass)) +
  geom_point(color = "grey", size = 0.3) +
  geom_smooth(fill = "blue") +
  labs(
    x = "Bottom depth (m)",
    y = expression(paste("Density (kg ", nmi^-2, ")"))
  ) +
  facet_grid(cruiseseries ~ commonname) +
  scale_y_continuous(
    trans = "log1p",
    breaks = c(0, 1, 10, 100, 1e3, 1e4, 5e4, 1e5, 5e5),
    expand = c(0.01, 0)
  ) +
  coord_cartesian(ylim = c(0, NA)) +
  theme_bw(base_size = 10) +
  scale_x_continuous(breaks = seq(0, 2000, 200), expand = c(0.01, 0))
```

## Maps

```{r, fig.cap = "Abundance (A, C, E) and biomass (B, D, F) swept area densities for beaked redfish (A, B), golden redfish (C, D) and uerslekten (E, F) averaged over all data for a grid covering the study area. Grey dots indicate bottom trawl stations.", fig.height=11}

grid <- st_as_sf(
  expand.grid(lon = attributes(p)$limits[1:2], lat = attributes(p)$limits[3:4]),
  coords = c("lon", "lat")
) %>%
  st_set_crs(attributes(p)$crs) %>%
  st_make_grid(n = 50) %>%
  st_transform(4326) %>%
  st_as_sf(data.frame(id = 1:length(.)), geometry = .)

plots <- lapply(levels(stn$commonname), function(k) {
  x <- stn %>%
    filter(commonname == k) %>%
    st_as_sf(coords = c("longitudestart", "latitudestart"), crs = 4326) %>%
    st_join(grid[, "id"])

  y <- x %>%
    st_set_geometry(NULL) %>%
    group_by(id) %>%
    summarise(
      bm = mean(rho_biomass, na.rm = TRUE),
      ab = mean(rho_abundance, na.rm = TRUE)
    )

  z1 <- grid %>% full_join(y, by = "id") %>% filter(!is.na(ab))

  z2 <- grid %>% full_join(y, by = "id") %>% filter(!is.na(bm))

  p1 <- basemap(data = stn, grid.col = NA) +
    geom_sf(data = z1, aes(fill = log10(ab + 1)), size = LS(0.05)) +
    geom_spatial_point(
      data = stn,
      aes(x = longitudestart, y = latitudestart),
      shape = ".",
      color = "grey40",
      size = 0.01,
      alpha = 0.1
    ) +
    labs(
      fill = "Mean abundance\ndensity (log10(rho + 1))\nind/nmi2",
      subtitle = k
    ) +
    scale_fill_distiller(palette = "Spectral", na.value = NA) +
    theme(
      legend.position = "bottom",
      axis.text = element_blank(),
      axis.title = element_blank()
    )

  p2 <- basemap(data = stn, grid.col = NA) +
    geom_sf(data = z2, aes(fill = log10(bm + 1)), size = LS(0.05)) +
    geom_spatial_point(
      data = stn,
      aes(x = longitudestart, y = latitudestart),
      shape = ".",
      color = "grey40",
      size = 0.01,
      alpha = 0.1
    ) +
    labs(
      fill = "Mean biomass\ndensity (log10(rho + 1))\nkg/nmi2",
      subtitle = k
    ) +
    scale_fill_distiller(palette = "Spectral", na.value = NA) +
    theme(
      legend.position = "bottom",
      axis.text = element_blank(),
      axis.title = element_blank()
    )

  list(reorder_layers(p1), reorder_layers(p2))
})

cowplot::plot_grid(plotlist = flattenlist(plots), ncol = 3, labels = "AUTO")
```

```{r, fig.cap = "Annual distribution of stations with biomass density of beaked redfish scaled to symbol color and size.", fig.height = 10}

x <- stn %>%
  filter(commonname == "snabeluer", !is.na(rho_biomass)) %>%
  mutate(value = rho_biomass)

bubble_map(x, ncol = 8)
```

```{r, fig.cap = "Annual distribution of stations with biomass density of golden redfish scaled to symbol color and size.", fig.height = 10}

x <- stn %>%
  filter(commonname == "vanlig uer", !is.na(rho_biomass)) %>%
  mutate(value = rho_biomass)

bubble_map(x, ncol = 8)
```

```{r, fig.cap = "Annual distribution of stations with biomass density of uerslekten scaled to symbol color and size.", fig.height = 10}

x <- stn %>%
  filter(commonname == "uerslekten", !is.na(rho_biomass)) %>%
  mutate(value = rho_biomass)

bubble_map(x, ncol = 8)
```

# Individual level data

```{r, cache = FALSE}
if (update_data) {
  tmp <- stn %>%
    dplyr::filter(biomass > 0 | (count > 0 & !is.na(count))) %>%
    dplyr::select(
      cruiseseries,
      platformname,
      cruise,
      serialnumber,
      commonname
    ) %>%
    distinct() %>%
    mutate(
      id = paste(cruise, platformname, serialnumber, commonname, sep = "_")
    )

  x <- indall %>%
    filter(
      commonname %in% c("snabeluer", "vanlig uer", "uerslekten", "uerfamilien"),
      gearcategory == "Bottom trawls",
      missionid %in% local(cs_list$missionid),
      !is.na(distance),
      !is.na(latitudestart),
      !is.na(longitudestart),
      !is.na(length)
    ) %>%
    collect() %>%
    dplyr::select(where(~ !all(is.na(.x)))) %>%
    mutate(
      id = paste(cruise, platformname, serialnumber, commonname, sep = "_")
    ) %>%
    filter(id %in% tmp$id) %>% # note that the station filtering is done here
    dplyr::select(-id) %>%
    mutate(sex = ifelse(!is.na(sex) & !sex %in% 1:2, NA, sex))

  x[
    x$commonname %in% c("uerslekten", "uerfamilien"),
  ]$commonname <- "uerslekten"

  # x$age[is.na(x$readingdate)] <- NA # Remove the old readings

  # Correct maturation stage

  # matStages <- readxl::read_excel("data/in/Atlantic halibut maturity stages.xlsx")
  # ms <- na.omit(matStages[c("maturationstage", "maturationstage_becomes")])
  # ms[,grep("becomes", names(ms))] <- suppressWarnings(as.integer(ms[[grep("becomes", names(ms))]]))
  # ss <- na.omit(matStages[c("specialstage", "specialstage_becomes")])
  # ss[,grep("becomes", names(ss))] <- suppressWarnings(as.integer(ss[[grep("becomes", names(ss))]]))
  #
  # x <- x %>% mutate(
  #   maturationstage =
  #     recode(maturationstage, !!!setNames(ms$maturationstage_becomes,
  #                                         ms$maturationstage)),
  #   specialstage =
  #     recode(specialstage, !!!setNames(ss$specialstage_becomes,
  #                                      ss$specialstage)))
  #
  # x[x$maturationstage %in% "1" & is.na(x$specialstage), ]$specialstage <- 1
  #
  # x[x$sex %in% "1" & !is.na(x$specialstage),]$maturationstage <-
  #   x[x$sex %in% "1" & !is.na(x$specialstage),]$specialstage

  ## Cruise series name

  x <- x %>%
    left_join(
      cs_list %>%
        select(
          startyear,
          cruise,
          callsignal,
          platformname,
          missionid,
          cruiseseries
        ),
      by = join_by(startyear, platformname, cruise, callsignal, missionid)
    ) %>%
    relocate(cruiseseries, .after = startyear)

  ## Reduce number of columns and save

  ind <- x %>%
    dplyr::select(-cruiseseriescode) %>%
    group_by(
      # new stuff test it.
      cruiseseries,
      startyear,
      cruise,
      callsignal,
      serialnumber,
      catchsampleid,
      commonname
    ) %>%
    mutate(lengthsamplecount = n()) %>%
    ungroup() %>%
    mutate(
      repN = catchcount / lengthsamplecount,
      repN = ifelse(is.na(repN), lengthsamplecount, repN)
    ) %>%
    rename("superindividuals" = "repN") %>%
    dplyr::select(
      cruiseseries,
      startyear,
      cruise,
      platformname,
      callsignal,
      serialnumber,
      commonname,
      catchpartnumber,
      specimenid,
      superindividuals,
      sex,
      maturationstage,
      catchweight,
      catchcount,
      individualweight,
      length,
      age
    ) %>%
    mutate(length = round(100 * length + 1e-6, 0)) %>%
    mutate(commonname = factor(commonname, levels = levels(stn$commonname)))

  write_rds(ind, 'data/survey data/ind_Database.rds', compress = "xz")
} else {
  ind <- read_rds('data/survey data/ind_Database.rds')
}
```

## Individual overview

```{r}
tmp <- ind %>%
  group_by(cruiseseries, startyear, serialnumber, commonname) %>%
  reframe(n = sum(superindividuals)) %>%
  group_by(cruiseseries, startyear, commonname) %>%
  reframe(total_measured = sum(n), stations_with_measured = n()) %>%
  full_join(
    stn %>%
      filter(biomass > 0) %>%
      group_by(cruiseseries, startyear, commonname) %>%
      reframe(total_positive_stations = n()),
    by = join_by(cruiseseries, startyear, commonname)
  ) %>%
  replace_na(list(total_measured = 0, stations_with_measured = 0)) %>%
  mutate(
    Yes = stations_with_measured,
    No = total_positive_stations - stations_with_measured
  ) %>%
  arrange(cruiseseries, startyear, commonname) %>%
  pivot_longer(cols = c(Yes, No)) %>%
  mutate(name = factor(name, levels = c("Yes", "No")))
```

`r tmp %>% filter(name == "No") %>% group_by(commonname) %>% reframe(value = sum(value)) %>% pull(value) %>% paste(collapse = ", ")` out of `r tmp %>% group_by(commonname) %>% reframe(value = sum(value)) %>% pull(value) %>% paste(collapse = ", ")` stations with `r tmp %>% filter(name == "No") %>% group_by(commonname) %>% reframe(value = sum(value)) %>% pull(commonname) %>% paste(collapse = ", ")` in catch were not length measured, respectively (Figure \@ref(fig:n-length-measured)).

```{r n-length-measured, fig.cap = paste("Number of stations with and without length measured", sp_name, "by year and cruise series."), fig.height=7}
tmp %>%
  ggplot(aes(x = startyear, y = value, fill = name)) +
  geom_col() +
  labs(
    x = "Year",
    y = "Number of stations",
    fill = "Length\nmeasured\nstation"
  ) +
  scale_y_continuous(expand = c(0.01, 0)) +
  scale_x_continuous(
    expand = c(0, 0),
    breaks = seq(1980, 2030, 4),
    labels = substr(seq(1980, 2030, 4), 3, 4)
  ) +
  facet_grid(commonname ~ cruiseseries) +
  theme_bw() +
  scale_fill_manual(values = c("Yes" = "grey", "No" = "darkred")) +
  theme(legend.position = "bottom")
```

## Length distributions

```{r, fig.cap = paste("Size distribution of", sp_name, "among years extrapolated to the entire catch from length and individual measured fish."), fig.height=12}
ind %>%
  mutate(
    sex = factor(
      ifelse(
        is.na(sex),
        "Unknown",
        ifelse(sex == 1, "Female", ifelse(sex == 2, "Male", "Wrong"))
      ),
      levels = rev(c("Female", "Male", "Unknown", "Wrong"))
    )
  ) %>%
  uncount(round(superindividuals, 0)) %>%
  ggplot(aes(x = length)) +
  geom_histogram(
    aes(fill = commonname),
    binwidth = 2,
    color = "black",
    linewidth = 0.1
  ) +
  facet_wrap(~startyear, ncol = 5, scales = "free_y", dir = "v") +
  labs(y = "Count", x = "Length (cm)", fill = "Species", color = "Species") +
  scale_y_continuous(breaks = function(k) round(c(0, max(k, na.rm = TRUE)))) +
  scale_x_continuous() +
  # scale_fill_manual(
  #   values = c("Female" = "#FF5F68", "Male" = "#449BCF", "Unknown" = "grey",
  #              "Wrong" = "black"), drop = TRUE, limits = force) +
  guides(fill = guide_legend(reverse = TRUE)) +
  coord_cartesian(expand = FALSE) +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(size = 10, face = 2)
  )
```

```{r, fig.cap = paste("Summed size distribution of", sp_name, " extrapolated to the entire catch from length and individual measured fish."), fig.height=6}
ind %>%
  mutate(
    sex = factor(
      ifelse(
        is.na(sex),
        "Unknown",
        ifelse(sex == 1, "Female", ifelse(sex == 2, "Male", "Wrong"))
      ),
      levels = rev(c("Female", "Male", "Unknown", "Wrong"))
    )
  ) %>%
  uncount(round(superindividuals, 0)) %>%
  ggplot(aes(x = length)) +
  geom_histogram(
    aes(fill = commonname),
    binwidth = 2,
    color = "black",
    linewidth = 0.1
  ) +
  facet_wrap(~cruiseseries, ncol = 3, scales = "free_y") +
  labs(y = "Count", x = "Length (cm)", fill = "Species", color = "Species") +
  scale_y_continuous(breaks = function(k) round(c(0, max(k, na.rm = TRUE)))) +
  scale_x_continuous() +
  # scale_fill_manual(
  #   values = c("Female" = "#FF5F68", "Male" = "#449BCF", "Unknown" = "grey",
  #              "Wrong" = "black"), drop = TRUE, limits = force) +
  guides(fill = guide_legend(reverse = TRUE)) +
  coord_cartesian(expand = FALSE) +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(size = 10, face = 2)
  )
```

## Age

```{r, fig.cap = "Beaked redfish growth for all survey data. A shows all data, B data with sex. Red refers to females and blue to males in B.", fig.height=8}

cowplot::plot_grid(
  ggFishPlots::plot_growth(
    ind %>%
      filter(commonname == "snabeluer") %>%
      filter(!is.na(age)), #%>%
    #uncount(round(superindividuals, 0))
    length.unit = "cm",
    legend.position = "none"
  )$plot,

  ggFishPlots::plot_growth(
    ind %>%
      filter(commonname == "snabeluer") %>%
      filter(!is.na(age), !is.na(sex)), #%>%
    #uncount(round(superindividuals, 0))
    female.sex = "1",
    male.sex = "2",
    split.by.sex = TRUE,
    length.unit = "cm",
    legend.position = "none"
  )$plot,

  labels = "AUTO",
  ncol = 1
)

# cowplot::plot_grid(
#   plotlist =
#     lapply(ind$cruiseseries %>% unique, function(k) {
#       print(k)
#
#       if(nrow(ind[ind$cruiseseries == k & !is.na(ind$age), ]) > 10) {
#
#         ggFishPlots::plot_growth(
#           ind %>% filter(cruiseseries == k, !is.na(age), !is.na(sex)) %>%
#             uncount(superindividuals),
#           female.sex = "1", male.sex = "2", split.by.sex = TRUE,
#           length.unit = "cm", legend.position = "none")$plot  +
#           ggtitle(k) +
#           coord_cartesian(ylim = c(0,130), xlim = c(0,30), expand = FALSE) +
#           scale_y_continuous(breaks = seq(0,150,25))
#
#       } else {
#
#         ggplot(data.frame(x=1:10,y=0:1), aes(x,y)) +
#           geom_blank() +
#           annotate("text", x = 15, y = 70, label = "Not enough\ndata") +
#           coord_cartesian(ylim = c(0,130), xlim = c(0,30), expand = FALSE) +
#           ggtitle(k) +
#           labs(x = "Age (years)", y = "Total length (cm)")
#       }
#     })
# )
```

```{r, fig.cap = "Golden redfish growth for all survey data. A shows all data, B data with sex. Red refers to females and blue to males in B.", fig.height=8}

cowplot::plot_grid(
  ggFishPlots::plot_growth(
    ind %>%
      filter(commonname == "vanlig uer") %>%
      filter(!is.na(age)), # %>%
    # uncount(round(superindividuals, 0))
    length.unit = "cm",
    legend.position = "none"
  )$plot,

  ggFishPlots::plot_growth(
    ind %>%
      filter(commonname == "vanlig uer") %>%
      filter(!is.na(age), !is.na(sex)), # %>%
    # uncount(round(superindividuals, 0))
    female.sex = "1",
    male.sex = "2",
    split.by.sex = TRUE,
    length.unit = "cm",
    legend.position = "none"
  )$plot,

  labels = "AUTO",
  ncol = 1
)
```

<!-- ```{r, fig.cap = "Uerslekten growth for all survey data.", fig.height=8} -->
<!--   ggFishPlots::plot_growth( -->
<!--     ind %>%  -->
<!--       filter(commonname == "uerslekten") %>%  -->
<!--       filter(!is.na(age)) %>% -->
<!--       uncount(superindividuals), -->
<!--     length.unit = "cm", legend.position = "none")$plot -->
<!-- ``` -->

## Maturity

```{r, fig.cap = "Beaked redfish maturity ogives for all survey data. A shows all data, B data with sex. Red refers to females and blue to males in B.", fig.height=8}

cowplot::plot_grid(
  ggFishPlots::plot_maturity(
    ind %>%
      filter(commonname == "snabeluer") %>%
      filter(!is.na(maturationstage)) %>%
      mutate(maturity = maturationstage > 1), # %>%
    # uncount(round(superindividuals, 0))
    legend.position = "none",
    length.bin.width = 5
  )$plot,

  ggFishPlots::plot_maturity(
    ind %>%
      filter(commonname == "snabeluer") %>%
      filter(!is.na(maturationstage), !is.na(sex), sex != 3) %>%
      mutate(maturity = maturationstage > 1), # %>%
    # uncount(superindividuals)
    female.sex = "1",
    male.sex = "2",
    split.by.sex = TRUE,
    legend.position = "none",
    length.bin.width = 5
  )$plot,
  labels = "AUTO",
  ncol = 1
)

# cowplot::plot_grid(
#   plotlist =
#     lapply(ind$cruiseseries %>% unique, function(k) {
#       # print(k)
#       ggFishPlots::plot_maturity(
#         ind %>%
#           filter(cruiseseries == k, !is.na(maturationstage), !is.na(sex), sex != 3) %>%
#           mutate(maturity = maturationstage > 1) %>%
#           uncount(superindividuals),
#         female.sex = "1", male.sex = "2", split.by.sex = TRUE,
#         legend.position = "none")$plot  +
#         ggtitle(k)
#     })
# )
```

```{r, fig.cap = "Golden redfish maturity ogives for all survey data. A shows all data, B data with sex. Red refers to females and blue to males in B.", fig.height=8}

cowplot::plot_grid(
  ggFishPlots::plot_maturity(
    ind %>%
      filter(commonname == "snabeluer") %>%
      filter(!is.na(maturationstage)) %>%
      mutate(maturity = maturationstage > 1), # %>%
    # uncount(round(superindividuals, 0))
    legend.position = "none",
    length.bin.width = 5
  )$plot,

  ggFishPlots::plot_maturity(
    ind %>%
      filter(commonname == "snabeluer") %>%
      filter(!is.na(maturationstage), !is.na(sex), sex != 3) %>%
      mutate(maturity = maturationstage > 1), # %>%
    # uncount(round(superindividuals, 0))
    female.sex = "1",
    male.sex = "2",
    split.by.sex = TRUE,
    legend.position = "none",
    length.bin.width = 5
  )$plot,
  labels = "AUTO",
  ncol = 1
)
```


# Combined station level length distributions

The maths for this part can be confusing. They have been through through, but have not been peer-reviewed. There are two methods to choose from. The first option (used currently) uses the densities calculated using station data and splits them to length groups such that the sum for each station equals that of in station data. The second option uses individual data to calculate the densities. Abundance densities between these two methods tend to be equal, but biomass densities differ somewhat because all fish were not weighted. 

Both methods require estimating weight for each length-measured fish. Since weight has not been measured for most fish, missing values need to be filled using a length-weight model.

## Length-weight relationships

Number of weight measurements varied among species but was generally high enough to calculate length-weight relationships (Table \@ref(tab:tbl-lwn)). All lengths in the database were total lengths in centimeters and weights round weights in kilograms.

```{r tbl-lwn}
ind %>%
  group_by(commonname) %>%
  reframe(weight = sum(individualweight > 0, na.rm = TRUE), length = n()) %>%
  make_table(
    caption = "Number of weight and length samples for each species in the dataset."
  )
```

The length-weight relationships were calculated using the [`plot_lw`](https://deepwaterimr.github.io/ggFishPlots/reference/plot_lw.html) function from the [ggFishPlots](https://CRAN.R-project.org/package=ggFishPlots) package. The nonlinear equation $weight = a \times length^b$ was fitted using nonlinear least squares (nls function in R) improving the fit to large individuals over the standard logarithmic linear regression method. Observations yielding absolute residual values from initial logarithmic regression, $log(weight) = a + b \times log(length)$, exceeding the 99.5 percentile were considered outliers and removed from the nonlinear models. These values had little influence on the coefficients, however. 

Estimation of the length-weight relationships:

```{r}
if (update_data) {
  tmp <- lapply(levels(ind$commonname), function(k) {
    ggFishPlots::plot_lw(
      ind %>% filter(commonname == k),
      weight = "individualweight",
      use.nls = TRUE,
      outlier.percentile = 99.5,
      annotate.coefficients = TRUE
    )
  }) %>%
    setNames(levels(ind$commonname))

  lw_coefficients <-
    lapply(seq_along(tmp), function(i) {
      tmp[[i]]$param %>% mutate(species = names(tmp[i]), .before = 1)
    }) %>%
    bind_rows()

  saveRDS(
    tmp,
    file = "data/definitions/length-weight_models.rda",
    compress = "xz"
  )
  write_csv(
    lw_coefficients,
    file = "data/definitions/length-weight_coefficients.csv"
  )
} else {
  tmp <- readr::read_rds("data/definitions/length-weight_models.rda")
  lw_coefficients <- readr::read_csv(
    "data/definitions/length-weight_coefficients.csv",
    show_col_types = FALSE
  )
}
```

The relationships are shown in the figure below. Coefficients and associated statistics in the tables under. 

```{r, fig.cap = "Length-weight relationships for beaked redfish, golden redfish and uerslekten, respectively."}
cowplot::plot_grid(
  plotlist = lapply(seq_along(tmp), function(i) {
    tmp[[i]]$plot +
      ggtitle(names(tmp)[i])
  }),
  nrow = 1,
  labels = "AUTO"
)
```

```{r}
knitr::kable(
  lw_coefficients,
  caption = "Length-weight coefficients and associated statistics.",
  format.args = list(big.mark = ""),
  digits = c(0, 0, 10, 8, 1, 4, 10, 10, 0, 0, 0, 0, 0, 0),
  row.names = FALSE
) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "condensed"))
```

Parameters for beaked redfish:

```{r, results = 'asis'}
htmlcat(tmp[[1]]$text)
```

Parameters for golden redfish:

```{r, results = 'asis'}
htmlcat(tmp[[2]]$text)
```

Parameters for uerslekten:

```{r, results = 'asis'}
htmlcat(tmp[[3]]$text)
```

## Calculus

Catches and biomass densities for each length group and species were calculted using the length-weight relationships above: 

```{r}
if (update_data) {
  stn_ldist <- ind %>%
    dplyr::select(-sex) %>% # the sex column multiplies the length of the final dataframe by 3. Not worth including for these species
    # mutate(sex = factor(ifelse(is.na(sex), "Unknown",
    #                            ifelse(sex > 2, "Unknown",
    #                                   ifelse(sex == 2, "Male", "Female"))))) %>%
    left_join(
      lw_coefficients %>%
        dplyr::select(species, term, estimate) %>%
        pivot_wider(names_from = term, values_from = estimate) %>%
        rename("commonname" = "species"),
      by = join_by(commonname)
    ) %>%
    mutate(weight = a * length^b) %>%
    dplyr::select(-a, -b) %>%
    full_join(
      stn,
      by = join_by(
        cruiseseries,
        startyear,
        cruise,
        platformname,
        callsignal,
        serialnumber,
        commonname
      )
    ) %>%
    mutate(
      length_group = cut(length, seq(0, 95, by = 5))
    ) %>%
    replace_na(list(superindividuals = 1)) %>%
    # uncount(superindividuals) %>%
    mutate(commonname = factor(commonname, levels = levels(stn$commonname)))

  if (!"individualweight" %in% colnames(stn_ldist)) {
    stn_ldist <- stn_ldist %>% mutate(individualweight = NA)
  }

  if (TRUE) {
    # This method uses densities from stn and splits them to length groups such that the sum for each station equals that of stn
    stn_ldist <-
      stn_ldist %>%
      group_split(commonname) %>%
      map(
        ~ {
          .x %>%
            mutate(
              weight = ifelse(is.na(individualweight), weight, individualweight)
            ) %>%
            group_by(
              cruiseseries,
              startyear,
              cruise,
              platformname,
              callsignal,
              serialnumber,
              stationstartdate,
              longitudestart,
              latitudestart,
              bottomdepthstart,
              gear,
              gearname,
              geartype,
              distance,
              gear_width,
              commonname,
              length_group,
              count,
              biomass,
              rho_abundance,
              rho_biomass
            ) %>%
            reframe(n = sum(superindividuals), weight = n * mean(weight)) %>%
            mutate(n = ifelse(is.na(length_group), 0, n)) %>%
            filter(!(is.na(length_group) & rho_biomass > 0)) %>% # removes stations and rows with missing length data
            group_by(
              cruiseseries,
              startyear,
              cruise,
              platformname,
              callsignal,
              serialnumber,
              stationstartdate,
              longitudestart,
              latitudestart,
              bottomdepthstart,
              gear,
              gearname,
              geartype,
              distance,
              gear_width,
              commonname
            ) %>%
            mutate(total_n = sum(n), total_w = sum(weight)) %>%
            ungroup() %>%
            mutate(weight = weight / (total_w / biomass)) %>%
            mutate(
              rho_biomass = ifelse(
                biomass == 0,
                0,
                rho_biomass * (weight / biomass)
              ),
              rho_abundance = ifelse(n == 0, 0, rho_abundance * (n / total_n))
            ) %>%
            mutate(n = n / (total_n / count)) %>%
            dplyr::select(-total_n, -total_w) %>%
            rename(total_weight = biomass, total_n = count) %>%
            group_by(
              cruiseseries,
              startyear,
              cruise,
              platformname,
              callsignal,
              serialnumber,
              stationstartdate,
              longitudestart,
              latitudestart,
              bottomdepthstart,
              gear,
              gearname,
              geartype,
              distance,
              gear_width,
              commonname
            ) %>%
            complete(
              length_group,
              fill = list(
                n = 0,
                weight = 0,
                rho_abundance = 0,
                rho_biomass = 0
              ),
              explicit = FALSE
            ) %>%
            mutate(total_n = sum(n), total_weight = sum(weight)) %>%
            replace_na(list(total_n = 0, total_weight = 0)) %>%
            relocate(rho_abundance, rho_biomass, .after = last_col()) %>%
            filter(!is.na(length_group))
        }
      ) %>%
      bind_rows()
  } else {
    # This method calculates densities for each length group from ground up. They will not match those in stn.

    stn_ldist <-
      stn_ldist %>%
      mutate(
        individualweight = ifelse(
          is.na(individualweight),
          weight,
          individualweight
        )
      ) %>%
      group_by(
        cruiseseries,
        startyear,
        cruise,
        platformname,
        serialnumber,
        longitudestart,
        latitudestart,
        bottomdepthstart,
        distance,
        gear_width,
        sex,
        length_group
      ) %>%
      reframe(n = n(), weight = sum(weight)) %>%
      mutate(
        n = ifelse(is.na(sex), 0, n),
        weight = ifelse(is.na(sex), 0, weight),
        rho_biomass = weight / (distance * (gear_width / 1852)),
        rho_abundance = n / (distance * (gear_width / 1852))
      ) %>%
      group_by(
        cruiseseries,
        startyear,
        cruise,
        platformname,
        serialnumber,
        longitudestart,
        latitudestart,
        bottomdepthstart
      ) %>%
      mutate(total_n = sum(n), total_weight = sum(weight)) %>%
      group_by(
        cruiseseries,
        startyear,
        cruise,
        platformname,
        serialnumber,
        longitudestart,
        latitudestart,
        bottomdepthstart
      ) %>%
      complete(
        sex,
        length_group,
        fill = list(
          total_n = 0,
          total_weight = 0,
          n = 0,
          weight = 0,
          rho_abundance = 0,
          rho_biomass = 0
        ),
        explicit = FALSE
      ) %>%
      filter(!is.na(sex), !is.na(length_group)) %>%
      relocate(rho_abundance, rho_biomass, .after = last_col())
  }

  # ind_method %>%
  #   group_by(cruiseseries, startyear, platformname, serialnumber) %>%
  #   reframe(ab_ind = sum(rho_abundance), bm_ind = sum(rho_biomass)) %>%
  #   full_join(
  #     stn_method %>%
  #       group_by(cruiseseries, startyear, platformname, serialnumber) %>%
  #       reframe(ab_stn = sum(rho_abundance), bm_stn = sum(rho_biomass))
  #   )

  write_rds(
    stn_ldist,
    'data/survey data/stn_ldist_Database.rds',
    compress = "xz"
  )
} else {
  stn_ldist <- read_rds('data/survey data/stn_ldist_Database.rds')
}
```

## Tests 

```{r}
nstn_ldist <- stn %>%
  group_by(cruiseseries, startyear, callsignal, serialnumber, commonname) %>%
  reframe(rho = sum(rho_biomass)) %>%
  group_by(cruiseseries, commonname) %>%
  reframe(
    `All stations` = n(),
    `Zero stations` = sum(rho == 0, na.rm = TRUE)
  ) %>%
  full_join(
    stn_ldist %>%
      group_by(
        cruiseseries,
        startyear,
        callsignal,
        serialnumber,
        commonname
      ) %>%
      reframe(rho = sum(rho_biomass)) %>%
      group_by(cruiseseries, commonname) %>%
      reframe(
        `Measured stations` = n(),
        `Zero stations 2` = sum(rho == 0, na.rm = TRUE)
      ),
    by = join_by(cruiseseries, commonname)
  ) %>%
  ungroup()
```

Number of zero stations in stn and stn_ldist match: `r all.equal(nstn_ldist[["Zero stations"]], nstn_ldist[["Zero stations 2"]])`

```{r}
# Make sure that the critical catch information match between stn and snt_ldist

tmp <- stn %>%
  group_by(cruiseseries, startyear, callsignal, serialnumber, commonname) %>%
  reframe(
    bm_stn = sum(rho_biomass, na.rm = TRUE),
    ab_stn = sum(rho_abundance, na.rm = TRUE),
    w_stn = sum(biomass, na.rm = TRUE),
    n_stn = sum(count, na.rm = TRUE)
  ) %>%
  left_join(
    stn_ldist %>%
      group_by(
        cruiseseries,
        startyear,
        callsignal,
        serialnumber,
        commonname
      ) %>%
      reframe(
        bm_ldist = sum(rho_biomass, na.rm = TRUE),
        ab_ldist = sum(rho_abundance, na.rm = TRUE),
        w_ldist = sum(weight, na.rm = TRUE),
        n_ldist = sum(n, na.rm = TRUE)
      )
  ) %>%
  mutate(
    bm_equal = abs(bm_stn - bm_ldist) < 0.1,
    ab_equal = abs(ab_stn - ab_ldist) < 0.1,
    n_equal = abs(n_stn - n_ldist) < 0.1,
    w_equal = abs(w_stn - w_ldist) < 0.1
  )
```

Sum of `rho_biomass` in stn equal to the sum of the same column in stn_ldist: `r all(tmp$bm_equal, na.rm = TRUE)`. `r tmp %>% filter(commonname == "snabeluer") %>% pull(bm_equal) %>% is.na %>% sum` removed beaked redfish and `r tmp %>% filter(commonname == "vanlig uer") %>% pull(bm_equal) %>% is.na %>% sum` removed golden redfish stations due to missing length measurements, respectively. 

Sum of `rho_abundance` in stn equal to the sum of the same column in stn_ldist: `r all(tmp$ab_equal, na.rm = TRUE)`. `r tmp %>% filter(commonname == "snabeluer") %>% pull(ab_equal) %>% is.na %>% sum` removed beaked redfish and `r tmp %>% filter(commonname == "vanlig uer") %>% pull(ab_equal) %>% is.na %>% sum` removed golden redfish stations due to missing length measurements, respectively. 

Sum of the `count` column in stn equal to the sum of the `n` column in stn_ldist: `r all(tmp$n_equal, na.rm = TRUE)`

Sum of the `biomass` column in stn equal to the sum of the `weight` column in stn_ldist: `r all(tmp$w_equal, na.rm = TRUE)`

```{r, fig.cap = "Control plot to compare individual length distribution data and combined station and length distribution data.", fig.height=7}
stn_ldist %>%
  group_by(cruiseseries, commonname, length_group) %>%
  reframe(n = sum(n, na.rm = TRUE)) %>%
  mutate(type = "stn_ldist") %>%
  bind_rows(
    ind %>%
      mutate(length_group = cut(length, seq(0, 95, by = 5))) %>%
      group_by(cruiseseries, commonname, length_group) %>%
      reframe(n = sum(superindividuals)) %>%
      mutate(type = "ind")
  ) %>%
  # pivot_wider(names_from = type, values_from = n) %>%
  # mutate(diff = 100*(stn_ldist - ind)/ind) %>%
  ggplot(aes(x = length_group, y = n, fill = type)) +
  geom_col(color = "black", linewidth = 0.1, position = "dodge") +
  facet_wrap(cruiseseries ~ commonname, scales = "free_y") +
  labs(y = "Count", x = "Length group (cm)") +
  scale_y_continuous(breaks = function(k) round(c(0, max(k, na.rm = TRUE)))) +
  coord_cartesian(expand = FALSE) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(size = 10, face = 2)
  )
```

## Combined overview

```{r}
nstn_ldist %>%
  dplyr::select(-`Zero stations 2`) %>% # Zero stations and Zero stations 2 have to match
  mutate(
    `Not measured\nstations` = `All stations` - `Measured stations`,
    `Measured stations` = `Measured stations` - `Zero stations`
  ) %>%
  rename("Survey" = "cruiseseries") %>%
  make_table(
    caption = paste(
      "All stations from station data, stations with zero",
      sp_name,
      ", stations with length data and stations excluded from the combined dataset due to lack of length measurements by data source."
    )
  )
```

```{r n-stn-ldist, fig.cap = paste("Number of stations in the length resolved dataset with and without", sp_name, "by year. White indicates the number of lost stations due to missing length data.")}
tmpdt <- stn_ldist %>%
  group_by(cruiseseries, startyear, callsignal, serialnumber, commonname) %>%
  reframe(rho_biomass = sum(rho_biomass, na.rm = TRUE)) %>%
  group_by(startyear, commonname) %>%
  reframe(
    yes = sum(rho_biomass > 0, na.rm = TRUE),
    no = sum(rho_biomass == 0, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = c(yes, no)) %>%
  mutate(name = factor(name, c("no", "yes"))) %>%
  left_join(
    stn %>%
      distinct(
        cruiseseries,
        startyear,
        callsignal,
        serialnumber,
        commonname
      ) %>%
      group_by(startyear, commonname) %>%
      count()
  )

ggplot() +
  geom_col(
    data = tmpdt %>% distinct(startyear, commonname, n),
    aes(x = startyear, y = n),
    fill = NA,
    color = "black"
  ) +
  geom_col(
    data = tmpdt,
    aes(x = startyear, y = value, fill = name),
    color = "black"
  ) +
  labs(x = "Year", y = "Number of stations", fill = "Contains\ntarget taxa") +
  facet_wrap(~commonname) +
  scale_fill_manual(values = c("yes" = "#66a182", "no" = "grey")) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(
    breaks = seq(1980, 2030, 2),
    labels = substr(seq(1980, 2030, 2), 3, 4),
    expand = c(0.01, 0)
  ) +
  theme(legend.position = "bottom")
```


```{r, fig.cap = "Average biomass density + 95% confidence interval by survey time-series and length group for the combined station and length dataset.", fig.height=7}

length_groups <- stn_ldist %>%
  filter(rho_biomass > 0 & !is.na(rho_biomass)) %>%
  droplevels() %>%
  pull(length_group) %>%
  levels()

stn_ldist %>%
  filter(length_group %in% length_groups) %>%
  droplevels() %>%
  group_by(cruiseseries, commonname, length_group) %>%
  reframe(
    value = mean(rho_biomass, na.rm = TRUE),
    se = sd(rho_biomass, na.rm = TRUE) / sqrt(length(!is.na(rho_biomass)))
  ) %>%
  ggplot(aes(
    x = length_group,
    y = value,
    ymin = value,
    ymax = value + 1.96 * se
  )) +
  geom_col(
    color = "black",
    linewidth = 0.3,
    position = "dodge",
    fill = "grey"
  ) +
  geom_errorbar(width = 0.2, linewidth = 0.3) +
  facet_wrap(cruiseseries ~ commonname, scales = "free") +
  labs(
    y = expression(paste("Average density (kg ", nmi^-2, ")")),
    x = "Length group (cm)"
  ) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  coord_cartesian(expand = FALSE) +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(size = 10, face = 2)
  )
```

```{r, fig.cap = "Average abundance density + 95% confidence interval by survey time-series for the combined station and length dataset.", fig.height=7}

length_groups <- stn_ldist %>%
  filter(rho_abundance > 0, !is.na(rho_abundance)) %>%
  droplevels() %>%
  pull(length_group) %>%
  levels()

stn_ldist %>%
  filter(length_group %in% length_groups) %>%
  droplevels() %>%
  group_by(cruiseseries, commonname, length_group) %>%
  reframe(
    value = mean(rho_abundance, na.rm = TRUE),
    se = sd(rho_abundance, na.rm = TRUE) / sqrt(length(!is.na(rho_abundance)))
  ) %>%
  ggplot(aes(
    x = length_group,
    y = value,
    ymin = value,
    ymax = value + 1.96 * se
  )) +
  geom_col(
    color = "black",
    linewidth = 0.3,
    position = "dodge",
    fill = "grey"
  ) +
  geom_errorbar(width = 0.2, linewidth = 0.3) +
  facet_wrap(cruiseseries ~ commonname, scales = "free_y") +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  labs(
    y = expression(paste("Average density (n ", nmi^-2, ")")),
    x = "Length group (cm)"
  ) +
  coord_cartesian(expand = FALSE) +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(size = 10, face = 2)
  )
```

```{r, fig.cap = paste("Average biomass (A) and abundance (B) density of", sp_names[1], "+ 95% confidence interval by year and length group for the combined station and length dataset."), fig.height=12}

length_groups <- stn_ldist %>%
  filter(commonname == filter_sp[1], rho_biomass > 0) %>%
  droplevels() %>%
  pull(length_group) %>%
  levels()

p1 <- stn_ldist %>%
  filter(commonname == filter_sp[1]) %>%
  filter(length_group %in% length_groups) %>%
  droplevels() %>%
  group_by(startyear, length_group) %>%
  reframe(
    value = mean(rho_biomass, na.rm = TRUE),
    se = sd(rho_biomass, na.rm = TRUE) / sqrt(length(!is.na(rho_biomass)))
  ) %>%
  ggplot(aes(
    x = length_group,
    y = value,
    ymin = value,
    ymax = value + 1.96 * se
  )) +
  geom_col(
    color = "black",
    linewidth = 0.3,
    position = "dodge",
    fill = "grey"
  ) +
  geom_errorbar(width = 0.2, linewidth = 0.3) +
  facet_wrap(~startyear, ncol = 6, dir = "v", scales = "free_y") +
  labs(
    y = expression(paste("Average density (kg ", nmi^-2, ")")),
    x = "Length group (cm)"
  ) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  coord_cartesian(expand = FALSE) +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(size = 10, face = 2)
  )

p2 <- stn_ldist %>%
  filter(commonname == filter_sp[1]) %>%
  filter(length_group %in% length_groups) %>%
  droplevels() %>%
  group_by(startyear, length_group) %>%
  reframe(
    value = mean(rho_abundance, na.rm = TRUE),
    se = sd(rho_abundance, na.rm = TRUE) / sqrt(length(!is.na(rho_abundance)))
  ) %>%
  ggplot(aes(
    x = length_group,
    y = value,
    ymin = value,
    ymax = value + 1.96 * se
  )) +
  geom_col(
    color = "black",
    linewidth = 0.3,
    position = "dodge",
    fill = "grey"
  ) +
  geom_errorbar(width = 0.2, linewidth = 0.3) +
  facet_wrap(~startyear, ncol = 6, dir = "v", scales = "free_y") +
  labs(
    y = expression(paste("Average density (n ", nmi^-2, ")")),
    x = "Length group (cm)"
  ) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  coord_cartesian(expand = FALSE) +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(size = 10, face = 2)
  )

cowplot::plot_grid(p1, p2, labels = "AUTO", ncol = 1)
```

```{r, fig.cap = paste("Average biomass (A) and abundance (B) density of", sp_names[2], "+ 95% confidence interval by year and length group for the combined station and length dataset."), fig.height=12}

length_groups <- stn_ldist %>%
  filter(commonname == filter_sp[2], rho_biomass > 0) %>%
  droplevels() %>%
  pull(length_group) %>%
  levels()

p1 <- stn_ldist %>%
  filter(commonname == filter_sp[2]) %>%
  filter(length_group %in% length_groups) %>%
  droplevels() %>%
  group_by(startyear, length_group) %>%
  reframe(
    value = mean(rho_biomass, na.rm = TRUE),
    se = sd(rho_biomass, na.rm = TRUE) / sqrt(length(!is.na(rho_biomass)))
  ) %>%
  ggplot(aes(
    x = length_group,
    y = value,
    ymin = value,
    ymax = value + 1.96 * se
  )) +
  geom_col(
    color = "black",
    linewidth = 0.3,
    position = "dodge",
    fill = "grey"
  ) +
  geom_errorbar(width = 0.2, linewidth = 0.3) +
  facet_wrap(~startyear, ncol = 6, dir = "v", scales = "free_y") +
  labs(
    y = expression(paste("Average density (kg ", nmi^-2, ")")),
    x = "Length group (cm)"
  ) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  coord_cartesian(expand = FALSE) +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(size = 10, face = 2)
  )

p2 <- stn_ldist %>%
  filter(commonname == filter_sp[2]) %>%
  filter(length_group %in% length_groups) %>%
  droplevels() %>%
  group_by(startyear, length_group) %>%
  reframe(
    value = mean(rho_abundance, na.rm = TRUE),
    se = sd(rho_abundance, na.rm = TRUE) / sqrt(length(!is.na(rho_abundance)))
  ) %>%
  ggplot(aes(
    x = length_group,
    y = value,
    ymin = value,
    ymax = value + 1.96 * se
  )) +
  geom_col(
    color = "black",
    linewidth = 0.3,
    position = "dodge",
    fill = "grey"
  ) +
  geom_errorbar(width = 0.2, linewidth = 0.3) +
  facet_wrap(~startyear, ncol = 6, dir = "v", scales = "free_y") +
  labs(
    y = expression(paste("Average density (n ", nmi^-2, ")")),
    x = "Length group (cm)"
  ) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  coord_cartesian(expand = FALSE) +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(size = 10, face = 2)
  )

cowplot::plot_grid(p1, p2, labels = "AUTO", ncol = 1)
```

```{r bm-density-depth-length, fig.cap = "Bottom depth versus swept area biomass density of each trawl haul (grey dots) allocated to length groups. The blue line and shading represents loess smoother and standard error. The y-axis is log + 1 transformed.", fig.height=12}

length_groups <- stn_ldist %>%
  filter(rho_biomass > 0) %>%
  droplevels() %>%
  pull(length_group) %>%
  levels()

ggplot(
  stn_ldist %>%
    filter(length_group %in% length_groups) %>%
    droplevels(),
  aes(x = bottomdepthstart, y = rho_biomass)
) +
  geom_point(color = "grey", size = 0.3) +
  geom_smooth(fill = "blue") +
  labs(
    x = "Bottom depth (m)",
    y = expression(paste("Density (kg ", nmi^-2, ")"))
  ) +
  facet_grid(length_group ~ commonname) +
  scale_y_continuous(
    trans = "log1p",
    # breaks = c(0,1,10,100,1e3,5e3,1e4,2.5e4,5e4,1e5,1.5e5),
    expand = c(0.01, 0)
  ) +
  coord_cartesian(ylim = c(0, NA)) +
  theme_bw(base_size = 10) +
  scale_x_continuous(breaks = seq(0, 2000, 200), expand = c(0.01, 0))
```

### Maps

```{r, fig.cap = paste("Spatial distribution of", sp_name, "taxa across all years and surveys measured using average biomass density. Size of the pie charts refers to the average total biomass density."), fig.height=10}
# cols <- c("Male" = "#449BCF", "Female" = "#FF5F68", "Unknown" = "grey")
by <- "commonname"
base_size <- 10
legend_lab <- "Redfish taxa"

bm <- basemap(
  stn,
  bathy.style = "rcg",
  legends = FALSE,
  base_size = base_size
) +
  ggnewscale::new_scale_fill() +
  geom_col(
    data = data.frame(
      z = factor(levels(stn_ldist[[by]]), levels(stn_ldist[[by]])),
      x = 1,
      y = 1
    ),
    aes(x = x, y = y, fill = z)
  ) +
  labs(xlab = NULL, ylab = NULL, fill = legend_lab) +
  # scale_fill_manual(values = cols, drop = FALSE) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(hjust = 0.5, size = base_size),
    plot.margin = unit(c(0.2, 0, 0.2, 0), "cm")
  )

roi <- stn %>%
  st_as_sf(coords = c("longitudestart", "latitudestart"), crs = 4326) %>%
  sf::st_transform(attributes(bm)$crs) %>%
  summarise(geometry = st_combine(geometry)) %>%
  concaveman::concaveman(concavity = 3) %>%
  st_make_valid() %>%
  st_buffer(5e4)

grid <- roi %>%
  st_make_grid(n = c(15, 15), what = "polygons") %>%
  st_sf()

grid_clip <- sf::st_intersection(grid, roi)

tmp <- stn_ldist %>%
  filter(!is.na(rho_biomass)) %>%
  group_by(
    cruiseseries,
    startyear,
    cruise,
    platformname,
    serialnumber,
    longitudestart,
    latitudestart,
    bottomdepthstart,
    commonname
  ) %>%
  reframe(value = sum(rho_biomass)) %>%
  st_as_sf(coords = c("longitudestart", "latitudestart"), crs = 4326) %>%
  sf::st_transform(attributes(bm)$crs) %>%
  add_column(grid = as.integer(st_intersects(., grid_clip))) %>%
  group_by(grid, commonname) %>%
  reframe(value = mean(value)) %>%
  group_by(grid) %>%
  mutate(total = sum(value)) %>%
  left_join(
    grid_clip %>%
      st_centroid() %>%
      st_coordinates() %>%
      as_tibble() %>%
      add_column(grid = 1:nrow(.), .before = 1)
  ) %>%
  ungroup() %>%
  mutate(scal = sqrt(total / max(.$total)) * 4e4 + 3e4) %>%
  group_by(grid, X, Y, total, scal)


df.grobs <- tmp %>%
  filter(total > 0) %>%
  do(
    subplots = ggplot(., aes(x = 1, y = value, fill = commonname)) +
      geom_col(
        position = "fill",
        alpha = 0.75,
        colour = "white",
        linewidth = LS(0.5)
      ) +
      coord_polar(theta = "y") +
      # scale_fill_manual(values = cols, drop = FALSE) +
      theme_void() +
      guides(fill = "none")
  ) %>%
  mutate(
    subgrobs = list(annotation_custom(
      ggplotGrob(subplots),
      xmin = (X - scal),
      ymin = (Y - scal),
      xmax = (X + scal),
      ymax = (Y + scal)
    ))
  )

bm + df.grobs$subgrobs
```

```{r, fig.cap = "Spatial distribution of beaked redfish size classes across all years and surveys measured using average biomass density. Size of the pie charts refers to the average total biomass density.", fig.height=10}

length_groups <- stn_ldist %>%
  filter(commonname == "snabeluer", rho_biomass > 0) %>%
  droplevels() %>%
  pull(length_group) %>%
  levels()

by <- "length_group"
base_size <- 10
legend_lab <- "Size class"

bm <- basemap(
  stn,
  bathy.style = "rcg",
  legends = FALSE,
  base_size = base_size
) +
  ggnewscale::new_scale_fill() +
  geom_col(
    data = data.frame(z = factor(length_groups, length_groups), x = 1, y = 1),
    aes(x = x, y = y, fill = z)
  ) +
  labs(xlab = NULL, ylab = NULL, fill = legend_lab) +
  scale_fill_viridis_d() +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(hjust = 0.5, size = base_size),
    plot.margin = unit(c(0.2, 0, 0.2, 0), "cm")
  )

roi <- stn %>%
  st_as_sf(coords = c("longitudestart", "latitudestart"), crs = 4326) %>%
  sf::st_transform(attributes(bm)$crs) %>%
  summarise(geometry = st_combine(geometry)) %>%
  concaveman::concaveman(concavity = 3) %>%
  st_make_valid() %>%
  st_buffer(5e4)

grid <- roi %>%
  st_make_grid(n = c(15, 15), what = "polygons") %>%
  st_sf()

grid_clip <- st_intersection(grid, roi)

tmp <- stn_ldist %>%
  filter(!is.na(rho_biomass)) %>%
  filter(commonname == "snabeluer") %>%
  filter(length_group %in% length_groups) %>%
  droplevels() %>%
  group_by(
    cruiseseries,
    startyear,
    cruise,
    callsignal,
    serialnumber,
    longitudestart,
    latitudestart,
    bottomdepthstart,
    length_group
  ) %>%
  summarise(value = sum(rho_biomass)) %>%
  st_as_sf(coords = c("longitudestart", "latitudestart"), crs = 4326) %>%
  sf::st_transform(attributes(bm)$crs) %>%
  add_column(grid = as.integer(st_intersects(., grid_clip))) %>%
  group_by(grid, length_group) %>%
  reframe(value = mean(value)) %>%
  group_by(grid) %>%
  mutate(total = sum(value)) %>%
  left_join(
    grid_clip %>%
      st_centroid() %>%
      st_coordinates() %>%
      as_tibble() %>%
      add_column(grid = 1:nrow(.), .before = 1)
  ) %>%
  ungroup() %>%
  mutate(scal = sqrt(total / max(.$total)) * 4e4 + 3e4) %>%
  group_by(grid, X, Y, total, scal)

df.grobs <- tmp %>%
  filter(total > 0) %>%
  do(
    subplots = ggplot(., aes(x = 1, y = value, fill = length_group)) +
      geom_col(
        position = "fill",
        alpha = 0.75,
        colour = "white",
        linewidth = LS(0.5)
      ) +
      coord_polar(theta = "y") +
      scale_fill_viridis_d(drop = FALSE) +
      theme_void() +
      guides(fill = "none")
  ) %>%
  mutate(
    subgrobs = list(annotation_custom(
      ggplotGrob(subplots),
      xmin = (X - scal),
      ymin = (Y - scal),
      xmax = (X + scal),
      ymax = (Y + scal)
    ))
  )


bm + df.grobs$subgrobs
```

```{r, fig.cap = "Spatial distribution of golden redfish size classes across all years and surveys measured using average biomass density. Size of the pie charts refers to the average total biomass density.", fig.height=10}
length_groups <- stn_ldist %>%
  filter(commonname == "vanlig uer", rho_biomass > 0) %>%
  droplevels() %>%
  pull(length_group) %>%
  levels()

by <- "length_group"
base_size <- 10
legend_lab <- "Size class"

bm <- basemap(
  stn,
  bathy.style = "rcg",
  legends = FALSE,
  base_size = base_size
) +
  ggnewscale::new_scale_fill() +
  geom_col(
    data = data.frame(z = factor(length_groups, length_groups), x = 1, y = 1),
    aes(x = x, y = y, fill = z)
  ) +
  labs(xlab = NULL, ylab = NULL, fill = legend_lab) +
  scale_fill_viridis_d() +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(hjust = 0.5, size = base_size),
    plot.margin = unit(c(0.2, 0, 0.2, 0), "cm")
  )

roi <- stn %>%
  st_as_sf(coords = c("longitudestart", "latitudestart"), crs = 4326) %>%
  sf::st_transform(attributes(bm)$crs) %>%
  summarise(geometry = st_combine(geometry)) %>%
  concaveman::concaveman(concavity = 3) %>%
  st_make_valid() %>%
  st_buffer(5e4)

grid <- roi %>%
  st_make_grid(n = c(15, 15), what = "polygons") %>%
  st_sf()

grid_clip <- st_intersection(grid, roi)

tmp <- stn_ldist %>%
  filter(!is.na(rho_biomass)) %>%
  filter(commonname == "vanlig uer") %>%
  filter(length_group %in% length_groups) %>%
  droplevels() %>%
  group_by(
    cruiseseries,
    startyear,
    cruise,
    callsignal,
    serialnumber,
    longitudestart,
    latitudestart,
    bottomdepthstart,
    length_group
  ) %>%
  summarise(value = sum(rho_biomass)) %>%
  st_as_sf(coords = c("longitudestart", "latitudestart"), crs = 4326) %>%
  sf::st_transform(attributes(bm)$crs) %>%
  add_column(grid = as.integer(st_intersects(., grid_clip))) %>%
  group_by(grid, length_group) %>%
  reframe(value = mean(value)) %>%
  group_by(grid) %>%
  mutate(total = sum(value)) %>%
  left_join(
    grid_clip %>%
      st_centroid() %>%
      st_coordinates() %>%
      as_tibble() %>%
      add_column(grid = 1:nrow(.), .before = 1)
  ) %>%
  ungroup() %>%
  mutate(scal = sqrt(total / max(.$total)) * 4e4 + 3e4) %>%
  group_by(grid, X, Y, total, scal)

df.grobs <- tmp %>%
  filter(total > 0) %>%
  do(
    subplots = ggplot(., aes(x = 1, y = value, fill = length_group)) +
      geom_col(
        position = "fill",
        alpha = 0.75,
        colour = "white",
        linewidth = LS(0.5)
      ) +
      coord_polar(theta = "y") +
      scale_fill_viridis_d(drop = FALSE) +
      theme_void() +
      guides(fill = "none")
  ) %>%
  mutate(
    subgrobs = list(annotation_custom(
      ggplotGrob(subplots),
      xmin = (X - scal),
      ymin = (Y - scal),
      xmax = (X + scal),
      ymax = (Y + scal)
    ))
  )


bm + df.grobs$subgrobs
```

<!-- ### Length group maps -->

<!-- ```{r, fig.cap = "Annual distribution of 1-17 cm (approximate age class 1) Atlantic halibut abundance density scaled to symbol color and size.", fig.height=15} -->
<!-- x <- stn_ldist %>% -->
<!--   filter(length_group == levels(stn_ldist$length_group)[1]) %>% -->
<!--   group_by(cruiseseries, startyear, platformname, serialnumber, longitudestart, latitudestart) %>% -->
<!--   reframe(value = sum(rho_abundance)) -->

<!-- bubble_map(x, label = expression(paste("Density (n ", nmi^-2, ")")), ncol = 7) -->
<!-- ``` -->

<!-- ```{r, fig.cap = "Annual distribution of 18-27 cm (approximate age class 2) Atlantic halibut abundance density scaled to symbol color and size.", fig.height=15} -->
<!-- x <- stn_ldist %>% -->
<!--   filter(length_group == levels(stn_ldist$length_group)[2]) %>% -->
<!--   group_by(cruiseseries, startyear, platformname, serialnumber, longitudestart, latitudestart) %>% -->
<!--   reframe(value = sum(rho_abundance)) -->

<!-- bubble_map(x, label = expression(paste("Density (n ", nmi^-2, ")")), ncol = 7) -->
<!-- ``` -->

<!-- ```{r, fig.cap = "Annual distribution of 28-47 cm Atlantic halibut abundance density scaled to symbol color and size.", fig.height=15} -->
<!-- x <- stn_ldist %>% -->
<!--   filter(length_group == levels(stn_ldist$length_group)[3]) %>% -->
<!--   group_by(cruiseseries, startyear, platformname, serialnumber, longitudestart, latitudestart) %>% -->
<!--   reframe(value = sum(rho_abundance)) -->
<!-- bubble_map(x, zero_color = "grey", label = expression(paste("Density (n ", nmi^-2, ")")), ncol = 7) -->
<!-- ``` -->

<!-- ```{r, fig.cap = "Annual distribution of 48-70 cm Atlantic halibut abundance density scaled to symbol color and size.", fig.height=15} -->
<!-- x <- stn_ldist %>% -->
<!--   filter(length_group == levels(stn_ldist$length_group)[4]) %>% -->
<!--   group_by(cruiseseries, startyear, platformname, serialnumber, longitudestart, latitudestart) %>% -->
<!--   reframe(value = sum(rho_abundance)) -->

<!-- bubble_map(x, zero_color = "grey", label = expression(paste("Density (n ", nmi^-2, ")")), ncol = 7) -->
<!-- ``` -->

<!-- ```{r, fig.cap = "Annual distribution of >71 cm Atlantic halibut abundance density scaled to symbol color and size.", fig.height=15} -->
<!-- x <- stn_ldist %>% -->
<!--   filter(length_group == levels(stn_ldist$length_group)[5]) %>% -->
<!--   group_by(cruiseseries, startyear, platformname, serialnumber, longitudestart, latitudestart) %>% -->
<!--   reframe(value = sum(rho_abundance)) -->

<!-- bubble_map(x, zero_color = "grey", label = expression(paste("Density (n ", nmi^-2, ")")), ncol = 7) -->
<!-- ``` -->